### 2016-04-07 paper/gen-graph-connected-heatmap.tex
Include generated stuff for now, to make it easier to build

\begin{tikzpicture}[gnuplot]
%% generated with GNUPLOT 5.0p0 (Lua 5.2; terminal rev. 99, script rev. 100)
%% Thu 07 Apr 2016 11:12:14 BST
\tikzset{every node/.append style={font={\scriptsize}}}
% \path (0.000,0.000) rectangle (10.922,7.366);
\gpcolor{color=gp lt color border}
\gpsetlinetype{gp lt border}
\gpsetdashtype{gp dt solid}
\gpsetlinewidth{1.00}
\draw[gp path] (1.320,5.785)--(1.320,2.197)--(4.908,2.197);
\node[gp node center,rotate=-270] at (0.246,3.991) {Association Runtime (ms)};
\node[gp node center] at (3.114,1.427) {MAC Both Runtime (ms)};
\begin{scope}
\clip (1.320,5.785) rectangle (4.908,2.197);
\def\gprawrgbimagedata{%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff846eff6f65ff846eff7e6cff6f65ff8c72ff7367ff7e6cff8c72ff7869ffffffffac88ff9576%
  ff9576ff7869ff8c72ffac88ff846eff9576ff9576ff9f7dff8c72ff9f7dff8c72ff9576ffac88ff9f7dff8c72ff8c72%
  ff8c72ff9576ff8c72ff9f7dff8c72ffffffff9f7dff9576ff9f7dff9576ff9576ff9576ff7e6cff9f7dff9f7dffffff%
  ff9f7dffac88ff9f7dffffffff7e6cffac88ffffffffffffffffffff9f7dff9f7dffffffffffffffac88ffac88ffac88%
  ff9576ffac88ffffffffac88ffffffffffffff9f7dffac88ff9f7dffffffff9f7dffac88ff9f7dff9576ff6f65ff7869%
  ff7e6cfc6260fc6260ff6b63ac0527ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffffffffffffff9f7dffac88%
  ffffffffffffffac88ffac88ffffffffac88ff9f7dffffffff9f7dee4156ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffac88ffac88ffac88ffac88ffffffffac88ff9f7ded4155ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ffac88ff9f7d%
  ff8c72ff9f7dff9f7dec3f55ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffff9576ffffffff9f7dffffff%
  ffac88ffffffffac88ff9f7dff9f7dff9576ff9f7dffac88e93c53ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffac88ffac88ffac88ffffffff9f7dff8c72ff9576ff9f7dff9f7dff9f7dffffffffffffec4055ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffac88ffac88ffffffffac88ffffffffac88ffffffff7e6cffac88ffac88ff7e6cff9f7d%
  ff9f7dff9576e33450ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffac88ff9f7dffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffac88ffffffffffffffac88ffffffffac88ffac88ffffffffffffff9576%
  ffac88ff846effac88ffac88ffac88ffac88ff9576ec3f55ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffac88ffac88ffffffffffffffac88ffffffffac88ff9f7dffac88ff9f7dffac88ffac88%
  ff9f7dffffffffffffffffffffac88ffac88ffac88ff9f7dff9576ffac88ffac88ffac88ee4156ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffac88ffffffffffffff9576%
  ffac88ffffffff9f7dff9f7dffac88ffac88ff9f7dff9f7dff9576ffffffff9f7dff9576ffffffff9576ffffffff9f7d%
  ffac88e1314fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffac88ffac88ff9f7dff9f7dffac88ff8c72ff7e6c%
  ff9576ffffffff9576ff9576ff846eff9576de2e4dffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88%
  ffac88ffffffffffffffffffff9f7dffac88ff9f7dff9f7dffffffff9f7dff9f7dff8c72ffac88ffffffff9576ff9f7d%
  ff9f7dff8c72ffffffff9576ffffffff8c72ffac88ffffffffffffff8c72ff9576dd2d4dffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffac88ffffffffffffff9f7dff9f7dffac88%
  ffffffffac88ff9576ffffffffac88ff9576ffac88ff9576ffac88ff9f7dff846effac88ff8c72ff846eff8c72ff8c72%
  d62449ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffffffffac88ffffffffac88ffffffffffffffffff%
  ffac88ffac88ffffffffac88ffffffffac88ff9576ff9f7dffac88ff9f7dffac88ffac88ff9f7dffac88ffac88ff9f7d%
  ff9f7dff9576ff9f7dff8c72ffac88da294bffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffac88ff9576ffac88ffac88%
  ffffffff9f7dff8c72ffac88ff9f7dffac88ff846effac88ff846efe6862da294bffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffff9f7dffffffffffffffac88ffac88ffac88ff9f7dffffffff8c72%
  ffac88ffffffffffffff9f7dffffffff9f7dff9f7dffac88ff8c72ff8c72ffffffff9576ff7869ff7367ff8c72d7264a%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffac88ffac88ffac88ffac88ffffff%
  ffffffffac88ffffffffac88ffffffff9f7dffac88ff9f7dff9f7dffac88ffffffff9f7dff9f7dff7e6cff9576ff9f7d%
  ff9576ff7869ff7869ff6f65d62449ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffff%
  ffac88ffffffffffffffffffff9f7dffffffffac88ffffffffffffffac88ff9f7dff9f7dff9f7dff9f7dff9f7dffffff%
  ff9f7dff9576ff8c72ff9576ff9576ff7869ff7e6cf6545cfc6260d52349ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffff%
  ffac88ffffffffac88ffac88ffffffffffffffffffffffffffffffff9f7dffffffffffffff8c72ff9f7dffac88ff9576%
  ffac88ff9576ff8c72ff9f7dff9f7dff9f7dff8c72ff9576ff6f65ff846eff6f65ff6b63fd6561fd6561d21f47ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffac88ffffff%
  ffffffffffffffffffffffffff9f7dff9f7dff8c72ff9f7dffffffffac88ff9576ff7e6cff846eff846efe6862ff6b63%
  fe6862f7565cfc6260d8264affffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffff9576ffffffffac88ffffffffac88%
  ffffffffffffff9576ffffffffac88ffac88ff9f7dffffffffac88ff8c72ffffffff9576ffac88ffac88ffac88ff7367%
  ff6f65ff7e6cfe6862fe6862f7585dfb605ffe6862f95c5ed32148ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ff9576ffac88ff9f7dffffffff9f7d%
  ff9f7dff9f7dff9576ff7e6cff8c72ff9576ff6f65ff6b63ff6b63fb605ffd6561ff6b63f85a5dda294bffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ff9f7dff9f7dffac88ffac88%
  ffffffffffffff9f7dff9f7dff9f7dffac88ff9f7dff9576ff6f65ff8c72ff7e6cff6b63fe6862ff6f65f34d59fe6862%
  f85a5dff6f65d7254affffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffffffffffff%
  ffac88ffac88ff9f7dffac88ffac88ff9576ff8c72ff9576ffac88ff9576ff846eff9576ff8c72ff846efa5e5ffb605f%
  fa5e5ff7585dfb605ff7565cf85a5dfe6862fc6260d8264affffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffffffff9f7dffffffffffff%
  ffffffff9576ffffffffac88ffffffffffffffac88ffac88ffac88ff8c72ff9f7dff8c72ffac88ff8c72ff846eff7869%
  ff7e6cfd6561fd6561fd6561fa5e5ffd6561fa5e5fff6b63f85a5dfd6561fb605fff7e6ce1314fffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffff9f7dffac88ff9576ffac88ffac88ffffffffac88ffffffff9f7dff7e6cff9576%
  ff7367ffac88ff7869ff7869fe6862ff6b63fc6260fd6561ff7367fe6862fc6260fe6862f85a5df7585dfe6862ff6f65%
  ff6b63e73952ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ffac88ffffffffffffffffffffffffffac88ffffffff9f7dffac88ffac88ff9f7dff9576%
  ffffffff9576ffac88ff8c72ff846eff9576ff846eff6f65ff846eff846eff7367f85a5dfb605ffc6260f7585df7565c%
  fd6561ff7367ff6f65fe6862ff7e6cff846eef4356ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffac88ffffffffac88ffffffffac88ffffff%
  ffac88ff9f7dff9576ffffffffffffff8c72ffac88ff9576ff9576ff9576ff9576ff9576ff9f7dff7e6cff846efe6862%
  ff7869ff8c72fe6862ff6b63ff7367ff7869ff8c72ff846eff846eff9576ff9f7de83b53ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffac88%
  ffffffffac88ffac88ffffffffffffffffffff9576ff9f7dff8c72ff9576ff9f7dff846effac88ff9576ff8c72ffac88%
  ff7e6cff7e6cff8c72ff8c72ff6b63ff846eff6f65ff7869ff7367ff9f7dff9576ff7e6cff846eff9f7dff8c72ff8c72%
  ed4155ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffac88ffffffffac88ff9f7dffac88ffac88ff9f7dff9576ff846eff9f7dffac88ff8c72ffac88%
  ff9576ff9576ff9f7dff846eff8c72ff7e6cfe6862ff6f65fc6260ff9576ff8c72ff846eff7e6cff9576ff9f7dff9576%
  ffffffff8c72ffac88ffac88ffac88f24a59ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffac88ffac88ffffffff8c72ffffffff9f7dff9576%
  ff8c72ff9f7dffac88ffac88ff9f7dff846eff7869ff9f7dff7e6cff7e6cff7367ff6f65fd6561ff7e6cfe6862fe6862%
  ff9576ff8c72ff9f7dff846effffffff9f7dff9f7dff9f7dffac88ffffffe93c53ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ff9f7dff9f7d%
  ffffffff9f7dff9576ff9f7dffac88ffac88ff9576ff8c72ff9f7dff9f7dff846efe6862ff6b63fd6561ff7367f95c5e%
  fe6862ff6f65ff846eff7e6cff7869ff6b63ff846eff9f7dff9f7dff846eff8c72ff9f7dff8c72ffffffff9f7df34d59%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffac88ffac88ffac88ffffffffac88ffffffffac88ff9f7dff7e6cff8c72ff9576ff8c72ff846eff8c72ff8c72%
  fd6561ff7e6cfc6260ff846eff7869fd6561f95c5ef95c5eff6f65ff7e6cff9f7dff846eff9576ff9f7dff9576ffac88%
  ffac88ff8c72ff9576ff9f7def4457ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffff9f7dffffffffffffffffffffac88ffac88ffac88ff9f7dff9f7dff9576ff8c72ff9f7d%
  ff8c72ff7367ff8c72fe6862ff7367ff6b63fb605ffc6260fe6862f85a5dfd6561f7565cff6f65fe6862ff6f65ff7869%
  ff8c72ff9576ff8c72ff9576ffac88ffffffffac88ffac88ff9f7df14858ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ff9f7dffffffffac88ffac88%
  ff8c72ff7869ff9f7dff9f7df95c5eff7869fe6862fd6561ff7367fa5e5ff95c5ef7565cf5515bf6545cf7585dfc6260%
  fb605fff6f65fc6260ff6f65ff6b63ff846eff8c72ff9f7dff9576ffac88ffac88ffac88ff9576ffac88ef4356ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffac88%
  ff9576ffac88ffac88ff7367ff846eff7367fe6862ff6f65fb605ffd6561f95c5eff6f65f7585dff6b63f95c5ef85a5d%
  f7565cf24a59f5535bf4505aff6b63f7585dfa5e5fff6f65ff6b63ff7e6cff8c72ff8c72ffac88ff9f7dff9f7dff8c72%
  ffffffff8c72ff846ef14958ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffffff%
  ffac88ffffffffffffffffffffac88ffac88ffac88ff9576ff9576ff7e6cff8c72ff7367f95c5eff7869f85a5dfb605f%
  fb605ff85a5df85a5dfc6260f6545cfa5e5ff7585dfd6561fa5e5ffc6260ff6b63ff7869ff7e6cff8c72fd6561ff846e%
  ffac88ffac88ff9f7dffac88ff9576ff9f7dff9576ff846ef7585dffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffff9f7dffac88ffffffffffffff846eff8c72ff7e6cff846eff7e6c%
  ff6f65fb605ffe6862ff6b63fc6260f95c5efb605ff85a5df7585dfa5e5ffd6561fe6862fb605ffd6561fe6862ff6b63%
  ff6f65ff846eff7e6cff9f7dff8c72ff9576ff9f7dffac88ff9576ff9576ffac88ff9f7dffac88f24b59ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffffffff9f7dff846effffff%
  ff846eff9576ff8c72ff7869ff7e6cff7367ff7367fa5e5ffc6260f7585dfb605ffb605ff95c5efe6862fc6260fb605f%
  ff7869ff6b63fe6862ff7367ff7869ff8c72ff9576ff7e6cff9576ff9576ff9f7dff9f7dff9f7dff8c72ff8c72ff7869%
  ff846eff9576f04557ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88%
  ffac88ffffffffac88ff9f7dff8c72ff9576ff846eff7367ff6f65ff7e6cf95c5efc6260fe6862fd6561fc6260fb605f%
  f95c5ef95c5ef95c5eff6f65fd6561ff8c72fd6561ff7e6cff6f65ff846eff8c72ff7869ff8c72ff9576ff9f7dffffff%
  ff846eff8c72ff9f7dffac88ff9f7dff846eff9f7df5515bffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffac88ff9576ffac88ffac88ff9576ff846eff846eff7e6cff7367fb605ffe6862ff6b63%
  fc6260fd6561ff846efb605fff8c72ff7e6cfb605fff6f65ff7869ff8c72ff7869ff6f65ff846eff6f65ff7e6cff846e%
  ff9576ff9576ff9576ff9576ff9576ff846eff846eff9576ff846eff9576ff9576ff9f7df24a59ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffff9f7dff9576ffffffff9576ff9f7dff846efe6862%
  f7565cff6f65f7565cf95c5eff7367fc6260fa5e5fff7367fe6862ff7869ff7869ff7e6cff7e6cff846eff7367ff9576%
  ff7e6cff7e6cff8c72ff846eff846eff9576ff7869ff7869ff8c72ff8c72ff9f7dff846eff9576ff7869ffffffff9576%
  ff9576e93c53ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffff9f7dff9f7dffac88ffac88%
  ff9f7dff9576ff9f7dff8c72fe6862fd6561fd6561fc6260ff7367ff7367ff7869ff7869fd6561ff6b63ff6b63ff7367%
  ff7e6cff846eff7367ff7869ff846eff7869ff7869ff846eff9f7dff7367ff8c72ff8c72ff8c72ff9576ff9f7dff9576%
  ff9f7dff8c72ff9576ff9576ff9f7dff9f7df34d59ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffffffff9f7dffac88%
  ffac88ff9f7dffffffffac88ffac88ffac88ff9f7dff8c72ff846eff7e6cff9576f85a5dff6f65ff6b63ff6b63ff7869%
  fc6260fd6561fe6862ff6b63fd6561ff6b63ff846eff7e6cff9576ff9f7dff8c72ff9576ff8c72ff7869ff8c72ff7e6c%
  ff8c72ffac88ff9f7dffac88ffac88ff9576ff9576ff9576ff8c72ff9576ffac88f4505affffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffac88ffac88ffffffffffffffac88ff8c72ff9f7dff8c72ff7367ff846eff846eff7e6cff7869ff8c72%
  fc6260ff7e6cff7367fb605fff7869ff7e6cff7367ff7e6cff7367ff7869ff8c72ff7e6cff8c72ffac88ff8c72ff7869%
  ff9576ff7367ff846eff7367ff7367ff7e6cff8c72ffac88ffac88ff7869ff846effffffff9576ff9576ff9f7dff9576%
  f14958ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88%
  ffac88ffffffffffffffffffffac88ffffffffac88ffffffffac88ff9f7dffac88ff8c72ff846eff9576ff9576ff9576%
  ff7869ff6b63ff846eff7869ff7e6cff7e6cfb605fff7367ff7e6cff9576ff8c72ff9576ff8c72ff8c72ff9576ff7367%
  ff6f65ff6f65ff7367ff8c72ff9576ff9576ff9f7dff7e6cff846eff7e6cff7e6cff846effac88ff9576ff9576ff9f7d%
  ff8c72ff9f7dffffffffac88ff9576ef4457ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffff9f7dffffffff9f7dffffffffffffffffffffac88ffffffff9f7dffac88ff9f7d%
  ff9f7dff846eff846eff8c72ff7367ff6f65ffac88ff6b63ff7869ff7869ff8c72ff9f7dff9f7dff7e6cff7e6cff8c72%
  ff8c72ff9576fe6862ffac88ff7869ff8c72ffac88ff7e6cff9576ff846eff7e6cff8c72ff7e6cff8c72ff7367ffac88%
  ff8c72ff8c72ff7869ff8c72ff9f7dff9576ffac88ff9f7dffac88ff846ef24a59ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffac88%
  ffac88ff846eff9576ff8c72ff9576ff9f7dff9f7dff846eff7e6cff7367ff7e6cff7869ff7e6cffac88ff9576ff6f65%
  ff9576ff7e6cff846eff9576ff7e6cff7e6cff7e6cff846eff7e6cff8c72ff8c72ff7869ff9f7dffac88ff9f7dff9f7d%
  ff8c72ff9576ff846eff9576ff7869ff8c72ff6f65ff846eff9576ff846eff9576ffac88ff846eff9f7dffac88f14958%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffffffffac88%
  ffffffffffffffffffff9f7dffac88ffac88ffffffffac88ff8c72ff846eff9f7dff8c72ff8c72ff7e6cff7869ff7869%
  ff8c72ff6f65ff846eff7e6cff6f65ff7e6cff8c72ff846efb605fff7e6cff6b63ffac88ff846eff9576ff846eff8c72%
  ff9576ff8c72ff7869ff9f7dff9f7dff846eff8c72ff9f7dff8c72ff8c72ff9576ff846eff846eff9576ff9576ff9f7d%
  ff9576ff9f7dff7e6cff9f7df24a59ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffac88ff9576ffffffff9f7dffac88ffffffffac88ff9f7dff8c72ff9576ffac88ff8c72%
  ff9576ff7367ff7367ff7869ff9576ff8c72ff9576ff8c72ff7e6cff846efe6862ff7869ff7367ffac88ff6f65ff7e6c%
  ff7e6cff7e6cff846eff7e6cff846eff8c72ff8c72ff7869ff9576ff846eff846eff8c72ff846eff7e6cff8c72ff9576%
  ff9576ff9576ff9576ffac88ff9576ff8c72ffac88ff846efffffff85a5dffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffffffff9576ff9576ffffffff846eff9f7dffac88%
  ff7869ff9576ff7e6cfe6862ff7869fe6862ff7367fd6561ff9576ff7367ff7367ff7367ff846eff9576fa5e5fff7e6c%
  fb605fff7367fe6862ff846eff6f65ff846eff8c72fd6561ff8c72ff7869ff846eff6f65ffffffff9f7dff8c72ff846e%
  ff7e6cff7869ff9576ff9576ff8c72ff9576ff8c72ff9f7dff7e6cff846eff9f7dff9f7dff8c72ffac88f85a5dffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ff9f7dffffffffffffffac88%
  ff9576ffffffff9576ff8c72ffac88ff7367ff7869ff9576fc6260ff6b63fe6862ff846eff846eff846eff6f65ff7869%
  ff6f65ff8c72ff9f7dff7e6cff846efd6561ff6b63ff846eff7869ff7e6cff7e6cff9f7dff7869ff846eff7e6cff846e%
  ff8c72ff846eff9576ff9f7dff9576ff6f65ff9f7dff846eff9f7dff7e6cffac88ff9576ff9576ff8c72ff8c72ffffff%
  ffac88ffac88ff9f7df24b59ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ff9576ff846effffffff9576ff9576ff9576ffac88ff846eff846eff7e6cff6b63ff7367fe6862ff7869ff6f65%
  ff7e6cff7869ff7e6cff7869ff846effac88ff846eff7367ff9576fd6561ff7869ff7367ff6f65ff7869ff7869ff846e%
  ff846effac88ff7367ff8c72ff846eff8c72ff7367ffac88ff846eff7e6cff846eff7869ff7869ff846eff846eff9576%
  ff9f7dff8c72ffac88ffffffffac88ff9576ff846eff9576f7565cffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffac88ffac88ffffffffffffffffffffac88ffac88ff9576ffffffff9576ff8c72ff9576ffac88fa5e5f%
  ff846eff7367ff7e6cff846eff9f7dff6f65ff7367ff9576ff9f7dff9576ff7869ff7869fd6561ff7367fb605fff846e%
  fe6862ff9576ffac88ff7367ff7e6cff7869ff7e6cff9576ff9576ff7869ff6f65ff846eff6f65ff9576ff7869ff7e6c%
  ff846eff846eff6b63ff7e6cff9576ff8c72ff9f7dffac88ff9576ff7869ff9f7dffac88ff9576f34e5affffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff9f7dffac88ffac88ffffffff9f7dffffffff9f7dffffffff9f7dff9576ff846eff9576ff8c72%
  ff9576ff9f7dff846eff846eff7e6cff9f7dff8c72ffac88ff7869ff7e6cff9f7dff9f7dff9576ff846eff8c72ff6f65%
  ff7367ff6b63ff7367ff846eff6b63fe6862ff846eff8c72ff6b63ff9576ff9f7dff9576ff7869ff846eff9576ff846e%
  ff846eff7869ff8c72ff7869ff6f65ff7e6cff8c72ff9576ff9576ff8c72ff9576ffac88ff9576ff9f7dff8c72ff846e%
  ff9576ffac88f5535bffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffffffff9f7dff8c72ffffffffac88ffffffff9f7d%
  ffac88ff8c72ff9f7dff9576ff9576ff846eff846eff7e6cff7869ff8c72ff7869ff9f7dff846eff7e6cff9576ff9f7d%
  ff8c72ffac88ff9f7dff846eff7367ff6b63ff846eff7869ff6f65ff7367ff9576ff846eff7367ff846eff8c72ff7e6c%
  fe6862ff6b63ff9576ff7e6cff7e6cffac88ffffffff846eff7e6cff8c72fe6862ff6f65ff7e6cffffffff8c72ff846e%
  ff9576ff846eff9576ff8c72ff9576ffffffffac88f24b59ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffac88%
  ffffffff8c72ffac88ff9f7dffac88ff9f7dff9f7dff7e6cff9f7dff7869ff846eff7e6cff9f7dff7367ff9576ff7869%
  ff7e6cff8c72ff6b63ff846eff7869ff9f7dff7e6cff846eff8c72ff7367ff7e6cff9f7dff7367ffac88ff7869ff846e%
  ffffffff8c72ff8c72ff8c72ff846effffffff7869ff9576ff7367ff7367ff9576fd6561ff7e6cff846effac88ff9f7d%
  ff8c72ff9576ff9576ff9f7dff6f65ff9576ffac88ff7869ff846eff9f7dff9576ff9576f6545cffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffff7367ff7e6cff9576ffffffff8c72ff8c72ff7e6c%
  ff7e6cff7869ff7e6cff846eff6b63ff7367ff7869ff7869ff7869ff7367ff9576ff846eff846eff846eff8c72fe6862%
  ff6b63ff7869ff846eff8c72ff9576ff9576ff846eff9576ff7e6cff846eff846eff846eff7367ff9576ff7e6cff6b63%
  ff846eff8c72ff9f7dff8c72ff9576ff846eff9576ff9f7dff7e6cff846eff9f7dff9f7dff8c72ff7869ff9576ffac88%
  ffac88f04557ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffac88ffffffffac88ffffffffac88ff9576ffffffffac88ff9576ff9f7dff9576ff9576%
  ff9f7dff9f7dff8c72ffac88ff9576ff7869ff7869ff7869fd6561ff7869ff6b63ff7869fb605fff846eff7869ff8c72%
  ff7e6cff8c72ff846eff7e6cff8c72ff6f65ff7e6cff9576ff9576ff9576ff9f7dff846eff7869ff846eff9576ff7869%
  f85a5dff846eff7869ff7869ff846eff846eff7869ff846eff6f65ff8c72ffac88ff846eff9f7dffac88ff8c72ff8c72%
  ff9f7dff9f7dff9576ff9f7dffac88fffffff95c5effffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffff9f7dffffffffac88ffffff%
  ff9f7dff846eff8c72ff9f7dff6f65ff9576ff7367fe6862ff7e6cff846eff7869ff6b63ff7367fd6561f85a5df85a5d%
  ff6b63ff7367ff7367ff7e6cff7869ff8c72ff8c72ff846eff8c72ff8c72ff846eff846eff8c72ff846eff7e6cff6b63%
  ff6f65ff7367ff9576ff7869ff6f65ff9576ff9576ff7e6cff846eff7367ff8c72ff7e6cff8c72ff9f7dff8c72ff9576%
  ff9576ff9576ff846eff9f7dff9576ff8c72ffac88ff8c72ff8c72ff8c72ffac88ff6b63ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffac88ffac88ff8c72ffffffffac88ff9576fe6862ff9576ff6f65ff9576ff7367ff7e6cff6f65ff6f65%
  f7585df7585dfd6561fa5e5fff7367ff7869ff6f65ff6f65ff846efe6862ff846eff9576ff8c72fe6862ff7869ff7367%
  ff7e6cff9f7dff7e6cff9576ff8c72ff8c72ff8c72ff7869ff9f7dff846eff6b63ff6f65ff7367ff7869fb605fff846e%
  ff6b63ff8c72ff9f7dff9576ff7e6cff7e6cff846eff7367ff7e6cff7e6cff9576ff8c72ff8c72ff9f7dff846eff9f7d%
  fa5e5fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ff9f7dffac88ff9576ff9576ff846eff8c72ff7869ff9576ff7367%
  ff846eff7e6cff6f65fb605ff5515bf4505af4505af5535bf95c5eff6f65ff7869fe6862ff7869ff6f65ff846eff7e6c%
  ff846eff6f65ff7e6cff846eff7367ff7e6cff9f7dffac88ff846eff9f7dff846eff7e6cff7e6cff846eff9576fd6561%
  fd6561ff9576ff7367ff7869ff7869ff8c72ff846eff846eff7e6cffac88ff7e6cff9f7dff9f7dff846eff8c72ff9f7d%
  ff7e6cff8c72ff9576ff9576fffffffe6862ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffff9f7dffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffff9f7dffffffffffffff9f7dff9576ffffffff9576%
  ff9576ff7367ff846eff7869fa5e5ffa5e5ffe6862fa5e5ff24b59f5535bf4505af34d59fc6260f7565cff7e6cfc6260%
  ff6f65fe6862ff7869ff8c72ff7367ff846eff9576ff8c72ff9576ff846eff846eff6b63ff7367ff9f7dff6f65ff8c72%
  ff6f65ff7869ff846eff846eff7367ff7e6cff7e6cff7869ff8c72ff846eff7869ff9f7dff9576ff846eff7869ff846e%
  ffac88ff9f7dffac88ff8c72ff8c72ff9576ff846eff9576ff8c72ff9576ff7367ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffff9f7dffffffff9f7dffac88%
  ffffffffffffffffffffac88ffac88ffac88ffffffffac88ffffffff9f7dffac88ffffffff9f7dff9f7dff9576ffac88%
  ffffffff9f7dff8c72ffffffff8c72ff7869ff8c72ff8c72fa5e5ff85a5dff6f65fa5e5ff85a5df24b59fa5e5ff5535b%
  f14958fb605ff85a5dfc6260ff7e6cff6b63ff6b63ff7869ff7e6cff7367ff9f7dff7367ff7e6cff846eff6f65ff7869%
  ff7869ff9576fe6862ff7869ff7869ff7e6cff7e6cff6f65ff7367ff7367ff8c72ff846eff6b63ffac88ffffffff846e%
  ff7869ff7e6cffac88ff9576ff9576ffffffffffffffac88ffac88ffac88ff9f7dff9576ffac88ffffffffac88ff7869%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7d%
  ff7e6cff9576ff8c72fc6260ff7367ff9f7dff7367ff9576ff9f7dff9576ff7869ff8c72ff9f7dff8c72ffffffff8c72%
  ff7869ff8c72ff9f7dff9576ff9576ff9576ff8c72ff9576ff9576fd6561ff8c72ff9f7dff7869fb605ff85a5dfc6260%
  f5515bf4505aee4256f5535bf34d59f04757f34e5af6545cf85a5dfb605ffd6561fb605fff7e6cff7869ff7e6cfe6862%
  ff7869ff7e6cff7367ff7367fe6862fe6862ff9576ff6b63ff6f65fc6260ff7869ff6b63ff7367ff6f65ff7869ff7869%
  ff7869ff846eff7367ffac88fe6862ff7367ff8c72ff9f7dff846eff9576ffac88ff9f7dff9576ff9576ff9f7dffffff%
  ffffffff9576ff9576ffffffff6b63ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffff7367fe6862ff7367f95c5efc6260f95c5ef95c5eff6b63fa5e5ffc6260fe6862f5515b%
  fc6260fa5e5ffd6561ff7367ff7367fe6862ff846eff9f7dff846eff7367ff7367ff846eff6b63fe6862fe6862fd6561%
  ff7367ff6b63fe6862f14958f4505aee4256f4505af04757f14958ef4356f04557f14958f4505afa5e5ff95c5ef95c5e%
  ff7367fe6862ff846efd6561fe6862ff7e6cfc6260ff846eff7e6cfd6561fd6561ff7869fe6862ff6f65f95c5eff6b63%
  ff7869ff8c72ff8c72ff7367ff7869ff9576ff8c72ff846eff846eff9576ff846eff9576ff846eff8c72ff9f7dff7e6c%
  ff9f7dffac88ff9576ff9f7dff9576ff9f7dffac88ffac88ffac88ff7869ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffac88fe6862f5535bf7585df7565cf34e5af14858f14858%
  f7565cf7565cf5515bf7565cfa5e5ff34d59f34d59f34e5afc6260f4505af5535bfd6561ff7367fe6862fc6260ff6f65%
  fc6260fa5e5ffc6260f7565cff6b63f4505af7565cf5535bf34d59f24a59ee4256f04557ef4356ed4155ef4457f04557%
  ea3d54f04757ee4156f5515bf5535bf6545cf7565cff846eff846eff6f65f95c5efc6260f85a5df6545cfd6561fc6260%
  fd6561fc6260fe6862fb605fff6b63ff7367ff846eff7367ff7869ff9576ff846eff9f7dff9f7dff846eff9576ff7869%
  ffac88ff7869ff9576ff9576ff8c72ff9576ff8c72ff9f7dffac88ff9576ff8c72ffac88ffffffff9f7df7585dffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff6b63fc6260f6545c%
  f5535bf14858f04557f7565cfa5e5ff5515bf6545cf95c5ef7565ced4155de2e4ed62449df2f4eef4457f14858f95c5e%
  f5515bf85a5dff7367f5535bfa5e5fff6b63f85a5df5535bf85a5dfb605ff5535bf14858ef4356e83a53eb3e54eb3e54%
  f14958e83b53e73952ee4156ec3f55ec3f55ee4156ed4155f85a5df04757f14858f14858f34d59f4505af5535bfe6862%
  f4505af5515bfa5e5ffd6561fc6260f6545cff6f65ff6b63fc6260fd6561ff6b63fc6260ff6f65ff7e6cff6b63ff9f7d%
  ff9f7dff7869ff7367ff9576ff7869ff7e6cff8c72ff846eff7869ff9f7dff9576ff846eff8c72ff9576ff9576ffac88%
  ffffffff9f7dffffffe63852ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffac88fa5e5ffc6260f95c5ef95c5eff6b63fe6862ff6f65fc6260ff7e6cfd6561ff7e6cfe6862f5515bd9284bbe1037%
  b80c32d21f47db2a4ce33550f24b59f6545cf7585df24a59f4505af6545cfe6862fa5e5ff6545cf6545cec4055ed4155%
  e83a53df304ee33450e73a52ef4356e83a53e63852e73952e73a52eb3e54f4505aed4155ef4457fb605ff4505aef4457%
  ee4256f5515bf7585df14958f4505afb605ff85a5dff6f65fe6862fe6862fa5e5ffe6862ff846eff7367ff7e6cfb605f%
  fe6862ff7e6cff6b63ff846eff7367ff7869ff9576ff8c72ff9f7dff8c72ff9576ff7e6cff9576ff8c72ff7e6cff8c72%
  ff9f7dffac88ff7e6cff9576ff9f7dffac88ff7e6cff9f7dec4055ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffff7e6cf04757f7585df5535bf34d59f34e5aff7367fb605ffd6561fd6561ff6f65%
  fd6561fd6561ff6b63e43651d7254ad7254ae83b53f34e5af24b59f5515bef4457f34e5aee4256f7565cf5515bf85a5d%
  f04757ef4356e73a52ea3d54e2324fd8264adb2b4ce23350e63852e43651ee4156e83a53e43651ea3d54e73a52e43651%
  f04757f04757ef4356f7565cf14858f14858f34d59f34e5af14858f34e5afa5e5ffa5e5ff5515bfa5e5ff6545cfa5e5f%
  ff7367ff7367ff6f65ff7e6cfe6862ff7e6cff846eff6b63ff7367ff7869ff8c72ff7869ff7367ff7e6cff7869ff8c72%
  ff6f65ff7869ff846eff9576ff846eff8c72ff846effffffff9576ffac88ff9f7dffac88ff9f7df24b59ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff7869ee4156e93c53f04557f14958e73a52%
  ec4055ec3f55f24b59fb605ff5515bf5515bf34e5af34e5aec4055f24a59f7585df24b59fc6260f5515bf95c5ef6545c%
  f6545cf6545cf14958ed4155e93c53ed4155e43651e33450e2324fdb2a4cde2e4de63852e33550e73a52ee4156e53751%
  ea3d54ec3f55ed4155ec4055e93c53ed4155e93c53f04557ee4256f14858f34d59f5515bef4457f24a59ef4356f85a5d%
  fa5e5ff95c5efc6260f7585dfe6862fa5e5ff85a5df85a5dff6f65ff6b63fe6862ff7e6cff6b63ff6b63ff6b63ff7869%
  ff7e6cfd6561ff8c72ff7367ff7367ff7367ff7e6cff8c72ff8c72ff7e6cff9f7dff9f7dff8c72ff9576ffac88ff9f7d%
  ff9576ffffffee4156ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff6b63%
  ec4055e73a52ef4356e83a53f34e5af14858f24b59fb605ff6545cf7585dea3d54d21f47ca1941c7173ed62449e63852%
  f34d59f85a5def4457f7565cf5515bf14858f4505af24a59f14858eb3e54e33550e73a52e53651e0314fdf2f4ee53751%
  dc2c4de43551e63852e53751ea3d54f24a59e83a53ee4256f04757ea3d54e63852ef4457ec4055ef4356f14958f14858%
  eb3e54f7565cf24a59f04557f4505af24b59fa5e5ff6545cf85a5df6545cf5535bf95c5ef7565cf5535bfc6260f7585d%
  ff8c72ff7e6cff9576ff6f65ff7e6cff7869ff9576ff7e6cff7367ff9576ff7869ff8c72ff7869ff846eff8c72ff846e%
  ff846eff9576ff9f7dff9f7dff9576ff9576ff9f7de53651ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ee4156f85a5df4505afd6561ff7869fa5e5fff846efb605fff846eff7e6cff9576fb605fda2a4c%
  c7173ebe1037c2133bd42248e53751eb3e54f04757f85a5df34e5af34e5aef4356f04557ee4256eb3e54e1314fe23450%
  e63852ea3d54e63852e43651e63852e1324feb3e54e23350ee4156f24a59ef4356f04757f14958f14858ed4155ee4256%
  f4505af04557f04557ee4156f5535bf6545cf4505af34d59f6545cfb605ffe6862f6545cfc6260fc6260fb605ff7565c%
  ff6b63fa5e5fff6f65ff6b63ff6f65ff7e6cff6b63ff9f7dff6f65ff8c72ff9576ff9576ffffffffac88ffac88ff8c72%
  ff9576ff7e6cffac88ff9f7dff9f7dff9f7dff846eff8c72ffac88ff8c72ff9f7dff9576e0314fffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff846ed21f47d11e46e2324fe53651ec4055ee4156f34d59f5515bff7367%
  f95c5efe6862fb605ff4505ae93c53e63852d8264ac6163ecf1d45d42148d9284be33450ea3d54e73952e23450d9284b%
  df2f4ee63852df2f4edd2d4ddc2c4cdf2f4ee43651e33550ea3d54e63852ea3d54eb3e54e93c53f4505aef4356ef4356%
  ef4356f14858f04757f5515bf5535bf24b59f6545cf24a59fc6260f7585df5535bf5515bfa5e5ff7565cff6b63ff7869%
  ff7367ff7e6cff6b63ff8c72ff7869ff6f65ff7869ff7367ff9f7dffac88ffac88ff7e6cffac88ff9f7dff9576ff8c72%
  ffffffff9f7dffffffffffffffac88ffac88ff9576ff8c72ffac88ffac88ffac88ff9f7dffac88ff9576ffac88ff9f7d%
  ffac88f7585dffffffffffffffffffffffffffffffffffffffffffffffffff7367e33450bd1036d11e46d62449dd2d4d%
  e23350e93c53ef4356f14858f24a59f7585df7585ded4155e0314fe23450de2d4dca1940c01239c91840d42148d7254a%
  e53651db2a4ce73952e33450dc2b4cda294be33450e2324fe2324fe63852df304ee0304ee43551ee4156e23450ed4155%
  f04757ee4256ef4457f24b59ea3d54e1324fed4155ee4256ef4356f04557fa5e5ff5535bf5515bf6545cff7869fc6260%
  fe6862ff6f65ff7367ff7367ff7869ff8c72ff7e6cff9576ff9576ffac88ff7e6cff9576ff9576ff9f7dff9576ff9f7d%
  ffac88ffac88ffac88ff8c72ffac88ffffffff8c72ffffffff9576ffffffffffffffac88ffac88ffffffffffffffac88%
  ffffffffffffffac88ffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffd52248%
  d42248df2f4ee83a53ee4256f5515bf14858fe6862fc6260f5535bee4256dd2d4de23350d52248ca1941c8183fd52349%
  da294bdb2b4ce0304ee33550e43551e73952e83b53e53751e83a53e43551e43551de2e4de23450e53651e53751ec4055%
  e93c53ea3d54ef4356ed4155f14958f14858f5535bef4356ef4356ee4256f85a5df85a5df7565cfa5e5ff7565cfd6561%
  ff6b63ff7367ff7869ff7869ff8c72ff9576ffac88ff9576ffffffff8c72ff9576ff9f7dffffffff9f7dff9f7dff9f7d%
  ff9f7dff8c72ffac88ff9f7dff8c72ffac88ffffffffac88ff9f7dffffffffac88ff9f7dffac88ff9f7dffac88ff9f7d%
  ffffffffffffffac88ffac88ffffffffac88ffffffffffffffffffffac88ffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffff7e6cc4153cce1c44d9274bdc2b4cea3d54f24b59f24b59f7565ce23450d8264ad42248d52349%
  d7264ad42148c7173ecb1a42d32048da294bea3d54de2e4de2324fdf304ee1324fe1324fdb2a4cdc2b4cdb2a4ce1314f%
  dd2d4de43551e73a52e93c53e33450f14958ef4457f24a59f34d59f4505af4505af7565cf5535bf6545cf85a5df85a5d%
  fe6862fd6561f85a5dff9576f7585dfd6561ff6f65ff6b63ff9576ff7367ff846eff8c72ff9576ff9f7dffac88ffac88%
  ffac88ffac88ffffffffffffff9f7dffffffffac88ffffffffac88ff9f7dffffffffac88ff9f7dffac88ffac88ffac88%
  ff9f7dffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffffb605fe23450e53651f04557f04557f6545cfa5e5ffa5e5f%
  fd6561d32148c5163dd7254aee4156f04757ee4256e93c53e73952ec3f55e1324fe2324fe73a52e43551dc2c4de23450%
  d9274bdb2a4ce0304edd2d4ddc2c4cdf2f4eeb3e54f4505af34e5af04757f24b59f34d59fb605ffd6561ff6f65fa5e5f%
  ff6b63f95c5ef7565cfc6260f95c5eff6f65fd6561ff9576ff7367ff7367ff6f65ff7e6cff9576ff9f7dff9f7dff9f7d%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffec4055d42148e33450%
  ed4155f34d59f7585df95c5ef7585df4505ae63852ee4156f24b59f24a59f14858ec3f55ef4356ec4055ee4156e63852%
  e83a53ec4055df2f4ee73952e43651e53651e73a52e83b53e33450ef4457f5535bf5515bf7565cf6545cfa5e5ff95c5e%
  ff6f65fa5e5ff5535bf4505afb605ff7565cfc6260fe6862ff7367fd6561fc6260ff6f65ff7e6cff7367ff8c72ff7e6c%
  ffac88ff8c72ff9f7dffac88ff9f7dff9f7dffac88ff8c72ffac88ffffffffffffffffffff9f7dffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ff846eff7367e93c53e73a52ef4457fd6561f95c5efd6561fb605fff7869e93c53da294be33550ef4457e1314fde2e4d%
  e63852f04757f04757f24a59ef4457e43551ea3d54e63852ec3f55e63852ee4156ea3d54eb3e54f04557ef4356f34e5a%
  f24b59ec3f55f5535bf14958f5535bf7565cf7585df85a5df7565cf7565cf85a5dff6f65f85a5dff6b63ff6b63ff6b63%
  ff846eff6b63fb605fff7869ff9f7dff9576ff846effac88ff8c72ffac88ff9f7dffac88ff9576ffac88ffac88ffac88%
  ffac88ff9f7dffffffffac88ffac88ffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffff7585df04557fd6561fd6561ff846eff8c72ff846eff7e6cff9576fa5e5fd11f46%
  d11e46e2324fe83b53dd2d4dd62449e93c53f5535bf7585dec3f55f24a59e83b53e73a52ef4356ee4156e73a52f14958%
  f04757f95c5ef14958f5535bf85a5dfd6561fb605ffe6862fb605ffb605fff7e6cfc6260ff7e6cff7869ff9576ff9576%
  ff9f7dff8c72ffac88ffac88ff8c72ff9f7dffac88ffac88ffac88ffffffffffffffac88ff9f7dffac88ffac88ff9f7d%
  ffffffffac88ffffffffac88ffac88ffffffffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffff7e6ce0314fe33450ec3f55ff6b63ff7e6cff846e%
  ff7869fb605fff7367ef4457d9274be73a52ed4155e0314fdf2f4ee63852f04757f24b59f4505af14858ef4356ea3d54%
  ea3d54f14858f34d59ec4055f14858f6545cf7585df95c5ef85a5dfd6561ff9576ff8c72ff846effac88ff846eff9f7d%
  ffac88ffffffffffffff9f7dffffffffffffffac88ffffffffffffffac88ffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffdb2a4cc01239%
  cb1a42d9274be73952f14958f14958ec4055e0304ee1324fdc2c4dd52248d8264ad32148cd1b43d01d45d42148d52349%
  dc2c4ddd2d4dd62449d52349d42248dc2c4cdb2b4cdd2d4ddd2d4dea3d54e83a53f6545cf6545cfc6260ff7367ff8c72%
  ff9f7dff7e6cffac88ffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffff9f7ded4155d7254ae23350e83b53ee4156f04757f6545cd52248c3143bd11e46d52248d7254aca1941%
  d11e46d01e45d62449da294bda294bdc2c4dda294bdd2d4de2324fe63852e83b53eb3e54f04557f34e5af7585dfb605f%
  ff7367fe6862ff7367ff846eff9f7dffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffff846ef95c5ef7565cf7585dff7367ff7e6cff7367f5535bdb2a4c%
  e1324fec4055ef4457e83b53e93c53ea3d54ed4155ef4356f24b59f24b59f04757f14858e53751f7565cf7585dfa5e5f%
  ff6f65fa5e5fff7869ff9576ff9576ff7367ffac88ffac88ffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff7565cda2a4cd21f47de2e4de23450%
  e93c53e53751e33550d72549ce1c44d01e46d9274bdf2f4ed42248da294bde2d4dd9284be53751e43551df304edf2f4e%
  db2a4cea3d54f7565cfb605ff85a5dff6f65ff9576ff8c72ffffffffffffffac88ff9576ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9576%
  e83b53d42148da294bdc2c4ce23450df2f4ede2e4dd42248d42248cd1b43d52349d62449d7254ad72549dc2c4ce73a52%
  e83a53e23450ec3f55e93c53ec3f55f5535bf5535bfd6561fc6260ff7e6cff846eff8c72ff8c72ff9f7dffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffa5e5fd52349d32047d9284bdc2b4cd8274ad7264ace1c44c6163dc5153dc9183fc8183f%
  ce1c44d11f46d21f47d42148da294bdc2c4dde2e4de2324fe53651f24a59f24a59fc6260fd6561ff6f65ff7367ff846e%
  ffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88e43651c3143bd32047de2d4dd9284bd52248c6163d%
  c01239bf1138c01138be1037c2133bc5163dc91840cc1b42cb1a42d32048cf1d45d62449e23450eb3e54f34e5af95c5e%
  fd6561ff7869ff7e6cffffffff9576ffac88ffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffd6561c91840bb0e35%
  d21f47d21f47d62449c91840c9183fc91840c6163dc3143bc3143bc7173ec7173ec91840c91840cc1a42cf1d45d22047%
  de2e4de53751ee4256f5515bf95c5eff9f7dffac88ffffffff9f7dffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffde2e4dbb0e35b2072dc8183fd01d45d52248cd1b43c4143cc3143b%
  bc0f36b90d33bf1138c4143cc3143bce1c44d32047d62449df2f4ee43651ec3f55f6545cff7e6cff7367ff9576ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffad0528c01239c01238c5163dc11239b80c33b3082eb40930b3082fb4092fba0d34bc0f36c1133acb1a41d72549%
  e33450ec4055f24a59f95c5eff6b63ff6b63ff7e6cffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff8e0104a4041d9b0313990211980210%
  96020d97020ea1041ba90524b2082ebd1037c6163dd42248e53651f7565cff7e6cff9f7dffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffdc2b4c%
  f04757ea3d54e0314fe0304ee73952dc2c4ce43551e23350ea3d54ef4457fb605fff6b63ff9f7dff9576ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffff8b0000f85a5dffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffff}%
\gprawimage{rgb}{1.302}{2.179}{101}{101}{3.624}{3.624}{\gprawrgbimagedata}{}
\end{scope}
\gpcolor{rgb color={0.000,0.000,0.000}}
\draw[gp path] (1.320,2.197)--(1.356,2.233)--(1.392,2.269)--(1.429,2.306)--(1.465,2.342)%
  --(1.501,2.378)--(1.537,2.414)--(1.574,2.451)--(1.610,2.487)--(1.646,2.523)--(1.682,2.559)%
  --(1.719,2.596)--(1.755,2.632)--(1.791,2.668)--(1.827,2.704)--(1.864,2.741)--(1.900,2.777)%
  --(1.936,2.813)--(1.972,2.849)--(2.009,2.886)--(2.045,2.922)--(2.081,2.958)--(2.117,2.994)%
  --(2.154,3.031)--(2.190,3.067)--(2.226,3.103)--(2.262,3.139)--(2.299,3.176)--(2.335,3.212)%
  --(2.371,3.248)--(2.407,3.284)--(2.444,3.321)--(2.480,3.357)--(2.516,3.393)--(2.552,3.429)%
  --(2.588,3.465)--(2.625,3.502)--(2.661,3.538)--(2.697,3.574)--(2.733,3.610)--(2.770,3.647)%
  --(2.806,3.683)--(2.842,3.719)--(2.878,3.755)--(2.915,3.792)--(2.951,3.828)--(2.987,3.864)%
  --(3.023,3.900)--(3.060,3.937)--(3.096,3.973)--(3.132,4.009)--(3.168,4.045)--(3.205,4.082)%
  --(3.241,4.118)--(3.277,4.154)--(3.313,4.190)--(3.350,4.227)--(3.386,4.263)--(3.422,4.299)%
  --(3.458,4.335)--(3.495,4.372)--(3.531,4.408)--(3.567,4.444)--(3.603,4.480)--(3.640,4.517)%
  --(3.676,4.553)--(3.712,4.589)--(3.748,4.625)--(3.784,4.661)--(3.821,4.698)--(3.857,4.734)%
  --(3.893,4.770)--(3.929,4.806)--(3.966,4.843)--(4.002,4.879)--(4.038,4.915)--(4.074,4.951)%
  --(4.111,4.988)--(4.147,5.024)--(4.183,5.060)--(4.219,5.096)--(4.256,5.133)--(4.292,5.169)%
  --(4.328,5.205)--(4.364,5.241)--(4.401,5.278)--(4.437,5.314)--(4.473,5.350)--(4.509,5.386)%
  --(4.546,5.423)--(4.582,5.459)--(4.618,5.495)--(4.654,5.531)--(4.691,5.568)--(4.727,5.604)%
  --(4.763,5.640)--(4.799,5.676)--(4.836,5.713)--(4.872,5.749)--(4.908,5.785);
\gpcolor{color=gp lt color border}
\draw[gp path] (1.320,2.197)--(1.500,2.197);
\node[gp node right] at (1.136,2.197) {1};
\draw[gp path] (1.320,2.744)--(1.500,2.744);
\node[gp node right] at (1.136,2.744) {10};
\draw[gp path] (1.320,5.481)--(1.500,5.481);
\draw[gp path] (1.320,5.785)--(1.500,5.785);
\node[gp node right] at (1.136,5.785) {1h};
\draw[gp path] (1.320,2.197)--(1.500,2.197);
\draw[gp path] (1.320,2.362)--(1.410,2.362);
\draw[gp path] (1.320,2.580)--(1.410,2.580);
\draw[gp path] (1.320,2.691)--(1.410,2.691);
\draw[gp path] (1.320,2.744)--(1.500,2.744);
\draw[gp path] (1.320,2.909)--(1.410,2.909);
\draw[gp path] (1.320,3.127)--(1.410,3.127);
\draw[gp path] (1.320,3.238)--(1.410,3.238);
\draw[gp path] (1.320,3.292)--(1.500,3.292);
\node[gp node right] at (1.136,3.292) {$10^{2}$};
\draw[gp path] (1.320,3.456)--(1.410,3.456);
\draw[gp path] (1.320,3.674)--(1.410,3.674);
\draw[gp path] (1.320,3.786)--(1.410,3.786);
\draw[gp path] (1.320,3.839)--(1.500,3.839);
\node[gp node right] at (1.136,3.839) {$10^{3}$};
\draw[gp path] (1.320,4.004)--(1.410,4.004);
\draw[gp path] (1.320,4.221)--(1.410,4.221);
\draw[gp path] (1.320,4.333)--(1.410,4.333);
\draw[gp path] (1.320,4.386)--(1.500,4.386);
\node[gp node right] at (1.136,4.386) {$10^{4}$};
\draw[gp path] (1.320,4.551)--(1.410,4.551);
\draw[gp path] (1.320,4.769)--(1.410,4.769);
\draw[gp path] (1.320,4.880)--(1.410,4.880);
\draw[gp path] (1.320,4.933)--(1.500,4.933);
\node[gp node right] at (1.136,4.933) {$10^{5}$};
\draw[gp path] (1.320,5.098)--(1.410,5.098);
\draw[gp path] (1.320,5.316)--(1.410,5.316);
\draw[gp path] (1.320,5.428)--(1.410,5.428);
\draw[gp path] (1.320,5.481)--(1.500,5.481);
\draw[gp path] (1.320,5.645)--(1.410,5.645);
\draw[gp path] (1.320,2.197)--(1.320,2.377);
\node[gp node center] at (1.320,1.889) {1};
\draw[gp path] (1.867,2.197)--(1.867,2.377);
\node[gp node center] at (1.867,1.889) {10};
\draw[gp path] (4.604,2.197)--(4.604,2.377);
\draw[gp path] (4.908,2.197)--(4.908,2.377);
\node[gp node center] at (4.908,1.889) {1h};
\draw[gp path] (1.320,2.197)--(1.320,2.377);
\draw[gp path] (1.485,2.197)--(1.485,2.287);
\draw[gp path] (1.703,2.197)--(1.703,2.287);
\draw[gp path] (1.814,2.197)--(1.814,2.287);
\draw[gp path] (1.867,2.197)--(1.867,2.377);
\draw[gp path] (2.032,2.197)--(2.032,2.287);
\draw[gp path] (2.250,2.197)--(2.250,2.287);
\draw[gp path] (2.361,2.197)--(2.361,2.287);
\draw[gp path] (2.415,2.197)--(2.415,2.377);
\node[gp node center] at (2.415,1.889) {$10^{2}$};
\draw[gp path] (2.579,2.197)--(2.579,2.287);
\draw[gp path] (2.797,2.197)--(2.797,2.287);
\draw[gp path] (2.909,2.197)--(2.909,2.287);
\draw[gp path] (2.962,2.197)--(2.962,2.377);
\node[gp node center] at (2.962,1.889) {$10^{3}$};
\draw[gp path] (3.127,2.197)--(3.127,2.287);
\draw[gp path] (3.344,2.197)--(3.344,2.287);
\draw[gp path] (3.456,2.197)--(3.456,2.287);
\draw[gp path] (3.509,2.197)--(3.509,2.377);
\node[gp node center] at (3.509,1.889) {$10^{4}$};
\draw[gp path] (3.674,2.197)--(3.674,2.287);
\draw[gp path] (3.892,2.197)--(3.892,2.287);
\draw[gp path] (4.003,2.197)--(4.003,2.287);
\draw[gp path] (4.056,2.197)--(4.056,2.377);
\node[gp node center] at (4.056,1.889) {$10^{5}$};
\draw[gp path] (4.221,2.197)--(4.221,2.287);
\draw[gp path] (4.439,2.197)--(4.439,2.287);
\draw[gp path] (4.551,2.197)--(4.551,2.287);
\draw[gp path] (4.604,2.197)--(4.604,2.377);
\draw[gp path] (4.768,2.197)--(4.768,2.287);
\draw[gp path] (1.320,5.785)--(1.320,2.197)--(4.908,2.197);
\node[gp node center] at (3.114,5.426) {33\% Labelled};
%% coordinates of the plot area
\gpdefrectangularnode{gp plot 1}{\pgfpoint{1.320cm}{2.197cm}}{\pgfpoint{4.908cm}{5.785cm}}
\draw[gp path] (6.780,5.785)--(6.780,2.197)--(10.368,2.197);
\node[gp node center,rotate=-270] at (5.706,3.991) {Association Runtime (ms)};
\node[gp node center] at (8.574,1.427) {MAC Both Runtime (ms)};
\begin{scope}
\clip (6.780,5.785) rectangle (10.368,2.197);
\def\gprawrgbimagedata{%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffac88ff9576ff9f7dff9f7dffffffffac88ff9f7d%
  ffffffff9f7dffac88ffffffff8c72ff8c72ffac88ffac88ffac88ff9f7dff846eff846efd6561fa5e5ffe6862ff7e6c%
  fe6862f85a5df7585df4505aef4457f04557e73952df2f4ee2324fd9284bd62449d42148d21f47d21f47cf1d44cd1b43%
  cc1b43d11e46d21f47cc1b43c7173ed01e46ca1940cd1c43cf1d45ce1c44d52349d42248dc2c4cda2a4cde2d4dd22047%
  d42248d8264ad7254ad7264aa00319ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffffffffffff%
  ffac88ff9f7dffffffff9f7dffffffffac88ff9f7dff9f7dffac88ffac88ffac88ff8c72ff6f65ff9f7dffac88ffac88%
  ff9f7dff8c72ff9f7dff7e6cff9576ff9576ffffffff9f7dff9f7dffffffffac88ffac88ffac88ffffffff9f7dff9f7d%
  ff9f7dffac88ffac88ff9f7dffac88ffffffffffffffffffffac88e1314fffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffffffffac88%
  ffffffffffffff9576ffffffff9576ffffffffac88ffffffffffffff9576ffac88ffac88ff9576ff846eff9576ff9f7d%
  ffac88ff8c72ffffffffac88ff9576ff9576ff846eff846eff846eff7869ff7e6cff7869ff7e6cff846eff9576ff9f7d%
  ffac88ff846effffffff9f7dff9576ffac88ff9f7dffac88ffffffffffffffffffffffffffffffffffffe2324fffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffffffffac88ffac88%
  ffffffffffffffffffffffffffffffffffffffac88ff8c72ffffffff9f7dff9f7dffffffff9f7dffac88ffffffff9576%
  ff9576ffffffffac88ffffffff9f7dffac88ff9576ff7869ff9576ff9576ff846eff8c72ff6f65ff7367fd6561ff7e6c%
  fd6561ff8c72ff846eff7869ff9f7dff846eff9f7dff9576ffac88ffffffff9f7dffffffffac88ffffffffffffffffff%
  ffffffffffffffffffe83b53ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffff%
  ffac88ff9f7dffac88ffac88ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88%
  ffac88ffffffffac88ffffffff846eff9f7dffffffffac88ff9576ffac88ff9f7dff8c72ff9576ff7e6cff7367f95c5e%
  fb605ffa5e5ffb605fff7367ff7367fc6260ff7869ff7367ff7869ff7e6cff7367ff9576ff9f7dff9f7dffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffe73952ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff846eff9f7dff7e6cffffffff9f7dffffffffffffffac88ffffffffac88%
  ff9576ffffffffffffffac88ffac88ffac88ffac88ff9f7dffac88ff9f7dff9f7dffac88ff8c72ff7869ff8c72ff7e6c%
  ff7869fd6561f95c5ef95c5ef7565cf7585df5515bf14958fa5e5ff85a5dfb605fff6f65ff7367ff6b63f95c5efb605f%
  ff8c72ff8c72ffac88ffffffffac88ffffffff9f7dffffffffffffffffffffffffffac88fffffff04557ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffff9f7dffffffffac88ff9576ffac88ffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffac88ffffffffffffff9576ff9f7dff7e6cff8c72%
  ffac88ff7e6cff6f65ff6f65ff7869ff7367f6545cfa5e5ff7585df5535bf24b59f85a5df5535bfc6260ff7e6cfc6260%
  f85a5df95c5ef6545cff6f65ff6f65ff9f7dff9576ffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffb605fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffffff9f7d%
  ff9f7dff9576ffac88ff9f7dffffffffac88ffffffffffffffac88ff8c72ff9576ffac88ffffffff9576ffffffff8c72%
  ff9f7dff9576ff846effac88ff8c72ff846eff7367f7585dff6b63f7585df85a5df24a59fb605fff6b63f6545cf24b59%
  f5535bf85a5df85a5df5535bf7585dff6b63f7565cff7367ff7e6cff9f7dff9576ffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffff8c72ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffac88ff8c72ff9576ff9f7dffac88ffffffffffffffffffffffffffffffffac88ffac88ffac88%
  ff9f7dffac88ffac88ffffffff9576ff8c72ff8c72ff9f7dff7e6cff7e6cf95c5eff7869ff6b63f7565cf5515bfe6862%
  fc6260f5515bf7585df95c5ef7565cfa5e5ff85a5dfd6561f5515bf7585df7565cfd6561ff7e6cff8c72ff9f7dffffff%
  ffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffffffffffffff8c72ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffff9f7dffac88ff9576ff8c72ff9f7dff9576ff9f7dffac88ffac88ffffff%
  ff9f7dffac88ff9f7dffac88ffac88ff9f7dffffffff9f7dff9f7dffac88ff9f7dff9f7dff7367ff7367fd6561f7565c%
  fe6862f85a5df5515bf7565cf85a5df14958f5535bf85a5dfa5e5fff846eff6b63ff7e6cf24b59fa5e5ff85a5dfd6561%
  ff9576ff9576ff8c72ff9576ff9f7dffac88ffffffff9f7dffac88ffac88ffffffffffffffac88ffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ff9f7dff9576%
  ffac88ff9576ffac88ffffffffffffff9576ff9f7dffac88ff9f7dff9f7dff8c72ff8c72ff8c72ff9576ff8c72ff7869%
  fe6862ff6f65fc6260fa5e5ff7565cf6545cfb605ff5535bf14958ff7e6cfc6260fd6561ff6f65fc6260fe6862ff7869%
  fe6862f7585dfc6260ff6b63ff846eff9576ff9f7dffac88ffac88ffffffffac88ff9f7dffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffac88%
  ffac88ff8c72ffac88ffac88ffffffffffffffac88ffffffff9f7dffac88ff9576ffffffff9f7dff9f7dff9f7dffffff%
  ffffffff8c72ff7367ff7367ff8c72ff6f65f34e5af4505af34e5af85a5df6545cfb605ffa5e5ffc6260fb605fff846e%
  ff6f65ff7e6cff8c72ff7e6cff7367fb605ffc6260ff6f65ff6b63ff9576ff9576ffac88ffffffffac88ffac88ffffff%
  ffac88ffac88ffac88ffac88ff9f7dffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ffac88ffac88ff9f7dffffffffffffff9f7dffffffffac88ffffffffffffff8c72ff9f7d%
  ff9f7dff9576ff9576ff9f7dffac88ff846eff7e6cff8c72ff6f65fc6260f7585df7565cf85a5df4505af7565cf6545c%
  f95c5eff7869ff7869fe6862ff846eff7367ff7869ff7367ff7367ff7869ff7e6cff6f65ff7869ff9f7dff8c72ffffff%
  ffac88ffffffff9f7dffffffffffffffffffffffffffffffffac88ffac88ffffffffffffff9f7dffac88ffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffac88ffffffffffffffffffffffffffffff%
  ff9f7dffac88ffac88ff9f7dff9576ffac88ff9f7dff9576ff9f7dff9576ff9f7dff7367ff9576ff6f65f7565cf7585d%
  f85a5df6545cf7585df85a5dff7367fc6260fb605ffe6862ff6f65ff9576ff8c72fc6260ff7367ff7e6cff846eff846e%
  ff7e6cff9576ff9f7dff9f7dff9f7dff9f7dffac88ffac88ffac88ff8c72ffffffffac88ffffffffac88ffac88ffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffff%
  ffffffff9f7dffac88ffac88ff9576ffac88ffffffffffffff9576ff9576ff9f7dffac88ff846eff846eff9576ff9576%
  ff7e6cff6b63fd6561f7585df85a5dfb605ff04757f4505afd6561fc6260ff6f65ff7869ff8c72ff7869ff8c72ff6f65%
  ff8c72fe6862ff9f7dff9576ff9576ffffffff9f7dffffffff9576ffac88ff8c72ff9576ffac88ffac88ffac88ff846e%
  ff9f7dff9576ffffffffffffffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffff9f7dff9576ffffffffac88ff9f7dffac88ffac88ffffffff9f7dffac88ff9576ffac88ff9f7dff9576%
  ffac88ffffffff9f7dff7869ff9576ff7e6cff6f65f85a5dfc6260f34e5af95c5efe6862fc6260ff6f65ff7869ff846e%
  ff6f65ff6f65ff7e6cff7e6cff7869ff6b63ff7869ff7367ff9576ff8c72ffac88ffac88ffffffff846eff7e6cff7869%
  ff846effac88ffffffffac88ffffffffffffffffffffffffffac88ffac88ffffffff9f7dffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffac88ff9576ffac88ffac88ffffffff9f7dffac88ff9f7dffffffffac88ffac88ffac88%
  ffffffff846effac88ffac88ff9f7dff9576ff8c72ff8c72ff7367fd6561ff846eff6f65fa5e5ffe6862fb605ff7565c%
  fb605fff6f65ff846eff8c72ff6b63f95c5efc6260ff7e6cff7367ff7e6cff7e6cff846eff9576ffac88ff9576ff9f7d%
  ff9f7dff9f7dff9f7dff7367ff6f65ff846eff9576ff9576ffffffff9576ffffffffffffffffffffac88ffffffffac88%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9576%
  ff9f7dffac88ffac88ff9576ff9f7dffac88ffac88ff8c72ff9576ff9576ffac88ff846eff9f7dff846eff7e6cff7869%
  ff6b63fc6260f85a5dfb605ffd6561ff7e6cff7e6cff7869ff9f7dff6b63ff7869ff7e6cff9576ff8c72ff7367ff7367%
  ff9576ffac88ffffffff9f7dffac88ffac88ff846eff7e6cff7869ff7869ff9f7dff9576ff9f7dffac88ff9576ffac88%
  ff9f7dffac88ff846effffffffffffffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9576ffac88%
  ffffffffffffffac88ffffffffffffff9576ffac88ff9576ffac88ff8c72ff9576ff9576ff7e6cff7e6cff846eff7e6c%
  ff9f7dff9576ff6f65fd6561ff7869ff846ef95c5eff8c72ff7869ff9576ff7e6cff6f65ff7869ff9576ff7367ff9f7d%
  ff846eff9f7dff7e6cff8c72ff9576ff9576ffffffff9f7dff9576ffffffffac88ff9576ff7869ff846eff7869ffffff%
  ff7e6cffffffff9f7dff9576ff9f7dff7e6cf85a5df85a5dff9f7dffffffffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ff9f7dffac88ff9576ffac88ffffffffffffffac88ffffffffac88ffac88ffac88ffac88ff7e6cff846eff9576ffffff%
  ffac88ff9f7dff8c72ff9f7dff846eff7869ff8c72ff7367fd6561ff9576ff7e6cff6b63ff7367ff6b63ff8c72ff7869%
  ff7e6cff8c72ff9576ff846eff7869ff7869ff9576ff9576ff9576ff9576ffac88ffac88ffffffff9f7dffffffff9f7d%
  ff9f7dff6f65ff7e6cff9f7dffffffffffffffac88ff9f7dff7e6cff9576fd6561f6545cff6b63ffac88ffac88ffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ff9f7dffffffffffffffffffffac88ffffffff9f7dff9f7dffffffffffffffac88ffffffffffffff9576%
  ffac88ff9f7dff7e6cff9f7dff9f7dff9576ff7e6cff9f7dffac88ff8c72ff8c72ff9f7dff7367ff7e6cff6b63ff846e%
  ff7367fd6561ff6f65ff6f65ff7e6cff9576ff8c72ff8c72ff846eff9576ff9f7dffac88ff9f7dffac88ffffffffffff%
  ff9576ff846eff7e6cff9576ffac88ff8c72ffffffffac88ff9f7dffac88ffac88ff7e6cfc6260fd6561ff7367f95c5e%
  f6545cff9576ffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffffff9576ffac88ff9576ffffff%
  ff8c72ff9f7dffac88ffffffff9f7dff9576ff9576ffac88ff846eff7367ff846eff7869ff7e6cff9576ff9576ff7e6c%
  ff8c72ff6b63ff7e6cff6f65ff7869ff7e6cff7367ff7367ff8c72ff8c72ff9576ff8c72ff9f7dffac88ff8c72ffffff%
  ffffffff9f7dff9f7dff9f7dff9f7dffffffff8c72ff8c72ff846eff9f7dff9f7dffac88ffac88ff8c72ffffffff9576%
  ff6f65fc6260fd6561ff6f65fd6561ff8c72ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffac88ffffffff9f7dffffffffac88%
  ffffffffffffff9f7dffffffffac88ffffffff9f7dff9576ff8c72ff9f7dff7e6cff9576ff846eff9f7dff7869ff8c72%
  ff9576ff846effac88ff9f7dff9f7dff9f7dff846eff846eff9576ff846eff9576ff9576ff9576ff7e6cff9f7dff9f7d%
  ff8c72ff8c72ffac88ffac88ff9f7dffffffffac88ff9576ffffffff9576ff9f7dff7e6cff846eff846effac88ff9f7d%
  ffffffffac88ffffffffac88ff9576ff6f65fd6561ff6b63ff7869ff7e6cffffffffffffffffffffffffffffffffac88%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ff9576ffffffff9f7dff9576ff9f7dffffffffffffff9576ffac88ffac88ffac88ff9f7d%
  ff9576ff846eff846eff6f65ff7e6cff6b63ffac88ff7e6cff8c72ff846effffffff7869ff8c72ff9f7dff8c72ff9f7d%
  ffac88ff7367ffac88ff9576ff846effac88ff9576ffac88ffac88ff9f7dffffffffffffffffffffac88ffffffffac88%
  ff846eff9f7dffffffff9f7dff9f7dffac88ffac88ffffffffac88ff8c72ff9576ff846eff7e6cffac88ffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffac88ffac88ffac88ffffffffac88ffac88ff9f7dff9f7dff8c72ffffffff9576ff9576%
  ffac88ff9576ffffffffac88ff9576ff9f7dff9f7dff6b63ff8c72ff7869ff9576ff846eff7869ff9f7dff9f7dff846e%
  ff9576ff846eff8c72ff846eff9576ff9576ffac88ff8c72ffac88ff9576ff9f7dffac88ffac88ffffffffac88ffac88%
  ffffffffffffffac88ffac88ff9f7dffac88ff7e6cff9576ffac88ffac88ffac88ff9f7dffffffffffffffffffff9f7d%
  ff9f7dffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffac88%
  ffac88ff9f7dff9f7dff9576ff9576ff9576ff9f7dffffffff9f7dff8c72ff8c72ff9f7dff9576ff8c72ff7367ffac88%
  ff9576ff9f7dff9f7dffffffff8c72ff7e6cff8c72ff9576ff8c72ff9576ff9576ff9576ffac88ffac88ffffffffffff%
  ff8c72ff9f7dff9576ffffffffffffff9f7dffffffff7e6cff9f7dff7e6cff9f7dffffffffac88ffffffff9f7dff9576%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffac88ffac88ffffffffffffff9f7dffffffff9f7dffffffffffffffac88ffffffff9576ffac88ffac88ff9f7d%
  ff9576ff9576ff9f7dff9f7dffac88ffac88ff9f7dffffffff8c72ffac88ff9f7dff7367ff9576ff8c72ff9576ff846e%
  ff7e6cff7e6cff9576ff9576ff9576ff8c72ffac88ffffffffffffffac88ff8c72ff9f7dffac88ff9576ff9576ffac88%
  ff9f7dffac88ffffffff9f7dffffffffffffffac88ff9576ffffffffac88ffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffac88ffffffffffffffffffff9576ff8c72ff9576ffac88ff9f7dff8c72%
  ff9576ff8c72ff8c72ff9576ff9f7dffac88ffac88ff9576ff7e6cffffffff8c72ff9576ff9f7dffac88ff846eff9576%
  ff7e6cff9576ffac88ff7869ff8c72ff7367ff7e6cff7869ff6f65ff9576ff8c72ffffffffffffffac88ff9576ffac88%
  ffac88ffac88ff9576ffac88ffffffff9f7dff9576ffffffffac88ffffffffac88ffffffffac88ffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffac88ff9f7dffac88ffac88ff9f7d%
  ff8c72ff7e6cff9576ffac88ff9f7dff9576ff9576ff9576ff9f7dff9f7dff9576ffffffffac88ff9f7dffac88ff9f7d%
  ffac88ff8c72ff9f7dffffffff9f7dff8c72ff7e6cff6b63ff846efc6260fc6260fb605ff95c5efd6561ff8c72ff9f7d%
  ff9f7dff8c72ff9f7dffffffffffffff9576ff9f7dffffffff9f7dffac88ffac88ff9576ff8c72ffffffff9f7dff9f7d%
  ffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffff9576ff9576ff8c72ff7e6cffac88ff8c72ff9576ff846eff7e6cff9576ff846eff9f7dff9f7d%
  ffffffff9f7dffffffff9f7dff8c72ff9f7dff9576ff8c72ff8c72ff846efb605ff95c5ef7585df85a5df5515bf85a5d%
  fb605ff6545cfe6862ff846eff9f7dffac88ff9f7dffffffff9f7dff9576ff9f7dff9f7dff8c72ffffffffac88ffac88%
  ff9f7dff9576ffffffffac88ffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffac88ffffffffac88ff9f7dff7367ff7367ff8c72ff846effffffffac88ff9576%
  ff7e6cff9576ff9f7dffffffffffffff7869ffac88ff9f7dff9f7dff846eff7367ff846eff7e6cfd6561f7585df95c5e%
  fa5e5ff5535bf04557f34d59f14958f24b59f5515bff6b63ff7e6cff9f7dffffffffffffff9f7dffffffff9576ff8c72%
  ffac88ff846eff9f7dffffffffac88ff9576ffac88ffffffffffffffac88ffffffffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffff9f7dffac88ff9f7dff7869%
  ff9f7dffac88ffac88ff846eff9576ff9f7dffac88ff8c72ff9576ffac88ff8c72ff7e6cff846eff7367ff7e6cff7367%
  ff7367ff6b63fa5e5ff7565cf95c5efa5e5ff34e5af4505af95c5ef24a59f5515bf4505afc6260ff8c72ffac88ff9576%
  ff9576ffac88ffffffffac88ff9f7dff846eff9576ff9576ffffffffac88ff9f7dffac88ffffffffffffffffffffac88%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffac88ff9576ffac88ffac88ffffffffac88ff9f7dff8c72ff9576ff9f7dff9f7dff9576ff8c72ff7e6c%
  ff846eff6f65ff6b63ff7367ff6b63fe6862fc6260f85a5df85a5df6545cf4505af4505af7585df5535bf4505af4505a%
  fc6260ff7e6cffac88ff9576ff9f7dffffffffffffff9f7dffac88ff8c72ff8c72ff9f7dff9f7dff9f7dff9f7dff9f7d%
  ff9f7dffffffffffffffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffff9f7dffffffffac88ffac88ffffffffffffff9f7dffffffffffffffffffff9f7dff9576ff8c72%
  ff9f7dff9f7dff9576ff846eff8c72ff6b63f85a5df95c5ef7585dff6b63fa5e5ffd6561fb605ff6545cf6545cf34e5a%
  f7565cf4505af5515bfb605ff4505afb605fff7869ff8c72ff9f7dffac88ffffffffac88ffac88ffac88ff9f7dff9f7d%
  ff9f7dffffffff9f7dffffffffffffff9576ffffffffffffffac88ffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffac88ffffffffac88ffffff%
  ff9f7dffac88ffac88ffac88ff9f7dff9576ff846eff6f65fe6862ff7367fb605ff7585df95c5efc6260f6545cfe6862%
  f6545cf95c5efc6260f95c5efb605ff5515bf7585dff6f65ff7367ff8c72ff7e6cff9f7dff9576ff9f7dff9f7dff8c72%
  ff9f7dff9f7dffffffff9f7dffffffffac88ff9f7dffac88ffffffff9f7dffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ff9f7d%
  ffffffffac88ff9f7dff9f7dffffffffac88ffffffffac88ff846eff9f7dff8c72ff7e6cfe6862f95c5ef7565cf14958%
  f4505afd6561fa5e5ff7585dfe6862ff7367ff6f65f7585dfb605fff6b63f7565cff6b63fa5e5ffa5e5ffc6260ff846e%
  ffac88ffffffffffffff8c72ff9576ffac88ffffffffac88ffffffffffffff9576ffffffffffffffffffff9f7dff9f7d%
  ff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffac88ffffffff9f7dff9f7dffac88ff9f7dff9576ff846eff9576ff846e%
  f85a5dff6b63f7565cf95c5ef4505af6545cf5535bff6b63fe6862ff6b63ff7e6cfe6862ff6b63ff6b63fe6862fb605f%
  f95c5eff6b63fe6862ff7e6cff846effffffffffffff9f7dffffffff7e6cffffffffffffff9576ff846effac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ffac88ff9f7dffac88ffffffffac88ff8c72ff9f7d%
  ff9f7dff9f7dff9f7dff7e6cfe6862fd6561f7565cf85a5df34d59f14958f95c5efa5e5ffe6862fd6561ff7367ff7e6c%
  ff7e6cff7e6cfb605ffd6561f6545cf85a5dff6f65ff6f65ff7e6cff9f7dff9576ffac88ff9f7dffac88ff8c72ffac88%
  ff7e6cff6b63ff6f65ff9f7dffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffac88ffac88ffffffff9f7dff9f7d%
  ffac88ffffffff9576ff9576ff9576ff9f7dff7869ff7e6cff7e6cfb605ff85a5dfa5e5ff7565cf5515bf95c5efc6260%
  ff6b63fe6862ff7e6cff7e6cff846eff9576fd6561ff6b63f95c5eff6b63ff6f65ffac88ff846eff8c72ff846eff7869%
  ff9576ff9576ff8c72ff9f7dff7367f7585ded4155f95c5eff9576ff9f7dff9f7dffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffff%
  ffffffffffffffac88ffac88ffac88ffffffff9576ffac88ff9576ffac88ff8c72ffac88ff846eff6f65ff7e6cfe6862%
  fe6862fe6862ff7367fb605fff7869ff846eff9576ff846eff9576ff846eff7869ff6b63ff7869ff7367ff8c72ff9576%
  ff8c72ff6f65ff7869ff7e6cfd6561ff7e6cff846eff846eff8c72fa5e5fed4155eb3e54ff7367ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ffffffff9576ffac88ff9f7dffac88ffac88ff9f7dff9f7dffac88ff9f7dff9f7dff7869%
  ff7e6cff7e6cff7367ff7869fc6260fb605fff7367ff9576ff7869ffac88ff9576ff8c72ff8c72ff8c72ff846eff8c72%
  ff6f65ff7869ff9576ff846effac88ff9576ff846eff6b63fa5e5fff7869ff8c72ff9f7dff7e6cff7e6cf7585df7585d%
  ff7869ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffac88ff9f7dffffffffac88ff9576ffac88ffffffffac88%
  ff9f7dffac88ff9576ff6f65ff7367ff7367fe6862ff846eff8c72ff7869ff6f65ff7869ff9f7dff8c72ff846effac88%
  ff8c72ff8c72ffac88ffffffff8c72ff9f7dff9576ff9f7dff7e6cff9576ff9f7dff7e6cfe6862fd6561ff7869ff6b63%
  ff9576ff9576ff846eff8c72ff7e6cff9576ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffac88ffac88ffffffffffffffac88ff9576%
  ffffffff9576ffffffffffffff8c72ffffffff9576ff9576ff7e6cff6f65ff6b63ff6f65ff9576ff7e6cff7e6cff9576%
  ff846eff7e6cff7869ff6f65ff846eff7e6cff8c72ff9f7dffac88ff9f7dff846eff7869ff9576ff8c72ff846eff9f7d%
  ff7367ff6f65ff8c72ff846effac88ff9576ff9f7dff9576ff9f7dffac88ffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ff9f7dffffffffffffffac88ffffffffffffffffffff9f7dffac88ff9f7dffac88ff9576ff9576ff9576ff7869ff7367%
  ff9576ff7367ff9576ff9f7dff846eff8c72ff9f7dff6f65ff6b63ff7367ff8c72ff9f7dff846eff9576ffac88ffac88%
  ff9576ffffffff9f7dff8c72ff9f7dff9f7dff7e6cff9f7dff846eff9576ff9f7dffac88ff9f7dffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffac88ffffffff9f7dff846effac88ffac88ffffffff9f7dff9f7dffac88ff9f7dffffff%
  ffffffffac88ff7869ffac88ff8c72ff8c72ff7869ffac88ff8c72ff9f7dff846eff7367ff9f7dff846eff8c72ffac88%
  ff9576ff9576ff846eff9f7dffac88ff9576ff9576ff9576ff7869ff7869ffac88ffac88ff9576ffac88ff9f7dff9576%
  ff9f7dff9f7dffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffac88ff9576ff6b63%
  ffac88ff7e6cffffffff9f7dff7869ffac88ff9f7dff9576ff9576ff9576ffac88ff8c72ff7e6cff9f7dff8c72ff846e%
  ff9576ff9576ff9576ffac88ffffffff9f7dff9f7dff9f7dffffffff9f7dff9576ff7e6cff846eff846eff7e6cffffff%
  ff9576ffac88ff9f7dff9576ff9f7dffffffffffffffac88ffffffffac88ffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffff9f7dffffffff9f7d%
  ffffffff9f7dff8c72ff7e6cff9576fe6862ff9576ffffffff9f7dff9576ff9f7dff9576ffac88ffac88ff846effac88%
  ff9576ffac88ffac88ffac88ff8c72ff8c72ff9f7dff9576ff9f7dffffffff9f7dff9576ffffffff8c72ff9576ffac88%
  ff9f7dff7869ff7367ff9576ff7e6cff9f7dff9576ff9f7dffac88ffac88ffffffffffffffac88ffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ff9f7dffffffffac88ff9576ff9576ff8c72fc6260ff7367ff6b63ffac88ffac88ffac88ffac88ff9f7d%
  ff9f7dffac88ffffffff846effffffff9f7dff7e6cff9576ff9576ffac88ff9f7dffffffffac88ff9f7dffffffff9f7d%
  ffffffff9f7dffffffff846eff9f7dff8c72ff7869ffac88ff7367ff9576ff9576ff9576ff9576ffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffff9f7dffffffff9f7dffffffff9576ffac88ff9f7dffac88ff846eff7e6cff7869%
  ffac88ff9f7dffac88ffac88ff9f7dffffffff9f7dffac88ffac88ffac88ff846effac88ffffffff9f7dffac88ffffff%
  ffffffffac88ff9f7dffac88ffac88ff9f7dffac88ff8c72ffffffffffffff846eff9f7dff7367ffffffff846effffff%
  ffac88ff9f7dff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffff8c72ff9f7dffac88ff846effac88%
  ff9576ff9576ff846eff6f65ff9f7dff8c72ffac88ff9576ff9f7dffffffffffffff9f7dffac88ffffffffffffff9576%
  ff846eff9f7dffac88ffac88ff9f7dff9f7dff9f7dff9f7dffac88ffac88ffac88ffac88ff9576ffac88ff9576ff9f7d%
  ff846effffffff9f7dff9576ff9576ffac88ffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7d%
  ff9f7dffac88ffffffff9576ff9576ff9f7dff9576ff9f7dff846eff9f7dffffffff9f7dffac88ffffffff9576ffffff%
  ffac88ff9f7dff9f7dff9576ffac88ffffffffffffff9576ffffffffac88ff9576ffffffff8c72ffac88ffffffffffff%
  ffffffff9576ff9576ffffffffffffffac88ff9f7dffac88ff9f7dffac88ffac88ffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffff9f7dffffffffffffffffffffffffff8c72ff846effac88ff8c72ff9f7dff9f7dff9576ffffff%
  ffac88ffac88ffac88ff9576ff9f7dff9576ffac88ff9f7dffac88ff9f7dff9f7dff8c72ffac88ff9576ffac88ff9576%
  ff8c72ff846eff9f7dff7367ff9f7dffffffff9f7dff7e6cffac88ffffffffac88ff9f7dffffffff9f7dffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffffffff9f7dff9576%
  ffac88ff9576ff9f7dffac88ff9f7dffffffff9f7dffac88ffac88ff8c72ffac88ff9576ff9f7dff9576ff7869ff8c72%
  ff9f7dff846eff9576ff9576ff846eff8c72ff8c72ff8c72ff9f7dff846eff8c72ff9576ffffffff846effffffffffff%
  ffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffac88ffffffffac88%
  ffac88ff9f7dff9f7dffac88ffac88ffac88ffffffffac88ff9576ff9576ff9f7dff9f7dff9f7dff9576ff9576ff846e%
  ff7e6cff7e6cff7e6cff7e6cff7e6cff7367ff7367ff9576ff9576ffac88ff846eff7869ff7869ff6f65ff9576ff7e6c%
  ff9576ff9576ffffffffac88ffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff9576ffffffffac88ffffffffac88ffffffff8c72ffac88ff8c72ff9576ff9576ff9f7dff8c72%
  ff846eff7e6cff7e6cff6f65fd6561ff6b63fb605ffd6561fa5e5fff7367fd6561fc6260ff8c72ff6b63ffac88ff846e%
  ff7869ff7869ff7367ff9576ff9576ff9f7dffffffff9576ffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffff9f7dffffffffffffffffffffffffffffffffffffff9f7d%
  ff8c72ff8c72ff9576ff9f7dff846eff7367ff7869ff7e6cfb605ff7585df24b59f24a59f34d59f34e5afb605ffc6260%
  fd6561ff8c72ff7e6cff7e6cffac88ffac88ff7e6cff7367ffac88ff9f7dffac88ffac88ffac88ffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffac88ff9576ffac88ff9576ff9576ff9576ff6f65ff846efc6260fc6260fa5e5ffc6260f85a5df5535bef4356%
  f4505af34e5aef4457f85a5df4505aff6b63ff846eff7e6cff7e6cff9f7dff8c72ff9f7dffffffff9576ff9f7dffffff%
  ff9f7dffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffac88ffffffff6b63ff7869ff8c72ff7367fd6561f95c5ef34d59%
  f5515bf14958f04757f04557f6545cf04757f34e5af5515bff6b63fc6260fb605fff8c72ff7367ff9576ff9f7dffac88%
  ffac88ff9f7dff7869ff846eff8c72ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffac88ffac88ffac88ff9576ffffffff9576ff9f7dff8c72ff8c72ff8c72%
  ff7869f85a5df5515bf24b59f6545cf24b59f5515bee4256f14958f85a5df5535bf95c5eff7367fb605fff7869ff6f65%
  ff7367ff9576ff8c72ff8c72ff9576ff9f7dff7e6cff9f7dff846effac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ff9f7dffffffffffffffffffffffffffffffffffffffffffffffffff9576ffac88%
  ff8c72ff9f7dff7869ff7e6cfb605ff7585df7585df4505af5515bfa5e5ff24b59f95c5ef34e5af85a5df85a5dff6f65%
  fc6260ff7869ff7367ff8c72ff846eff6b63ff7e6cff6f65ff6f65ff7e6cff6f65ff6f65ff846eff9f7dffac88ffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffff9f7dff9576ff8c72ff846eff8c72ff6f65fc6260fc6260f4505af34e5aec3f55f14958f95c5ef7565cf85a5d%
  fc6260fa5e5ffe6862fd6561fe6862ff8c72ff6f65ff7e6cff6f65fb605fff7367fe6862fa5e5ff5535bf85a5dff7367%
  ff8c72ff7869ff9576ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88%
  ffffffffac88ffffffffac88ffffffff8c72ff9576ff9576ff9f7dff7367ff7869fd6561fc6260f24b59f34e5af14958%
  f24a59f85a5dfa5e5ffc6260ff6b63ff6f65ff6f65ff7869ff8c72ff7869ff846eff7e6cff846eff7e6cff7e6cfc6260%
  f85a5dfa5e5ffd6561fb605fff7e6cff8c72ffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffff9f7dffac88ffffffff9f7dff9f7dff8c72ff846eff6f65ff6b63fa5e5ffe6862%
  f5515bf34d59f5535bf85a5df7585df7585dfc6260f7585dfc6260fe6862ff7869ff8c72ff8c72ff7e6cff6f65fc6260%
  fc6260ff9576fe6862ff7367ff6b63ff6f65f85a5dfd6561fd6561ff7367ffac88ffac88ffac88ffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffac88ffffffff846effac88ff7869%
  ff7e6cfe6862ff9576f95c5efc6260f24b59f34d59f7565cf34e5afa5e5ffa5e5fff7869ff7869ff7e6cffac88ff846e%
  ff7367ffac88ff7367ff846eff9576ff7367ff6b63ff7e6cff9f7dff7869ff6f65ff7367ffac88ff8c72ff9f7dffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffffffff9f7dffac88%
  ff9576ffac88ff8c72ff8c72fd6561ff7869ff6b63ff6b63ff7869fa5e5ffb605ffc6260f85a5dfc6260ff6b63ff7367%
  ff7869ff7869ff6b63ff9576ff7367ff6b63f95c5eff6f65ff7869f95c5eff6f65ff846eff7869fd6561ff9576ff7869%
  ff9f7dffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffff9f7dff8c72ff9f7dff9576ff9576ff8c72ff846efe6862ff7e6cff7367fa5e5ffa5e5ffd6561fa5e5f%
  ff7367f85a5dff7367ff6b63fc6260ff7869ff7e6cff7367ff846eff6f65ff7869ff7e6cf95c5eff7e6cff7367ff846e%
  ff8c72ff7869ff8c72ff8c72ff9f7dffac88ffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffff9f7dffffffff9f7dff9576ffac88ff846eff7e6cff846eff7e6cff9f7dff8c72ff846e%
  fe6862ff7e6cfd6561ff7e6cfc6260fa5e5fff7869ff7869ff7367ff7e6cff8c72ff7367fe6862ff7367fe6862ff8c72%
  f95c5efb605fff6b63ff7869ff7869ff7869ff846eff9576ffac88ffffffff8c72ffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff846effac88ff8c72ff846eff846eff8c72%
  ff846eff8c72fd6561ff8c72ff846eff7367ff8c72ff6b63ff6f65ff7e6cff7e6cff8c72ff6b63ff9576ff8c72ff846e%
  ff6f65ff7367ff846eff7367ff7367f85a5dff6f65fd6561ff7367ff6f65ff9f7dff9576ffac88ffac88ffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffac88ff9f7d%
  ff846eff9f7dff8c72ff9576ff7869fd6561ff7367ff8c72ff8c72ff8c72ff9576ff846eff7e6cff7367ffac88ffac88%
  ff9576ff7367ff8c72ff846efd6561ff6b63ff6f65ff7367ff7e6cff8c72ff9576ff7869ff9576ff8c72ffac88ff9f7d%
  ffac88ff8c72ffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ff9f7dff9f7dffac88ff9576ff8c72ff8c72ff846eff6b63ff846eff9576ff846eff8c72%
  ff9f7dff8c72ff846eff9576ff7e6cff6f65ff9576ff7e6cff6f65fc6260f85a5dfe6862ff7869ff7869ff9576ff846e%
  ff7e6cff846eff9f7dff9f7dffac88ffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffff9576ff9f7dff7e6cff846eff7869ff7869ff7367%
  fb605fff6f65ff7367ff846eff9576ff8c72ffac88ffffffff8c72ff9576ff846eff8c72fc6260ff846eff6f65fe6862%
  ff7367ff9576ff7869ff9f7dff9f7dffffffff9f7dffac88ff9f7dffac88ffffffffffffffffffffac88ffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffff%
  ff9576ffac88ff9f7dff9f7dfe6862f7585df6545cff7e6cff8c72ff9f7dff9576ffffffff9576ff9576ffffffff9f7d%
  ff7e6cff9576ff7869ff846eff9576ff9f7dff7869ff9f7dffffffffac88ff9f7dffffffffac88ffac88ffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ff9f7dffac88ffffffffac88ffffffff8c72ff8c72ff8c72ff8c72ff9576ff846eff7869ff7e6cffffffffffffff9576%
  ffac88ffac88ffffffff8c72ff8c72ffac88ff9f7dffffffff8c72ff9576ffac88ff9f7dff9576ff9f7dffac88ffac88%
  ffac88ffac88ffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9f7dff8c72ff9f7dff9f7dff7367ff8c72ff846eff9576%
  ff9576ff9f7dff9f7dff9f7dff9f7dffffffffffffffffffff9576ffffffffac88ff9f7dff9f7dff7e6cffffffff9576%
  ff9576ffac88ffac88ffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ff9576ff8c72ff7869ff7367ff8c72f7585dff7367ff9576ff8c72ff9576ffac88ffac88ffffffffac88ffffffffac88%
  ffac88ffac88ff9576ffffffff9576ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffac88ffac88ff9f7dff846eff7367f85a5df6545cfe6862ff8c72ffac88ffac88ffffff%
  ffffffffac88ffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffffffffffffffac88ff9f7dff8c72ff9f7dff9576%
  ff7e6cff7869ff846eff8c72ff9f7dffac88ff9576ffac88ffac88ffac88ffac88ffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dff9576%
  ff7367ff7e6cfe6862fc6260ff6f65ff6f65ff7869fd6561ff6b63ff7869ff846eff7e6cff9576ff8c72ffac88ffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffff9576ffffffffac88ff846eff7e6cff6b63fd6561f7565cf34e5af85a5df5535bf4505af5515bfb605ffe6862%
  ff6f65ff7869ff8c72ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffac88ff9576ff9576ff846efc6260f6545cf5535bf14958e53651e53751e93c53%
  ec3f55e63852f14858f24b59f34d59f24a59f85a5dfe6862ff846eff9576ff9f7dff9576ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffac88ff9f7dffac88ff846eff7367fd6561f6545cf14858%
  eb3e54e93c53ea3d54e73a52e53751f04557ea3d54ea3d54f04557ea3d54f14858f34d59ff7367ff8c72ff8c72ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9576ff8c72ff9576ff9f7dff846e%
  ff7e6cfe6862f95c5ef7585df5535bf34e5aed4155ef4356eb3e54f24b59f34e5aef4457f24b59f14858ef4356ef4356%
  f7565cff7e6cffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88%
  ff9576ff9576ff7869fd6561fb605ff5535bf04557f5515bef4457e93c53ea3d54e63852ea3d54ed4155f04557ee4256%
  f04757ea3d54e23450e73a52ee4156f7565cff7367ff7e6cffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffff9f7dff9f7dff9576ffffffff7e6cff6b63fe6862f7565cfe6862f34e5afb605ff7585df7585df7565c%
  f7585de83b53f5535bfa5e5ff6545cf85a5df34e5afb605ff7585dff6f65ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9f7dff9f7dffac88ff9f7dff9576ff9576ff846efc6260%
  fa5e5ffc6260ff6f65ff6f65fa5e5ffa5e5ff7585df7585df7585df85a5df95c5eff7869ff8c72ffffffffac88ffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ff8c72fc6260ff6f65f5515bf04557f6545cf5515bff6f65ff6b63f95c5ef95c5efa5e5ffd6561f7585dff7e6cff6b63%
  ff9f7dffac88ff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffac88ffac88ff8c72ff9f7dff9576ffac88ffffffffac88ffffffffffffffac88ffffffff9576%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffb3082eda294bd11f46d42248dc2b4c%
  df304eea3d54f6545cff6b63ff8c72ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff900106%
  bb0e35b3082ebb0e35c3143bc8183fd52349e0314ffb605fffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffff}%
\gprawimage{rgb}{6.762}{2.179}{101}{101}{3.624}{3.624}{\gprawrgbimagedata}{}
\end{scope}
\gpcolor{rgb color={0.000,0.000,0.000}}
\draw[gp path] (6.780,2.197)--(6.816,2.233)--(6.852,2.269)--(6.889,2.306)--(6.925,2.342)%
  --(6.961,2.378)--(6.997,2.414)--(7.034,2.451)--(7.070,2.487)--(7.106,2.523)--(7.142,2.559)%
  --(7.179,2.596)--(7.215,2.632)--(7.251,2.668)--(7.287,2.704)--(7.324,2.741)--(7.360,2.777)%
  --(7.396,2.813)--(7.432,2.849)--(7.469,2.886)--(7.505,2.922)--(7.541,2.958)--(7.577,2.994)%
  --(7.614,3.031)--(7.650,3.067)--(7.686,3.103)--(7.722,3.139)--(7.759,3.176)--(7.795,3.212)%
  --(7.831,3.248)--(7.867,3.284)--(7.904,3.321)--(7.940,3.357)--(7.976,3.393)--(8.012,3.429)%
  --(8.048,3.465)--(8.085,3.502)--(8.121,3.538)--(8.157,3.574)--(8.193,3.610)--(8.230,3.647)%
  --(8.266,3.683)--(8.302,3.719)--(8.338,3.755)--(8.375,3.792)--(8.411,3.828)--(8.447,3.864)%
  --(8.483,3.900)--(8.520,3.937)--(8.556,3.973)--(8.592,4.009)--(8.628,4.045)--(8.665,4.082)%
  --(8.701,4.118)--(8.737,4.154)--(8.773,4.190)--(8.810,4.227)--(8.846,4.263)--(8.882,4.299)%
  --(8.918,4.335)--(8.955,4.372)--(8.991,4.408)--(9.027,4.444)--(9.063,4.480)--(9.100,4.517)%
  --(9.136,4.553)--(9.172,4.589)--(9.208,4.625)--(9.244,4.661)--(9.281,4.698)--(9.317,4.734)%
  --(9.353,4.770)--(9.389,4.806)--(9.426,4.843)--(9.462,4.879)--(9.498,4.915)--(9.534,4.951)%
  --(9.571,4.988)--(9.607,5.024)--(9.643,5.060)--(9.679,5.096)--(9.716,5.133)--(9.752,5.169)%
  --(9.788,5.205)--(9.824,5.241)--(9.861,5.278)--(9.897,5.314)--(9.933,5.350)--(9.969,5.386)%
  --(10.006,5.423)--(10.042,5.459)--(10.078,5.495)--(10.114,5.531)--(10.151,5.568)--(10.187,5.604)%
  --(10.223,5.640)--(10.259,5.676)--(10.296,5.713)--(10.332,5.749)--(10.368,5.785);
\gpcolor{color=gp lt color border}
\draw[gp path] (6.780,2.197)--(6.960,2.197);
\node[gp node right] at (6.596,2.197) {1};
\draw[gp path] (6.780,2.744)--(6.960,2.744);
\node[gp node right] at (6.596,2.744) {10};
\draw[gp path] (6.780,5.481)--(6.960,5.481);
\draw[gp path] (6.780,5.785)--(6.960,5.785);
\node[gp node right] at (6.596,5.785) {1h};
\draw[gp path] (6.780,2.197)--(6.960,2.197);
\draw[gp path] (6.780,2.362)--(6.870,2.362);
\draw[gp path] (6.780,2.580)--(6.870,2.580);
\draw[gp path] (6.780,2.691)--(6.870,2.691);
\draw[gp path] (6.780,2.744)--(6.960,2.744);
\draw[gp path] (6.780,2.909)--(6.870,2.909);
\draw[gp path] (6.780,3.127)--(6.870,3.127);
\draw[gp path] (6.780,3.238)--(6.870,3.238);
\draw[gp path] (6.780,3.292)--(6.960,3.292);
\node[gp node right] at (6.596,3.292) {$10^{2}$};
\draw[gp path] (6.780,3.456)--(6.870,3.456);
\draw[gp path] (6.780,3.674)--(6.870,3.674);
\draw[gp path] (6.780,3.786)--(6.870,3.786);
\draw[gp path] (6.780,3.839)--(6.960,3.839);
\node[gp node right] at (6.596,3.839) {$10^{3}$};
\draw[gp path] (6.780,4.004)--(6.870,4.004);
\draw[gp path] (6.780,4.221)--(6.870,4.221);
\draw[gp path] (6.780,4.333)--(6.870,4.333);
\draw[gp path] (6.780,4.386)--(6.960,4.386);
\node[gp node right] at (6.596,4.386) {$10^{4}$};
\draw[gp path] (6.780,4.551)--(6.870,4.551);
\draw[gp path] (6.780,4.769)--(6.870,4.769);
\draw[gp path] (6.780,4.880)--(6.870,4.880);
\draw[gp path] (6.780,4.933)--(6.960,4.933);
\node[gp node right] at (6.596,4.933) {$10^{5}$};
\draw[gp path] (6.780,5.098)--(6.870,5.098);
\draw[gp path] (6.780,5.316)--(6.870,5.316);
\draw[gp path] (6.780,5.428)--(6.870,5.428);
\draw[gp path] (6.780,5.481)--(6.960,5.481);
\draw[gp path] (6.780,5.645)--(6.870,5.645);
\draw[gp path] (6.780,2.197)--(6.780,2.377);
\node[gp node center] at (6.780,1.889) {1};
\draw[gp path] (7.327,2.197)--(7.327,2.377);
\node[gp node center] at (7.327,1.889) {10};
\draw[gp path] (10.064,2.197)--(10.064,2.377);
\draw[gp path] (10.368,2.197)--(10.368,2.377);
\node[gp node center] at (10.368,1.889) {1h};
\draw[gp path] (6.780,2.197)--(6.780,2.377);
\draw[gp path] (6.945,2.197)--(6.945,2.287);
\draw[gp path] (7.163,2.197)--(7.163,2.287);
\draw[gp path] (7.274,2.197)--(7.274,2.287);
\draw[gp path] (7.327,2.197)--(7.327,2.377);
\draw[gp path] (7.492,2.197)--(7.492,2.287);
\draw[gp path] (7.710,2.197)--(7.710,2.287);
\draw[gp path] (7.821,2.197)--(7.821,2.287);
\draw[gp path] (7.875,2.197)--(7.875,2.377);
\node[gp node center] at (7.875,1.889) {$10^{2}$};
\draw[gp path] (8.039,2.197)--(8.039,2.287);
\draw[gp path] (8.257,2.197)--(8.257,2.287);
\draw[gp path] (8.369,2.197)--(8.369,2.287);
\draw[gp path] (8.422,2.197)--(8.422,2.377);
\node[gp node center] at (8.422,1.889) {$10^{3}$};
\draw[gp path] (8.587,2.197)--(8.587,2.287);
\draw[gp path] (8.804,2.197)--(8.804,2.287);
\draw[gp path] (8.916,2.197)--(8.916,2.287);
\draw[gp path] (8.969,2.197)--(8.969,2.377);
\node[gp node center] at (8.969,1.889) {$10^{4}$};
\draw[gp path] (9.134,2.197)--(9.134,2.287);
\draw[gp path] (9.352,2.197)--(9.352,2.287);
\draw[gp path] (9.463,2.197)--(9.463,2.287);
\draw[gp path] (9.516,2.197)--(9.516,2.377);
\node[gp node center] at (9.516,1.889) {$10^{5}$};
\draw[gp path] (9.681,2.197)--(9.681,2.287);
\draw[gp path] (9.899,2.197)--(9.899,2.287);
\draw[gp path] (10.011,2.197)--(10.011,2.287);
\draw[gp path] (10.064,2.197)--(10.064,2.377);
\draw[gp path] (10.228,2.197)--(10.228,2.287);
\draw[gp path] (6.780,5.785)--(6.780,2.197)--(10.368,2.197);
\gpfill{rgb color={1.000,0.998,0.996}} (10.920,2.197)--(11.138,2.197)--(11.138,2.226)--(10.920,2.226)--cycle;
\gpfill{rgb color={1.000,0.934,0.895}} (10.920,2.225)--(11.138,2.225)--(11.138,2.254)--(10.920,2.254)--cycle;
\gpfill{rgb color={1.000,0.871,0.793}} (10.920,2.253)--(11.138,2.253)--(11.138,2.282)--(10.920,2.282)--cycle;
\gpfill{rgb color={1.000,0.807,0.692}} (10.920,2.281)--(11.138,2.281)--(11.138,2.310)--(10.920,2.310)--cycle;
\gpfill{rgb color={1.000,0.743,0.590}} (10.920,2.309)--(11.138,2.309)--(11.138,2.313)--(10.920,2.313)--cycle;
\gpfill{rgb color={1.000,0.737,0.580}} (10.920,2.312)--(11.138,2.312)--(11.138,2.338)--(10.920,2.338)--cycle;
\gpfill{rgb color={1.000,0.731,0.575}} (10.920,2.337)--(11.138,2.337)--(11.138,2.366)--(10.920,2.366)--cycle;
\gpfill{rgb color={1.000,0.724,0.570}} (10.920,2.365)--(11.138,2.365)--(11.138,2.394)--(10.920,2.394)--cycle;
\gpfill{rgb color={1.000,0.717,0.564}} (10.920,2.393)--(11.138,2.393)--(11.138,2.422)--(10.920,2.422)--cycle;
\gpfill{rgb color={1.000,0.710,0.559}} (10.920,2.421)--(11.138,2.421)--(11.138,2.450)--(10.920,2.450)--cycle;
\gpfill{rgb color={1.000,0.703,0.553}} (10.920,2.449)--(11.138,2.449)--(11.138,2.478)--(10.920,2.478)--cycle;
\gpfill{rgb color={1.000,0.696,0.548}} (10.920,2.477)--(11.138,2.477)--(11.138,2.506)--(10.920,2.506)--cycle;
\gpfill{rgb color={1.000,0.689,0.542}} (10.920,2.505)--(11.138,2.505)--(11.138,2.534)--(10.920,2.534)--cycle;
\gpfill{rgb color={1.000,0.682,0.537}} (10.920,2.533)--(11.138,2.533)--(11.138,2.562)--(10.920,2.562)--cycle;
\gpfill{rgb color={1.000,0.675,0.531}} (10.920,2.561)--(11.138,2.561)--(11.138,2.590)--(10.920,2.590)--cycle;
\gpfill{rgb color={1.000,0.668,0.526}} (10.920,2.589)--(11.138,2.589)--(11.138,2.618)--(10.920,2.618)--cycle;
\gpfill{rgb color={1.000,0.661,0.520}} (10.920,2.617)--(11.138,2.617)--(11.138,2.646)--(10.920,2.646)--cycle;
\gpfill{rgb color={1.000,0.654,0.515}} (10.920,2.645)--(11.138,2.645)--(11.138,2.674)--(10.920,2.674)--cycle;
\gpfill{rgb color={1.000,0.647,0.509}} (10.920,2.673)--(11.138,2.673)--(11.138,2.702)--(10.920,2.702)--cycle;
\gpfill{rgb color={1.000,0.640,0.504}} (10.920,2.701)--(11.138,2.701)--(11.138,2.730)--(10.920,2.730)--cycle;
\gpfill{rgb color={1.000,0.633,0.498}} (10.920,2.729)--(11.138,2.729)--(11.138,2.758)--(10.920,2.758)--cycle;
\gpfill{rgb color={1.000,0.626,0.493}} (10.920,2.757)--(11.138,2.757)--(11.138,2.786)--(10.920,2.786)--cycle;
\gpfill{rgb color={1.000,0.619,0.487}} (10.920,2.785)--(11.138,2.785)--(11.138,2.814)--(10.920,2.814)--cycle;
\gpfill{rgb color={1.000,0.612,0.482}} (10.920,2.813)--(11.138,2.813)--(11.138,2.842)--(10.920,2.842)--cycle;
\gpfill{rgb color={1.000,0.605,0.476}} (10.920,2.841)--(11.138,2.841)--(11.138,2.870)--(10.920,2.870)--cycle;
\gpfill{rgb color={1.000,0.598,0.471}} (10.920,2.869)--(11.138,2.869)--(11.138,2.892)--(10.920,2.892)--cycle;
\gpfill{rgb color={1.000,0.592,0.467}} (10.920,2.891)--(11.138,2.891)--(11.138,2.898)--(10.920,2.898)--cycle;
\gpfill{rgb color={1.000,0.590,0.466}} (10.920,2.897)--(11.138,2.897)--(11.138,2.926)--(10.920,2.926)--cycle;
\gpfill{rgb color={1.000,0.581,0.462}} (10.920,2.925)--(11.138,2.925)--(11.138,2.954)--(10.920,2.954)--cycle;
\gpfill{rgb color={1.000,0.573,0.458}} (10.920,2.953)--(11.138,2.953)--(11.138,2.982)--(10.920,2.982)--cycle;
\gpfill{rgb color={1.000,0.564,0.454}} (10.920,2.981)--(11.138,2.981)--(11.138,3.010)--(10.920,3.010)--cycle;
\gpfill{rgb color={1.000,0.555,0.450}} (10.920,3.009)--(11.138,3.009)--(11.138,3.038)--(10.920,3.038)--cycle;
\gpfill{rgb color={1.000,0.546,0.446}} (10.920,3.037)--(11.138,3.037)--(11.138,3.066)--(10.920,3.066)--cycle;
\gpfill{rgb color={1.000,0.538,0.442}} (10.920,3.065)--(11.138,3.065)--(11.138,3.095)--(10.920,3.095)--cycle;
\gpfill{rgb color={1.000,0.529,0.438}} (10.920,3.094)--(11.138,3.094)--(11.138,3.123)--(10.920,3.123)--cycle;
\gpfill{rgb color={1.000,0.520,0.434}} (10.920,3.122)--(11.138,3.122)--(11.138,3.151)--(10.920,3.151)--cycle;
\gpfill{rgb color={1.000,0.511,0.430}} (10.920,3.150)--(11.138,3.150)--(11.138,3.179)--(10.920,3.179)--cycle;
\gpfill{rgb color={1.000,0.503,0.426}} (10.920,3.178)--(11.138,3.178)--(11.138,3.207)--(10.920,3.207)--cycle;
\gpfill{rgb color={1.000,0.494,0.422}} (10.920,3.206)--(11.138,3.206)--(11.138,3.235)--(10.920,3.235)--cycle;
\gpfill{rgb color={1.000,0.485,0.418}} (10.920,3.234)--(11.138,3.234)--(11.138,3.263)--(10.920,3.263)--cycle;
\gpfill{rgb color={1.000,0.476,0.414}} (10.920,3.262)--(11.138,3.262)--(11.138,3.291)--(10.920,3.291)--cycle;
\gpfill{rgb color={1.000,0.468,0.410}} (10.920,3.290)--(11.138,3.290)--(11.138,3.319)--(10.920,3.319)--cycle;
\gpfill{rgb color={1.000,0.459,0.406}} (10.920,3.318)--(11.138,3.318)--(11.138,3.347)--(10.920,3.347)--cycle;
\gpfill{rgb color={1.000,0.450,0.402}} (10.920,3.346)--(11.138,3.346)--(11.138,3.375)--(10.920,3.375)--cycle;
\gpfill{rgb color={1.000,0.441,0.398}} (10.920,3.374)--(11.138,3.374)--(11.138,3.403)--(10.920,3.403)--cycle;
\gpfill{rgb color={1.000,0.433,0.394}} (10.920,3.402)--(11.138,3.402)--(11.138,3.431)--(10.920,3.431)--cycle;
\gpfill{rgb color={1.000,0.424,0.390}} (10.920,3.430)--(11.138,3.430)--(11.138,3.459)--(10.920,3.459)--cycle;
\gpfill{rgb color={1.000,0.415,0.386}} (10.920,3.458)--(11.138,3.458)--(11.138,3.471)--(10.920,3.471)--cycle;
\gpfill{rgb color={1.000,0.412,0.384}} (10.920,3.470)--(11.138,3.470)--(11.138,3.487)--(10.920,3.487)--cycle;
\gpfill{rgb color={0.998,0.407,0.383}} (10.920,3.486)--(11.138,3.486)--(11.138,3.515)--(10.920,3.515)--cycle;
\gpfill{rgb color={0.995,0.400,0.381}} (10.920,3.514)--(11.138,3.514)--(11.138,3.543)--(10.920,3.543)--cycle;
\gpfill{rgb color={0.992,0.393,0.378}} (10.920,3.542)--(11.138,3.542)--(11.138,3.571)--(10.920,3.571)--cycle;
\gpfill{rgb color={0.988,0.385,0.376}} (10.920,3.570)--(11.138,3.570)--(11.138,3.599)--(10.920,3.599)--cycle;
\gpfill{rgb color={0.985,0.378,0.374}} (10.920,3.598)--(11.138,3.598)--(11.138,3.627)--(10.920,3.627)--cycle;
\gpfill{rgb color={0.982,0.370,0.372}} (10.920,3.626)--(11.138,3.626)--(11.138,3.655)--(10.920,3.655)--cycle;
\gpfill{rgb color={0.979,0.363,0.369}} (10.920,3.654)--(11.138,3.654)--(11.138,3.683)--(10.920,3.683)--cycle;
\gpfill{rgb color={0.975,0.356,0.367}} (10.920,3.682)--(11.138,3.682)--(11.138,3.711)--(10.920,3.711)--cycle;
\gpfill{rgb color={0.972,0.348,0.365}} (10.920,3.710)--(11.138,3.710)--(11.138,3.739)--(10.920,3.739)--cycle;
\gpfill{rgb color={0.969,0.341,0.362}} (10.920,3.738)--(11.138,3.738)--(11.138,3.767)--(10.920,3.767)--cycle;
\gpfill{rgb color={0.966,0.333,0.360}} (10.920,3.766)--(11.138,3.766)--(11.138,3.795)--(10.920,3.795)--cycle;
\gpfill{rgb color={0.963,0.326,0.358}} (10.920,3.794)--(11.138,3.794)--(11.138,3.823)--(10.920,3.823)--cycle;
\gpfill{rgb color={0.959,0.319,0.356}} (10.920,3.822)--(11.138,3.822)--(11.138,3.851)--(10.920,3.851)--cycle;
\gpfill{rgb color={0.956,0.311,0.353}} (10.920,3.850)--(11.138,3.850)--(11.138,3.879)--(10.920,3.879)--cycle;
\gpfill{rgb color={0.953,0.304,0.351}} (10.920,3.878)--(11.138,3.878)--(11.138,3.907)--(10.920,3.907)--cycle;
\gpfill{rgb color={0.950,0.296,0.349}} (10.920,3.906)--(11.138,3.906)--(11.138,3.935)--(10.920,3.935)--cycle;
\gpfill{rgb color={0.946,0.289,0.347}} (10.920,3.934)--(11.138,3.934)--(11.138,3.963)--(10.920,3.963)--cycle;
\gpfill{rgb color={0.943,0.282,0.344}} (10.920,3.962)--(11.138,3.962)--(11.138,3.992)--(10.920,3.992)--cycle;
\gpfill{rgb color={0.940,0.274,0.342}} (10.920,3.991)--(11.138,3.991)--(11.138,4.020)--(10.920,4.020)--cycle;
\gpfill{rgb color={0.937,0.266,0.340}} (10.920,4.019)--(11.138,4.019)--(11.138,4.048)--(10.920,4.048)--cycle;
\gpfill{rgb color={0.933,0.259,0.337}} (10.920,4.047)--(11.138,4.047)--(11.138,4.049)--(10.920,4.049)--cycle;
\gpfill{rgb color={0.933,0.259,0.337}} (10.920,4.048)--(11.138,4.048)--(11.138,4.076)--(10.920,4.076)--cycle;
\gpfill{rgb color={0.928,0.252,0.334}} (10.920,4.075)--(11.138,4.075)--(11.138,4.104)--(10.920,4.104)--cycle;
\gpfill{rgb color={0.923,0.246,0.332}} (10.920,4.103)--(11.138,4.103)--(11.138,4.132)--(10.920,4.132)--cycle;
\gpfill{rgb color={0.918,0.239,0.329}} (10.920,4.131)--(11.138,4.131)--(11.138,4.160)--(10.920,4.160)--cycle;
\gpfill{rgb color={0.912,0.232,0.326}} (10.920,4.159)--(11.138,4.159)--(11.138,4.188)--(10.920,4.188)--cycle;
\gpfill{rgb color={0.907,0.226,0.323}} (10.920,4.187)--(11.138,4.187)--(11.138,4.216)--(10.920,4.216)--cycle;
\gpfill{rgb color={0.902,0.219,0.320}} (10.920,4.215)--(11.138,4.215)--(11.138,4.244)--(10.920,4.244)--cycle;
\gpfill{rgb color={0.896,0.213,0.317}} (10.920,4.243)--(11.138,4.243)--(11.138,4.272)--(10.920,4.272)--cycle;
\gpfill{rgb color={0.891,0.206,0.315}} (10.920,4.271)--(11.138,4.271)--(11.138,4.300)--(10.920,4.300)--cycle;
\gpfill{rgb color={0.886,0.199,0.312}} (10.920,4.299)--(11.138,4.299)--(11.138,4.328)--(10.920,4.328)--cycle;
\gpfill{rgb color={0.880,0.193,0.309}} (10.920,4.327)--(11.138,4.327)--(11.138,4.356)--(10.920,4.356)--cycle;
\gpfill{rgb color={0.875,0.186,0.306}} (10.920,4.355)--(11.138,4.355)--(11.138,4.384)--(10.920,4.384)--cycle;
\gpfill{rgb color={0.870,0.179,0.303}} (10.920,4.383)--(11.138,4.383)--(11.138,4.412)--(10.920,4.412)--cycle;
\gpfill{rgb color={0.864,0.173,0.300}} (10.920,4.411)--(11.138,4.411)--(11.138,4.440)--(10.920,4.440)--cycle;
\gpfill{rgb color={0.859,0.166,0.297}} (10.920,4.439)--(11.138,4.439)--(11.138,4.468)--(10.920,4.468)--cycle;
\gpfill{rgb color={0.854,0.159,0.295}} (10.920,4.467)--(11.138,4.467)--(11.138,4.496)--(10.920,4.496)--cycle;
\gpfill{rgb color={0.848,0.153,0.292}} (10.920,4.495)--(11.138,4.495)--(11.138,4.524)--(10.920,4.524)--cycle;
\gpfill{rgb color={0.843,0.146,0.289}} (10.920,4.523)--(11.138,4.523)--(11.138,4.552)--(10.920,4.552)--cycle;
\gpfill{rgb color={0.838,0.139,0.286}} (10.920,4.551)--(11.138,4.551)--(11.138,4.580)--(10.920,4.580)--cycle;
\gpfill{rgb color={0.833,0.133,0.283}} (10.920,4.579)--(11.138,4.579)--(11.138,4.608)--(10.920,4.608)--cycle;
\gpfill{rgb color={0.827,0.126,0.280}} (10.920,4.607)--(11.138,4.607)--(11.138,4.628)--(10.920,4.628)--cycle;
\gpfill{rgb color={0.823,0.121,0.278}} (10.920,4.627)--(11.138,4.627)--(11.138,4.636)--(10.920,4.636)--cycle;
\gpfill{rgb color={0.822,0.120,0.277}} (10.920,4.635)--(11.138,4.635)--(11.138,4.664)--(10.920,4.664)--cycle;
\gpfill{rgb color={0.815,0.115,0.272}} (10.920,4.663)--(11.138,4.663)--(11.138,4.692)--(10.920,4.692)--cycle;
\gpfill{rgb color={0.809,0.111,0.267}} (10.920,4.691)--(11.138,4.691)--(11.138,4.720)--(10.920,4.720)--cycle;
\gpfill{rgb color={0.802,0.106,0.262}} (10.920,4.719)--(11.138,4.719)--(11.138,4.748)--(10.920,4.748)--cycle;
\gpfill{rgb color={0.796,0.101,0.256}} (10.920,4.747)--(11.138,4.747)--(11.138,4.776)--(10.920,4.776)--cycle;
\gpfill{rgb color={0.789,0.096,0.251}} (10.920,4.775)--(11.138,4.775)--(11.138,4.804)--(10.920,4.804)--cycle;
\gpfill{rgb color={0.783,0.092,0.246}} (10.920,4.803)--(11.138,4.803)--(11.138,4.832)--(10.920,4.832)--cycle;
\gpfill{rgb color={0.776,0.087,0.241}} (10.920,4.831)--(11.138,4.831)--(11.138,4.860)--(10.920,4.860)--cycle;
\gpfill{rgb color={0.770,0.082,0.236}} (10.920,4.859)--(11.138,4.859)--(11.138,4.889)--(10.920,4.889)--cycle;
\gpfill{rgb color={0.763,0.077,0.231}} (10.920,4.888)--(11.138,4.888)--(11.138,4.917)--(10.920,4.917)--cycle;
\gpfill{rgb color={0.757,0.073,0.225}} (10.920,4.916)--(11.138,4.916)--(11.138,4.945)--(10.920,4.945)--cycle;
\gpfill{rgb color={0.750,0.068,0.220}} (10.920,4.944)--(11.138,4.944)--(11.138,4.973)--(10.920,4.973)--cycle;
\gpfill{rgb color={0.744,0.063,0.215}} (10.920,4.972)--(11.138,4.972)--(11.138,5.001)--(10.920,5.001)--cycle;
\gpfill{rgb color={0.737,0.058,0.210}} (10.920,5.000)--(11.138,5.000)--(11.138,5.029)--(10.920,5.029)--cycle;
\gpfill{rgb color={0.731,0.054,0.205}} (10.920,5.028)--(11.138,5.028)--(11.138,5.057)--(10.920,5.057)--cycle;
\gpfill{rgb color={0.725,0.049,0.200}} (10.920,5.056)--(11.138,5.056)--(11.138,5.085)--(10.920,5.085)--cycle;
\gpfill{rgb color={0.718,0.044,0.195}} (10.920,5.084)--(11.138,5.084)--(11.138,5.113)--(10.920,5.113)--cycle;
\gpfill{rgb color={0.712,0.039,0.190}} (10.920,5.112)--(11.138,5.112)--(11.138,5.141)--(10.920,5.141)--cycle;
\gpfill{rgb color={0.705,0.035,0.184}} (10.920,5.140)--(11.138,5.140)--(11.138,5.169)--(10.920,5.169)--cycle;
\gpfill{rgb color={0.699,0.030,0.179}} (10.920,5.168)--(11.138,5.168)--(11.138,5.197)--(10.920,5.197)--cycle;
\gpfill{rgb color={0.692,0.025,0.174}} (10.920,5.196)--(11.138,5.196)--(11.138,5.207)--(10.920,5.207)--cycle;
\gpfill{rgb color={0.690,0.024,0.172}} (10.920,5.206)--(11.138,5.206)--(11.138,5.225)--(10.920,5.225)--cycle;
\gpfill{rgb color={0.686,0.023,0.167}} (10.920,5.224)--(11.138,5.224)--(11.138,5.253)--(10.920,5.253)--cycle;
\gpfill{rgb color={0.678,0.022,0.159}} (10.920,5.252)--(11.138,5.252)--(11.138,5.281)--(10.920,5.281)--cycle;
\gpfill{rgb color={0.671,0.020,0.150}} (10.920,5.280)--(11.138,5.280)--(11.138,5.309)--(10.920,5.309)--cycle;
\gpfill{rgb color={0.664,0.019,0.142}} (10.920,5.308)--(11.138,5.308)--(11.138,5.337)--(10.920,5.337)--cycle;
\gpfill{rgb color={0.657,0.018,0.134}} (10.920,5.336)--(11.138,5.336)--(11.138,5.365)--(10.920,5.365)--cycle;
\gpfill{rgb color={0.650,0.017,0.125}} (10.920,5.364)--(11.138,5.364)--(11.138,5.393)--(10.920,5.393)--cycle;
\gpfill{rgb color={0.643,0.016,0.117}} (10.920,5.392)--(11.138,5.392)--(11.138,5.421)--(10.920,5.421)--cycle;
\gpfill{rgb color={0.636,0.015,0.109}} (10.920,5.420)--(11.138,5.420)--(11.138,5.449)--(10.920,5.449)--cycle;
\gpfill{rgb color={0.629,0.014,0.100}} (10.920,5.448)--(11.138,5.448)--(11.138,5.477)--(10.920,5.477)--cycle;
\gpfill{rgb color={0.622,0.013,0.092}} (10.920,5.476)--(11.138,5.476)--(11.138,5.505)--(10.920,5.505)--cycle;
\gpfill{rgb color={0.615,0.011,0.083}} (10.920,5.504)--(11.138,5.504)--(11.138,5.533)--(10.920,5.533)--cycle;
\gpfill{rgb color={0.608,0.010,0.075}} (10.920,5.532)--(11.138,5.532)--(11.138,5.561)--(10.920,5.561)--cycle;
\gpfill{rgb color={0.601,0.009,0.067}} (10.920,5.560)--(11.138,5.560)--(11.138,5.589)--(10.920,5.589)--cycle;
\gpfill{rgb color={0.594,0.008,0.058}} (10.920,5.588)--(11.138,5.588)--(11.138,5.617)--(10.920,5.617)--cycle;
\gpfill{rgb color={0.587,0.007,0.050}} (10.920,5.616)--(11.138,5.616)--(11.138,5.645)--(10.920,5.645)--cycle;
\gpfill{rgb color={0.580,0.006,0.042}} (10.920,5.644)--(11.138,5.644)--(11.138,5.673)--(10.920,5.673)--cycle;
\gpfill{rgb color={0.573,0.005,0.033}} (10.920,5.672)--(11.138,5.672)--(11.138,5.701)--(10.920,5.701)--cycle;
\gpfill{rgb color={0.566,0.003,0.025}} (10.920,5.700)--(11.138,5.700)--(11.138,5.729)--(10.920,5.729)--cycle;
\gpfill{rgb color={0.559,0.002,0.017}} (10.920,5.728)--(11.138,5.728)--(11.138,5.757)--(10.920,5.757)--cycle;
\gpfill{rgb color={0.552,0.001,0.008}} (10.920,5.756)--(11.138,5.756)--(11.138,5.785)--(10.920,5.785)--cycle;
\draw[gp path] (10.920,2.197)--(11.138,2.197)--(11.138,5.785)--(10.920,5.785)--cycle;
\draw[gp path] (11.138,2.197)--(10.958,2.197);
\node[gp node left] at (11.139,2.197) {0};
\draw[gp path] (10.920,2.197)--(11.100,2.197);
\draw[gp path] (11.138,2.557)--(10.958,2.557);
\node[gp node left] at (11.139,2.557) {1};
\draw[gp path] (10.920,2.557)--(11.100,2.557);
\draw[gp path] (11.138,3.393)--(10.958,3.393);
\node[gp node left] at (11.139,3.393) {10};
\draw[gp path] (10.920,3.393)--(11.100,3.393);
\draw[gp path] (11.138,4.589)--(10.958,4.589);
\node[gp node left] at (11.139,4.589) {$10^2$};
\draw[gp path] (10.920,4.589)--(11.100,4.589);
\draw[gp path] (11.138,5.785)--(10.958,5.785);
\node[gp node left] at (11.139,5.785) {$\ge10^3$};
\draw[gp path] (10.920,5.785)--(11.100,5.785);
\node[gp node center] at (8.574,5.426) {Unlabelled};
%% coordinates of the plot area
\gpdefrectangularnode{gp plot 2}{\pgfpoint{6.780cm}{2.197cm}}{\pgfpoint{10.368cm}{5.785cm}}
\end{tikzpicture}
%% gnuplot variables