### 2016-04-07 paper/gen-graph-unconnected-heatmap.tex

\begin{tikzpicture}[gnuplot]
%% generated with GNUPLOT 5.0p0 (Lua 5.2; terminal rev. 99, script rev. 100)
%% Thu 07 Apr 2016 11:12:12 BST
\tikzset{every node/.append style={font={\scriptsize}}}
% \path (0.000,0.000) rectangle (10.922,7.366);
\gpcolor{color=gp lt color border}
\gpsetlinetype{gp lt border}
\gpsetdashtype{gp dt solid}
\gpsetlinewidth{1.00}
\draw[gp path] (1.320,5.785)--(1.320,2.197)--(4.908,2.197);
\node[gp node center,rotate=-270] at (0.246,3.991) {Clique Runtime (ms)};
\node[gp node center] at (3.114,1.427) {MAC Runtime (ms)};
\begin{scope}
\clip (1.320,5.785) rectangle (4.908,2.197);
\def\gprawrgbimagedata{%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffac88ffffffffffffffffffffffffffffffffac88ffac88ffac88ffffffff8c72ff9f7d%
  ffac88ff8c72ff9f7dff95768b0000ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffd01d45ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffffffffffffffffffffac88ca1941ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffac88ffffffffffffc7173effffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffc6163effffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffffffc6163dffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffc6163effffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffac88ffffffffffffffffffffffffffac88c7173effffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ff9f7dffffffffac88ffffffffffffffffffffffffc1133affffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffc8183fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffc7173fffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffac88%
  ffffffffac88ffffffffffffffffffffac88ffffffffac88ffffffffac88ffffffc6163effffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffff9576ffffffffffff%
  c91840ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffac88ffffffffffffffffffffac88c4153cffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffff9f7dffac88ffffffffffffffffffffac88c4143cffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7d%
  ffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffac88ffac88ff9f7dff9f7dffac88ffffffc3143b%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffac88%
  ffac88ffffffff9f7dffffffc11239ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffac88ffffffffffffffac88ffffffffac88ffffffffffffffffffffffffffac88%
  ffac88ffac88ffffffffac88ffffffffffffffffffffac88ffffffc4153cffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffac88ffffffffffffffffffffffffffffffffac88ffffffff9f7dff9f7dff9f7dffac88ffffffc2143bffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffac88ffac88ff9576ff9f7dff9f7dffac88%
  ffffffffffffffac88c6163dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffffffffffff%
  ffffffffffffff9576ffac88ff9576ffffffff8c72ffffffc11239ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffff9f7dffffffffac88ffac88ffffffff9f7dff9f7dffffffff9f7dffffffff9f7dc5153cffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffac88ffac88ffac88ffffffff8c72ffffffff9576%
  ffffffff9576bd1037ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ff9f7dffac88ff9f7dffffff%
  ffac88ffac88ff9576ff9f7dff9f7dff9576ffac88c2133affffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88%
  ffac88ff9f7dffac88ffffffffac88ff9f7dff9f7dff9f7dff9f7dff9576ffac88ff9f7dc2133bffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffac88ffffffffffff%
  ffffffffffffffac88ffffffffac88ffffffff9576ffac88ffac88ffffffffac88ff8c72ffffffffac88ff8c72ff8c72%
  ffffffc01239ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff9f7dffffffff9f7dffac88ffac88ffac88ffffffffac88ff9f7dffac88%
  ff9f7dffac88ff9576ff9f7dff9f7dff9f7dc11239ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffac88ffffffffac88ffac88ffac88%
  ffac88ff8c72ff9576ffac88ffac88ff9f7dff9576ff846effac88ff846eff846ec4143cffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffac88ff9f7dffac88ffffffffffffffffffffffff%
  ffac88ffac88ff9f7dff9f7dff9576ff9f7dff9f7dff9576ff9f7dff9576ff9f7dff9576ffac88ff9576ff9f7dff8c72%
  c1133affffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffac88ffffffffac88ff9f7dffac88ffac88ff9576ffac88ffac88ff846effac88ff9f7d%
  ff9f7dff9576ff9f7dffffffff9576bd1036ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ff9f7dffac88ff9f7dff9f7dff9f7dff9576ffffff%
  ffac88ffac88ffffffff9576ffffffff9576ff846eff9f7dffac88ff9f7dc11239ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffff9576ff9576%
  ff8c72ffffffff9f7dff9f7dff9576ff8c72ff9f7dff9f7dffac88ff9f7dff9f7dff9576ff9f7dff9576ff9576bf1138%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffac88ffffffffffffffffffffffffffac88ffffff%
  ffffffff9f7dffffffffffffff9f7dff9576ff9576ff9576ffac88ff9576ffffffffffffffac88ff846eff9f7dff9576%
  ff846effffffffac88ff9f7dc11239ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffff9f7dffffffffffffffac88ffffffffffffff9f7dff9f7dffffffffac88ff8c72ff8c72ff9f7dff7e6cff9f7d%
  ff8c72ffffffff9576ffac88ff7e6cff9576ff7e6cff846eff7869c2133affffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffac88ffffffff9576ffac88ffac88ffac88ffac88%
  ff9576ffac88ff9f7dffac88ff9f7dffac88ff8c72ffac88ff9576ffac88ff9576ff8c72ff7e6cff846ec2133affffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffff%
  ffffffffffffff9f7dffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffffffffffffffac88ffffff%
  ffac88ff846eff9576ff9f7dff9f7dffac88ff8c72ffac88ff9576ffac88ff7869ff8c72ff846eff9f7dff9f7dff846e%
  ff7e6cff6b63ffffffc5163dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ff9f7dffac88ffac88ffac88ffffffff9576ffac88ff9576ff846eff9576ff9576ff8c72ff9576ff9f7d%
  ff8c72ff9f7dff7869ff846eff7e6cff7e6cff8c72ff8c72c1133affffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ffffffffac88ffffffffffffffac88ffffffffffffffac88ffac88ff9f7dff846effac88%
  ff8c72ff9f7dff8c72ff9576ff846eff846effac88ff7367ff7e6cff8c72ffffffff7367ff8c72c2133bffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffac88ffffffffac88ffac88ff9f7dffffffffac88ffac88ff846eff8c72%
  ffac88ffac88ff7e6cff9f7dff9576ff846eff9576ff9f7dff9f7dff9f7dff7869ff7367ff846eff9576ff846eff7e6c%
  ff7e6cff7e6cc4153cffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffff%
  ffac88ffac88ffffffff9576ffffffff9f7dff9576ff9576ffffffff9576ff8c72ff8c72ff7869ff9576ff8c72ff8c72%
  ff8c72ff9576ff7e6cff7e6cff7367ff9f7dff7e6cc7173effffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffac88ffffffffac88%
  ffac88ffffffff9f7dffffffffffffffffffff9f7dffffffff9f7dffac88ff9576ff9576ff846eff9576ff7869ff8c72%
  ffffffffac88ffac88ff8c72ff9576ff7e6cff9576ff7e6cfe6862ff9f7dff9f7dff7367c6163effffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffac88ffffffffac88ffac88ffffffff9f7dffffffff9f7dffffffff9f7dffffffffac88ff9576%
  ff9576ff8c72ff9576ff8c72ff9576ff7e6cff8c72ff8c72ff8c72ff9576ff8c72ff846eff7e6cff846eff7e6cff7367%
  ff7869c6163dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffac88ff9f7dff8c72%
  ff9576ff9576ffac88ff9f7dff9f7dff8c72ff9576ff9f7dff7e6cffac88ff7e6cff9f7dff846eff7e6cff7869ff8c72%
  ff8c72ff6f65ff7367ff9576ff8c72ff7e6ccd1b43ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffff9f7dffffffffffffffffffffffffffffffffffffffffffff9f7dffac88ff9576%
  ffac88ffac88ffac88ff9f7dffac88ff9576ff9f7dff9f7dff8c72ff9f7dff7e6cff846eff7869ff7e6cff9f7dff846e%
  ff846eff8c72ff9f7dff8c72ff7869ff846eff7e6cff9f7dff7e6cff7e6cff9f7dc91940ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffff%
  ffffffffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffac88ffac88ffffff%
  ffac88ffffffffffffffac88ffac88ff9f7dff9f7dff9f7dff9f7dffac88ffffffff846eff9576ff846eff9f7dff8c72%
  ff9576ff9f7dff8c72ff8c72ff7869ff9f7dfd6561ff7869ff8c72ff846eff7367ffac88ff9576ff8c72ff9576ff7367%
  ca1940ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffac88ff9f7dffffffffac88ffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffac88ffffffffffffff9576ff9576ffffffffffffff9576ffffffffac88ffffff%
  ffac88ff8c72ff9f7dff9f7dff8c72ffac88fe6862ff7e6cff8c72ff9f7dff7e6cff8c72ff9f7dff7869ff9576ff846e%
  ff8c72ff7e6cff9f7dfd6561ff7869ca1940ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffac88ff9f7d%
  ff9f7dff846eff8c72ff9576ff9576ff846eff9576ff9f7dff846eff9f7dff9f7dff6b63ff7e6cff9576ff846effac88%
  ff846eff8c72ff9576ff846eff7e6cff9576ff9576ff846eff8c72ff7367c91840ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffff9f7dff9f7dffac88ff9f7d%
  ffffffffac88ffffffff9f7dffffffff8c72ffac88ff9f7dff9576ffac88ffffffffac88ff9576ff9f7dffac88ff7367%
  ff7869ffac88ff8c72ff9576ff7869ff9f7dfd6561ff9576ff9f7dff846efc6260ff7367ff6f65ff846eff846ec7173e%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffac88ffffffffac88ffffffff9f7dff9f7dffffffffffffffac88ff9f7dff9576ff9f7dff9f7dff9f7dff9f7dffac88%
  ff6b63ff7e6cff846eff7e6cff7e6cff6f65ff7367ff7869ff7e6cff8c72ff7869ff9f7dff9f7dff7e6cff9576ff9f7d%
  ff9f7dff846efe6862ff9576c91940ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffffff%
  ffac88ffffffffffffffac88ffac88ffffffffffffffffffffac88ffac88ff9f7dff9576ff9f7dffac88ff9f7dffffff%
  ffffffffac88ff846effac88ff9f7dff9576ff7869ff9f7dff7367ff6f65ff7869ff6f65ff846eff7367ff7367ff846e%
  ff846eff7367ff7367ff7869ff8c72ff8c72ff7e6cff7367ff9f7dce1c44ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffac88ffac88ffffffffffffffac88ffac88ffac88ffac88ffac88ffac88ffac88ffffffff9f7dffac88%
  ffffffff9f7dff8c72ff9f7dff8c72ffffffff9576ff846eff8c72ff8c72ff9576ff846eff7869ff7869ff7367ff7869%
  ff7367fe6862ff8c72ff7367ff846eff7e6cff8c72ff8c72ff7869ff846eff6f65ffac88fd6561ff846eca1941ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffac88ffac88%
  ffac88ff9f7dffffffff9f7dffffffffac88ff9576ff9f7dff9f7dff9f7dff7e6cffac88ff846eff7869ff846eff9576%
  ff846eff7e6cff6b63ff846eff7869ff6b63ff9576ff9576ff6f65ff846eff846eff7367ffac88ff6b63ff7e6cff7e6c%
  ffac88ff8c72ff846ecc1b42ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffac88ffffffffffff%
  ffffffffffffff9f7dffffffff9f7dffac88ffac88ffffffffffffffac88ffac88ffac88ff9f7dff9f7dff8c72ff9576%
  ff8c72ff9576ff8c72ff7869ff9576ff6f65ff6f65ff7367ff9576ff846eff8c72ff846eff7869ff7367ff6f65ff7869%
  ff7e6cff6f65ff9576ff8c72ff7869ff7367ff6f65ff9576ca1941ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffffffffac88ffffffffac88ffffffffffff%
  ffac88ffffffffac88ffffffffac88ffac88ffac88ffac88ffac88ffffffff9f7dff9f7dffffffffac88ff9576ff9576%
  ffac88ffac88ff9f7dff9576ff9576ff9f7dffac88ff7367ffac88ff7869ff7e6cff6f65ff8c72fc6260fe6862ff9576%
  ff6b63ff8c72ff7367ff7869ff7869ff6f65ff7e6cff7869ff7367ff7869ff8c72ff7e6cff7e6cd52349ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffac88ffac88ffffffffffffffffffffffffffac88ffac88ffac88ffac88ffac88ff9f7dff9576ffffffff9576%
  ff9f7dffac88ffac88ff9f7dff9f7dff9576ff9576ff9576ff9f7dffac88ff8c72ff7e6cff6f65ff8c72ff7869ff7869%
  ff7869ff7e6cff6b63ff8c72ff7869ff846eff7e6cff7869ff7869ff846eff7367ff8c72ff7e6cff7367ff6f65ff7869%
  ff6f65ff9576d32148ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffac88ffac88ffac88ffffffffffffffffffffac88ff9f7dffffffffac88ffffff%
  ff8c72ffac88ff9576ff9576ff9f7dffffffffac88ff9f7dffac88ff9576ff9f7dff9f7dff9576ffac88ff8c72ff8c72%
  ff8c72ff7869ff7367fd6561ff6f65fe6862ff7869ff7e6cff846eff6b63ff7869ff846eff7869ff6f65ff7e6cff6f65%
  ff6b63ff9576ff7869fe6862ff7367ff7869ff846ed42248ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffff%
  ffffffffffffffffffffffffff9576ffffffff9576ffffffffffffffac88ffffffffffffff9f7dffac88ff9576ff846e%
  ff9f7dffac88ff7367ff846eff7869ff8c72ff8c72ff7e6cff7e6cfd6561ff6f65ff6b63ff9576ff7e6cff9576ff7e6c%
  ff6f65ff7869ff7e6cff6b63ff7869ff846eff6b63ff846eff7869ff7869ff846eff8c72d42248ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffac88ff9576ffac88ff9f7dffac88ffac88ffac88%
  ffffffff9f7dffac88ffac88ffac88ffac88ff8c72ff7367ff846eff6f65ff6b63ff7e6cff8c72ff7e6cff7e6cfb605f%
  ff6b63ff7367ff9576ff7367ff7869ff9576ff6f65fd6561ff846eff9576ff8c72ff9576ff7367ff7e6cff7367ff7367%
  fe6862d11e46ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffac88ff9576ffffffff846e%
  ffffffff9f7dffffffffac88ffac88ffac88ffac88ff9f7dffac88ff9f7dff8c72ff9f7dff846eff9576ff7367ff8c72%
  ff8c72fe6862ff8c72fc6260ff6b63ffac88ff7869ff7367ff8c72ff846eff6f65ff7e6cff8c72ff7367ff6b63ff6f65%
  ff9576ff8c72ff7869ff846eff846eff7e6cd32047ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffac88ffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffac88ffac88ffffffff9f7d%
  ff9576ffffffffffffff9576ff9f7dff9f7dff8c72ff9576ff9f7dffac88ff8c72ff8c72ff9576ff846effac88ffac88%
  ff9576ff846eff8c72ff6f65fd6561fe6862ff6b63ff846eff6f65fe6862ff6b63ff7e6cfb605fff9576ff7869ff9576%
  ff7367ff7367ff846eff7869ff7367f85a5dff846eff8c72ff7869fe6862ff846ed7254affffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffffff%
  ffac88ffffffffffffffac88ffac88ffffffff9576ff9f7dffffffffac88ffac88ff9f7dff9f7dff9576ffffffffac88%
  ff9576ff9576ffac88ff9576ff846eff6f65ff846eff7367ff7367ff9576ff8c72fe6862fd6561ff6f65fb605fff8c72%
  ff7367ff7367ff7869ff846eff7869ff7e6cff7e6cfe6862ff846eff7367ff846eff8c72ff8c72ff7e6cff846efd6561%
  da294bffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffff9f7dffffffffac88ffffffffac88ff9f7dff9f7dffffffffffffffac88ff8c72ff9576ff9f7dff9576ff9576%
  ffac88ff9f7dff8c72ff8c72ff9f7dff9f7dff7869ff846eff7869ff7e6cff6f65ff846eff846eff6f65ff7367ff6f65%
  ff846efd6561fe6862fc6260ff7367ff7367ff7e6cfe6862ff846eff9f7dff6b63ff7367ff7869ff7869ff6b63ff6f65%
  ff6b63ff6f65ff6f65ff7367ff6f65da2a4cffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffac88ffac88ffffffffac88ff8c72ffffffff846e%
  ff8c72ff8c72ffffffff8c72ff8c72ff846eff9576ff8c72ff6b63ff846eff9576ff8c72fe6862ff846eff7e6cff6f65%
  ff7367ff7869ff7367ff6f65ff7367ff6b63fe6862f95c5efd6561ff9576ff6b63ff846eff6f65ff7e6cff7e6cff846e%
  ff6f65ff7e6cff9f7dff6b63fe6862ff7367ff7367ff6f65ff7367ff7e6cd8274affffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffffffff9f7dffffff%
  ffac88ffac88ff9f7dffac88ffffffff9576ff7e6cff8c72ff9f7dff9576ffac88ff846eff9f7dffac88ff9f7dff9f7d%
  ff7869ff7e6cfd6561ff846eff7869fe6862ff6f65f85a5dff7869ff9576ff846efe6862fd6561ff7367fd6561fe6862%
  ff7367ff6f65fb605ffd6561ff6b63ff6f65ff8c72ff8c72fd6561ff6f65ff7367ff846eff8c72ff846eff7367d8264a%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffffff9f7d%
  ffac88ffffffffac88ffac88ffffffffac88ffffffff7e6cff8c72ff846eff8c72ff9f7dff8c72ff9f7dff8c72ff9f7d%
  ff6f65ff846eff6f65fa5e5fff846eff846eff7e6cff8c72ff7869ff9576ff7367ff7869fc6260ff6b63fc6260fe6862%
  ff6b63fe6862ff846efd6561ff7367fc6260f85a5dff6f65fe6862ff8c72fe6862fe6862ff6b63fe6862ff7869ff846e%
  ff8c72ff846eff8c72ff7e6cde2e4effffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88%
  ffac88ffffffffffffffac88ffffffffac88ffac88ff9f7dff9576ff9f7dff9f7dff846eff9f7dff9576ff9576ff9f7d%
  ff9576ff846eff9576ff8c72ff7e6cff6f65ff7869ff6b63ff7367ff7367ff7869ff7869ff6f65ff6f65fd6561ff9576%
  ff6f65ff7367fe6862ff6b63f7565cff7367ff846ef95c5efb605fff6b63fe6862ff7e6cff6b63ff6f65fa5e5ffd6561%
  fa5e5fff846eff6b63fd6561ff6f65fe6862ff7869ff7869ff846ede2e4effffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ffffffffac88ff8c72%
  ff9f7dffffffff7869ff846effac88ff846eff8c72ff8c72ff8c72ff8c72ff7e6cff7869ff6f65ff7367ff7869ff7e6c%
  ff9f7dff7e6cff7e6cff7869ff7869ff9576ff6f65fd6561fe6862fe6862ff7367fc6260ff7869fc6260ff6b63ff6f65%
  ff6f65ff6b63fe6862fd6561fc6260fa5e5ffe6862fb605fff846eff9576ff7367ff7869ff6f65ff7e6cdf2f4effffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ff9f7dffffffff9f7dffac88ffffffffffffffffffffac88ffffff%
  ffffffffffffff9f7dff9f7dff7869ff9f7dff9f7dff8c72ff846eff7869ff7869ff9f7dff846eff6f65ff8c72fe6862%
  ff7e6cfd6561ff6b63ff6b63fd6561ff6b63fe6862ff6f65f7585dfc6260ff6f65fe6862ff7869f95c5efd6561ff846e%
  fe6862fb605ffc6260f6545cfc6260ff7367ff6b63fd6561ff7869fa5e5fff7869fd6561fc6260ff7367ff6f65ff6f65%
  ff7869fe6862ff7869e23350ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffffffffffffffac88ff9f7d%
  ffac88ffac88ffac88ffac88ffac88ff9f7dff9576ff9f7dffac88ffac88ff846eff9576fe6862ff6b63ff7869ff8c72%
  ff7367fb605ffd6561ff6f65fd6561fb605fff7869fc6260ff7367ff7367fe6862fe6862f95c5eff7e6cff7367f5535b%
  fb605fff6b63ff7367f7585dfe6862ff6b63fc6260fe6862fb605ff7565cf85a5dfe6862f7585dff7e6cff6b63ff7367%
  ff6b63ff846eff6b63ff6f65ff6b63ff7e6cff8c72ff846ee63852ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9576ff9576ff9f7dff9f7dffffffff9576ff8c72ffac88%
  ff9576ff9f7dff9f7dff9f7dff9576ffffffff9f7dff9576ffffffffac88ff9f7dff9576ff9f7dffffffffffffff9f7d%
  ffac88ffffffff8c72ffffffffffffff9576ffac88ff9f7dff9576ffac88ff9576ff8c72ff846eff9f7dff7e6cff7367%
  ff9576ff9576ff846efc6260f95c5efd6561ff6b63f95c5eff7367ff7e6cff7367ff6b63fc6260fd6561fa5e5fff6b63%
  ff6f65fb605ffb605ff7565cfc6260f7565cf7565cf14958f7565cfc6260fb605ff34e5af5515bfa5e5ff5515bfb605f%
  fc6260fd6561fe6862ff6f65ff8c72fa5e5ffe6862ff8c72ff6b63ff7367fe6862ff846eff7869e73a52ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88fa5e5ffd6561fd6561fc6260%
  f5515bfa5e5ffc6260f7585dff6b63fc6260fc6260ff9576ff6f65ff6f65ff7367ff846eff9576ff846eff846effac88%
  ffac88ff9f7dff8c72ff7e6cffac88ff9576ffac88ff9f7dff846eff846eff9576ffac88ff7367ff7e6cff846eff9f7d%
  ff846eff7869ff8c72fe6862ff8c72ff6f65ff6f65fa5e5ffe6862fb605ff34d59f85a5df7585df7565cff6f65fd6561%
  fd6561fc6260fb605ff95c5efb605ffe6862f7565cf95c5efb605ffa5e5ff5535bf85a5df5515bf5535bfe6862fd6561%
  f95c5ef5515bf14958fc6260f95c5eff9576fb605ffa5e5ffd6561ff7e6cfb605fff6b63ff7367ff7367ff6f65ff6f65%
  fe6862ff7367ea3d54ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  fc6260ec3f55f14858f24b59f34e5ae33550f14858f34d59f95c5ef4505af5535bfe6862ff6b63ff7367ff6b63ff6f65%
  ff6f65fa5e5fff7367ff6b63ff7e6cff6f65ff7869ff7869ff6f65ff8c72ff846eff8c72ff7e6cff8c72ff6f65ff846e%
  ff9576ff846eff7e6cff6b63ff9576ff6f65ff7869fd6561fe6862ff7e6cfb605ffe6862ff7367fa5e5ff95c5ef5515b%
  f7585dfb605ffd6561f85a5dff7869fa5e5ffd6561ff6b63fb605ff7585dfc6260fb605ffc6260fa5e5ff6545cf85a5d%
  f6545cf14958f4505af7565cfb605ff4505af5535bfa5e5ff95c5efb605ffb605ff5535bfb605fff846eff7e6cff7869%
  f85a5dfc6260ff8c72ff6b63ffffffff7869ff9576e63852ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88f5515be73952e2324fe73952ea3d54ea3d54e73952ea3d54f34d59f34d59f5535b%
  f7585df7585df5535bf7565cff6f65fa5e5ff85a5dfc6260fd6561ff7367fb605fff7367ff6b63ff6f65ff8c72ff8c72%
  ff7869ff6f65ff7869ff7367fb605fff846eff8c72ff7367fc6260fc6260fa5e5ffb605fff6f65f95c5eff7367fd6561%
  f7585dfa5e5ff4505af34d59fe6862f34e5af7565cf95c5ef34e5afc6260f5515bff6b63fc6260f7585df5535bf7585d%
  f7565cf85a5df85a5df7565cf7565cf5535bf85a5dfd6561fe6862fc6260f6545cfa5e5ff7565cff6b63fa5e5ffd6561%
  fc6260fa5e5fff6b63ff6b63ff6f65ff6b63ff7e6cff7367ff7e6cff6b63ff846effac88e53651ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff7e6cf5535bef4356e33450ee4256f04757ee4156f34d59%
  f7565cf95c5ef85a5dfd6561fc6260f95c5eff6f65ff7869ff6f65ff6f65fc6260ff6f65ff846eff6b63ff7e6cff6b63%
  fe6862ff6f65ff6b63ff7367ff7367fe6862ff6f65fa5e5fff6b63fe6862ff6b63fa5e5fff7367f5535bfc6260f7585d%
  fd6561f85a5df7585df14858f7565cf5535bfc6260f34e5af85a5df7565cf5535bf7565cf95c5eff7869fd6561fe6862%
  f7585dfa5e5ff7585dfb605ff6545cf24b59f95c5efc6260fc6260f5515bfa5e5ff7565cfb605fff6b63fc6260fb605f%
  fa5e5ff7585df85a5dff6f65ff6b63ff7869fe6862fd6561fd6561ff8c72ff846efd6561ff7869ff7869ff9576ff8c72%
  ff9f7deb3e54ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88f04557e33550e33550%
  ec4055ec3f55ef4457f34d59f6545cfa5e5ffe6862f34d59f7585dfb605fff6f65ff7869ff7367fc6260fa5e5ffc6260%
  fe6862ff846eff6f65ff6b63f95c5ef34d59ff846ef95c5efd6561f7565cf7565cf85a5dfc6260f34e5afc6260f7585d%
  f95c5ef5535bf4505af95c5ef6545cf4505af14858f34d59f7565cf34d59f7585df7585df7585df24b59fa5e5ff7585d%
  fa5e5ff6545cf7585df6545cf95c5efa5e5ff7585dfb605ff34e5af7585df7565cf85a5df4505af24a59fc6260f5535b%
  f6545cf4505af7585dff6b63f7585df95c5ef7585dfd6561fd6561ff6b63ff6b63fd6561fc6260fe6862ff7e6cfe6862%
  ff6b63ff6b63ff846eff846eff6f65ff7869e83b53ffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ff9576f95c5ee1324fe0314fe23450e33550e43651ea3d54ef4457f04757f24a59f7585dfb605ffa5e5ffc6260f95c5e%
  fa5e5fff6b63fa5e5ff7585dff7367fd6561fb605ff7565cfe6862fa5e5ffb605ff7585df95c5ef7585df7565cf34e5a%
  f04757f34d59f34e5af4505af04757f5535bf24a59f34d59f14858f34d59f04757f14958f24b59f04757f14958f4505a%
  f24a59f04557f7585def4457f14958fc6260f5535bf34e5af85a5df7565cf04757f34e5af5535bf7565cf6545cf85a5d%
  ff6b63f4505af7565cf95c5ef7585dfe6862fd6561fa5e5ffa5e5ff5535bf34e5afa5e5ffc6260fe6862fa5e5fff6b63%
  fd6561ff6f65fe6862ff7e6cff6f65ff7869ff7869fe6862ff8c72ff6f65ff8c72e73a52ffffffffffffffffffffffff%
  fffffffffffffffffffffffffe6862e43551df2f4ecc1b42d9284bdd2d4dde2e4ee43551e33450ec4055f14858f5515b%
  f4505af5515bff6b63f4505afb605ff6545cf85a5df5535bfb605ffb605ff6545cfc6260f85a5dfb605ff24a59f6545c%
  f85a5df5515bf5535bf5535bf14858ef4457ec4055f4505aeb3e54f34e5aef4356ef4457f24b59ee4156ee4256e83a53%
  e63852ec3f55ee4156f14958e73a52e83b53e83a53ec3f55ea3d54e53751ef4356f14958ef4356f5515bed4155f95c5e%
  f34d59f14958f6545cee4156f5515bf7565cf6545cf04557f5535bfb605ff34d59f7565cf34e5af6545cf24a59fe6862%
  fe6862f85a5dfa5e5ff85a5dfc6260fc6260f85a5dff6f65ff7e6cff7e6cff6f65ff9576ff846eff7e6cff8c72ff6b63%
  ec4055fffffffffffffffffffffffffffffffffffffffffffe6862dc2c4dd21f47cb1a42c8183fdb2a4ce23450e43651%
  ed4155ee4256f24b59f4505afc6260f5515bf4505af5515bf85a5df6545cfd6561f5515bf7565cf85a5df85a5df85a5d%
  ff7e6cf5515bf34e5aff6b63fd6561fa5e5ff14958f95c5ef24a59f24b59f04757f14858f04557f24a59f34d59f04757%
  f4505af04557f04757e93c53f04757e33550e63852e43551ea3d54ea3d54ec3f55e93c53e63852ea3d54ec4055ee4156%
  ea3d54f14958f34d59f6545cf7585df34d59f34e5af14958f14958f6545cf04557f6545cf04557fa5e5ff7565cf95c5e%
  f95c5efd6561fa5e5fff846efc6260ff6b63ff6f65ff7367ff6f65ff6f65ff6b63ff6f65ff7367ff7869ff9576ff846e%
  ff7869ff7367ff7869ff9576ff846ef24a59ffffffffffffffffffffffffffffffffffffffffffde2d4dd01e45c4153c%
  cb1a42ce1c44de2e4ee0314fea3d54f34d59f5515bf5535bf34e5af6545cf34e5afb605ffb605ffa5e5ffc6260fa5e5f%
  fe6862fc6260fa5e5ff5535bfb605ff95c5ef6545cf7565cfc6260f24a59f34e5af24a59f04757ef4457f34d59f14858%
  f04557f14958f4505af34e5af14858e83a53f34d59eb3e54ec3f55ef4457ea3d54ee4156ee4156ea3d54e63852ef4356%
  e83a53ef4356e83b53e73a52f14958ee4156f14958ef4457f14958f14858f5515bf14958f5515bf5515bf4505afa5e5f%
  fc6260ff7869f4505afd6561fa5e5fff6f65f85a5dff8c72ff6f65ff7869ff9f7dff7869ff8c72ff9576ff846eff7e6c%
  ff9f7dff9576ffac88ff846effac88ff9f7dffac88ffac88ffac88ffac88ff9f7dffffffffffffffffffffffffffffff%
  ffffffff9576c4153cc11239c3143bd11e46d52349e0304edf304ee63852f14958f14858f34e5afb605ff34d59f34e5a%
  f34e5af24a59f34e5af24a59ee4256f5515bf34e5aef4457f4505af7565cf24a59f34d59f24a59f04557e53751ea3d54%
  e83b53e63852ee4256ea3d54ef4356e93c53ec3f55e33550f14858f04757ef4356ea3d54eb3e54e93c53e63852f04557%
  e73a52ea3d54e83b53eb3e54ee4156ed4155e83a53ed4155f04757f14958ee4156f14958f5515bfd6561f5515bf6545c%
  f4505afd6561f85a5df85a5dfc6260ff6b63ff7869fe6862ff7e6cff846eff8c72ff846effac88ff9576ffac88ff8c72%
  ff8c72ffac88ff9f7dffac88ffffffffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffff4505ac8183fcb1a42e93c53dd2d4ded4155ef4457f04757f14958f14858%
  f34d59f24a59fb605ff34d59f14958f85a5df04557f34d59f14858f24a59f7565cf85a5dfd6561f24b59ef4356f5535b%
  f4505af14958e83b53f04757f04557ec3f55f34d59f14958f24a59ed4155ec3f55f04757f24b59f14958f34d59f34d59%
  e43651ef4457ee4256f14858f24a59f04757fa5e5ff04557ef4457f14858ef4356f5535bf14958f7585df85a5df95c5e%
  f85a5dfc6260f7585dff7367ff7e6cff7869ff7367ff7367ff846eff8c72ffac88ff9576ffac88ffffffffffffffac88%
  ffac88ffac88ff9f7dffffffff9576ff9576ffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9576e83b53d42248e1324ff14958f24b59%
  f24b59f4505af7565cf7565cf95c5ef95c5eff6b63fa5e5ffa5e5ff5535bf4505af34d59f34d59f6545cf85a5dfc6260%
  fe6862f7585df6545cf5535bf34e5af4505af6545cf7565cf95c5ef85a5dfa5e5fff8c72fa5e5ff14858fa5e5ff14858%
  f7585df34e5af34d59f34d59f34e5af4505af4505af4505af6545cf95c5ef34d59f7585df34e5af24b59fe6862fe6862%
  f6545cf7565cf85a5dff7367f95c5eff7869ff6b63ff7e6cff7869ff7869ff846effac88ff7e6cff9576ff9f7dffffff%
  ff846eff846eff9576ffac88ffffffffac88ff9576ff9f7dffffffffac88ffffffffffffffffffffffffffac88ffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff4505a%
  d21f47d52248db2a4ce33550e23350e73952f24a59ee4256e73952f04757ef4457ee4156ef4457ef4356eb3e54f14858%
  ef4457f34e5af85a5df4505afb605ffa5e5feb3e54e83a53f04757f04757f04557ea3d54f4505afb605ff4505aef4356%
  f24a59f14858f14858f14858ee4156e83a53e83a53ee4256e73a52e83a53e93c53e83a53e83b53f04557f24b59ef4457%
  e93c53ec3f55eb3e54ed4155f34d59ef4457fc6260f24b59f5515bf5515bf95c5efb605ffa5e5ffe6862ff8c72ff7e6c%
  ff846eff7869ff9576ffffffffac88ff8c72ff9576ff9576ffac88ffac88ff9f7dffac88ffffffffffffffac88ff9f7d%
  ffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff8c72dd2d4dcb1a42e2324ff14858ef4356f6545cf5535bf4505af5535bf5515bfb605ff34d59%
  f34d59f4505af5535bf7585df5515bfa5e5fff7367f95c5efc6260f7565cfb605ff24a59ef4356ef4356f14858f5535b%
  f24b59f7565cf6545cfd6561f5515bf24a59ee4156f4505af24b59ef4356ef4356ef4356f34d59ea3d54ee4256f04757%
  f14858ef4356f04557ef4356f7565cf34e5af24b59f95c5ef5515bf95c5ef5515bfa5e5ffc6260ff7e6cff7e6cff846e%
  ff7869ff9576ff846eff7869ff846eff9576ff8c72ff9f7dffac88ffac88ffac88ffac88ffffffffac88ffffffffffff%
  ffffffffac88ffffffffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffff34d59e1324ff04557fd6561ff7869fd6561ff9f7dff7869%
  ff9576fc6260ff8c72fc6260ff6f65ff6f65fd6561ff846eff7e6cfb605fff6f65ff7367ff8c72ff8c72fa5e5ffc6260%
  fd6561ff6b63ff6f65fa5e5ffe6862fb605ffd6561ff6f65ff7e6cfb605ffd6561ff6b63fc6260ff7e6cfb605ff5515b%
  f24b59ff7e6cfd6561ff6f65fd6561f95c5efc6260fd6561fd6561ff6f65ff7e6cff6f65ff9576ff846eff6b63ff9f7d%
  ff9f7dffac88ffac88ff9f7dff9f7dff9576ff9f7dffffffffac88ffac88ffffffff9f7dffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffac88e53751bc0f36c4153cda294b%
  ea3d54ec3f55f6545cf24a59f7565cf34e5af6545cf24a59f14958f34d59f34e5af34d59f34e5af04757f4505af04757%
  f24a59f04557eb3e54ec4055e83a53f34d59ef4457ef4457ef4356ef4356ee4156ec4055ea3d54ee4156f34d59ee4156%
  f24a59f14858ed4155f14858f14958f04557ef4457f4505af34d59f4505af7585df5535bfb605ff85a5dff6f65fd6561%
  f85a5dff6b63ff8c72ff9f7dff846effac88ff846effac88ff846eff9576ffffffffffffffffffffac88ffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7d%
  ff9576d7254ac7173edd2d4dee4256f5515bf5515bfc6260f4505af24a59f34d59f5535bf6545cf24a59f34e5af4505a%
  f04557f04757f34d59f34d59ee4256f04557f14958f5535bf7565cf14958f14858f6545cef4457ef4457f04557f24a59%
  f14958f4505aec4055ef4457f34e5af95c5eef4356f14858fd6561fa5e5ff24b59f85a5df95c5efc6260f85a5dfa5e5f%
  ff846eff8c72ff846eff6b63ff846eff8c72ff9f7dffac88ffffffff9f7dffffffffac88ff9f7dffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff8c72ff6f65fe6862d7254ad01e46ef4457f24b59f04757f04557f4505ae83b53ef4457f34d59%
  f34d59f34e5af04757f24b59f4505af5515bef4356f04757f34e5aef4457ef4457ec3f55f04557e53751ee4156f24b59%
  f24b59ef4356f14958f14858e93c53ea3d54ea3d54eb3e54f34e5af04557f4505af24a59f4505af7585dff6f65fc6260%
  fd6561fe6862ff7367ff8c72ff7869ff7367ff8c72ff8c72ff9f7dffffffff9f7dff9f7dff9f7dffac88ffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffff7585df34e5ae0304ed11e46da294be2324fe63852e23450%
  e23350e73a52e43651ea3d54ea3d54e73952e93c53ef4457ef4457f14858ef4356eb3e54e73952e83a53ea3d54e2324f%
  e73a52e73a52ed4155ee4156ea3d54ef4356ec4055f34d59f24b59e83b53ee4156ef4356ee4156f34e5af34d59f5515b%
  f7565cff7367f85a5dff9576ff7e6cff846eff9576ff8c72ff8c72ff9576ff8c72ffffffffffffffffffffac88ffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  fffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff24a59db2a4cc2133a%
  c01138cd1c43d21f47d62449d62449d8274ad42248d42148d7264ada294bdb2a4cdc2c4ddc2c4cd9284bde2d4dde2e4d%
  e1314fde2d4de23450e0314fda294bdc2c4de1314fdb2b4cda294be23450d9274be23450e0314fe33550e73a52eb3e54%
  ee4156f14858f24b59f4505afe6862f34e5aff7e6cff7367ff9f7dff7869ff9576ff9576ffac88ff8c72ff9f7dffffff%
  ffac88ffffffffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffcc1b42c1133aba0e34ca1940d32047d52349d11e46d01e45d62449%
  d7254ad62449d52349da294bd7254add2d4dda2a4cd8264ad32148d7254ad62449da294bdc2b4cd7254ad8264ad7264a%
  db2a4cd62449d42248d52349e0304ee43551e1314fe83b53e83b53ea3d54f4505af95c5eff7869ff7367ff6f65ff7e6c%
  ff7367ff7869ff9576ffac88ff9f7dffac88ffac88ffac88ff9576ffac88ffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffff9a0212af062bb80c32c01138cb1a42cf1d45d32047d52349d21f47d62449d8264ad9284bd42148d7254ad22047%
  d01e46d21f47d32047d01e46d11f46d9274bd32047d8264ad9284bd7264ad8264adc2c4cdc2b4ce1314fe73a52e63852%
  ea3d54f24b59f34d59fb605fff6f65ff7869ff7869ff8c72ff9576ff9576ffac88ffffffff9f7dffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff8b0000b70b32b50a30b60b31b70b32%
  bb0e35bc0f36bf1138c01239bf1138bf1138ba0d34bf1138c1133ac1133ac91840c9183fca1941c6163dca1941c91840%
  cd1b43d11e46d01d45d52248dc2c4cda2a4ce53751e73a52ec3f55f34e5afa5e5fff6b63ff7e6cff846effac88ff9f7d%
  ff9576ff9f7dffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff8b0000%
  b3082eae062aac0527aa0525ac0527b0062cb4092fba0e34bd1037c5163dc7173eca1940d21f47d7264ade2e4deb3e54%
  ed4155fb605ff5515bff6b63ff9576ff7367ff9f7dff8c72ffac88ffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffff8b0000ff6b63ff9576ffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffff}%
\gprawimage{rgb}{1.302}{2.179}{101}{101}{3.624}{3.624}{\gprawrgbimagedata}{}
\end{scope}
\gpcolor{rgb color={0.000,0.000,0.000}}
\draw[gp path] (1.320,2.197)--(1.356,2.233)--(1.392,2.269)--(1.429,2.306)--(1.465,2.342)%
  --(1.501,2.378)--(1.537,2.414)--(1.574,2.451)--(1.610,2.487)--(1.646,2.523)--(1.682,2.559)%
  --(1.719,2.596)--(1.755,2.632)--(1.791,2.668)--(1.827,2.704)--(1.864,2.741)--(1.900,2.777)%
  --(1.936,2.813)--(1.972,2.849)--(2.009,2.886)--(2.045,2.922)--(2.081,2.958)--(2.117,2.994)%
  --(2.154,3.031)--(2.190,3.067)--(2.226,3.103)--(2.262,3.139)--(2.299,3.176)--(2.335,3.212)%
  --(2.371,3.248)--(2.407,3.284)--(2.444,3.321)--(2.480,3.357)--(2.516,3.393)--(2.552,3.429)%
  --(2.588,3.465)--(2.625,3.502)--(2.661,3.538)--(2.697,3.574)--(2.733,3.610)--(2.770,3.647)%
  --(2.806,3.683)--(2.842,3.719)--(2.878,3.755)--(2.915,3.792)--(2.951,3.828)--(2.987,3.864)%
  --(3.023,3.900)--(3.060,3.937)--(3.096,3.973)--(3.132,4.009)--(3.168,4.045)--(3.205,4.082)%
  --(3.241,4.118)--(3.277,4.154)--(3.313,4.190)--(3.350,4.227)--(3.386,4.263)--(3.422,4.299)%
  --(3.458,4.335)--(3.495,4.372)--(3.531,4.408)--(3.567,4.444)--(3.603,4.480)--(3.640,4.517)%
  --(3.676,4.553)--(3.712,4.589)--(3.748,4.625)--(3.784,4.661)--(3.821,4.698)--(3.857,4.734)%
  --(3.893,4.770)--(3.929,4.806)--(3.966,4.843)--(4.002,4.879)--(4.038,4.915)--(4.074,4.951)%
  --(4.111,4.988)--(4.147,5.024)--(4.183,5.060)--(4.219,5.096)--(4.256,5.133)--(4.292,5.169)%
  --(4.328,5.205)--(4.364,5.241)--(4.401,5.278)--(4.437,5.314)--(4.473,5.350)--(4.509,5.386)%
  --(4.546,5.423)--(4.582,5.459)--(4.618,5.495)--(4.654,5.531)--(4.691,5.568)--(4.727,5.604)%
  --(4.763,5.640)--(4.799,5.676)--(4.836,5.713)--(4.872,5.749)--(4.908,5.785);
\gpcolor{color=gp lt color border}
\draw[gp path] (1.320,2.197)--(1.500,2.197);
\node[gp node right] at (1.136,2.197) {1};
\draw[gp path] (1.320,2.744)--(1.500,2.744);
\node[gp node right] at (1.136,2.744) {10};
\draw[gp path] (1.320,5.481)--(1.500,5.481);
\draw[gp path] (1.320,5.785)--(1.500,5.785);
\node[gp node right] at (1.136,5.785) {1h};
\draw[gp path] (1.320,2.197)--(1.500,2.197);
\draw[gp path] (1.320,2.362)--(1.410,2.362);
\draw[gp path] (1.320,2.580)--(1.410,2.580);
\draw[gp path] (1.320,2.691)--(1.410,2.691);
\draw[gp path] (1.320,2.744)--(1.500,2.744);
\draw[gp path] (1.320,2.909)--(1.410,2.909);
\draw[gp path] (1.320,3.127)--(1.410,3.127);
\draw[gp path] (1.320,3.238)--(1.410,3.238);
\draw[gp path] (1.320,3.292)--(1.500,3.292);
\node[gp node right] at (1.136,3.292) {$10^{2}$};
\draw[gp path] (1.320,3.456)--(1.410,3.456);
\draw[gp path] (1.320,3.674)--(1.410,3.674);
\draw[gp path] (1.320,3.786)--(1.410,3.786);
\draw[gp path] (1.320,3.839)--(1.500,3.839);
\node[gp node right] at (1.136,3.839) {$10^{3}$};
\draw[gp path] (1.320,4.004)--(1.410,4.004);
\draw[gp path] (1.320,4.221)--(1.410,4.221);
\draw[gp path] (1.320,4.333)--(1.410,4.333);
\draw[gp path] (1.320,4.386)--(1.500,4.386);
\node[gp node right] at (1.136,4.386) {$10^{4}$};
\draw[gp path] (1.320,4.551)--(1.410,4.551);
\draw[gp path] (1.320,4.769)--(1.410,4.769);
\draw[gp path] (1.320,4.880)--(1.410,4.880);
\draw[gp path] (1.320,4.933)--(1.500,4.933);
\node[gp node right] at (1.136,4.933) {$10^{5}$};
\draw[gp path] (1.320,5.098)--(1.410,5.098);
\draw[gp path] (1.320,5.316)--(1.410,5.316);
\draw[gp path] (1.320,5.428)--(1.410,5.428);
\draw[gp path] (1.320,5.481)--(1.500,5.481);
\draw[gp path] (1.320,5.645)--(1.410,5.645);
\draw[gp path] (1.320,2.197)--(1.320,2.377);
\node[gp node center] at (1.320,1.889) {1};
\draw[gp path] (1.867,2.197)--(1.867,2.377);
\node[gp node center] at (1.867,1.889) {10};
\draw[gp path] (4.604,2.197)--(4.604,2.377);
\draw[gp path] (4.908,2.197)--(4.908,2.377);
\node[gp node center] at (4.908,1.889) {1h};
\draw[gp path] (1.320,2.197)--(1.320,2.377);
\draw[gp path] (1.485,2.197)--(1.485,2.287);
\draw[gp path] (1.703,2.197)--(1.703,2.287);
\draw[gp path] (1.814,2.197)--(1.814,2.287);
\draw[gp path] (1.867,2.197)--(1.867,2.377);
\draw[gp path] (2.032,2.197)--(2.032,2.287);
\draw[gp path] (2.250,2.197)--(2.250,2.287);
\draw[gp path] (2.361,2.197)--(2.361,2.287);
\draw[gp path] (2.415,2.197)--(2.415,2.377);
\node[gp node center] at (2.415,1.889) {$10^{2}$};
\draw[gp path] (2.579,2.197)--(2.579,2.287);
\draw[gp path] (2.797,2.197)--(2.797,2.287);
\draw[gp path] (2.909,2.197)--(2.909,2.287);
\draw[gp path] (2.962,2.197)--(2.962,2.377);
\node[gp node center] at (2.962,1.889) {$10^{3}$};
\draw[gp path] (3.127,2.197)--(3.127,2.287);
\draw[gp path] (3.344,2.197)--(3.344,2.287);
\draw[gp path] (3.456,2.197)--(3.456,2.287);
\draw[gp path] (3.509,2.197)--(3.509,2.377);
\node[gp node center] at (3.509,1.889) {$10^{4}$};
\draw[gp path] (3.674,2.197)--(3.674,2.287);
\draw[gp path] (3.892,2.197)--(3.892,2.287);
\draw[gp path] (4.003,2.197)--(4.003,2.287);
\draw[gp path] (4.056,2.197)--(4.056,2.377);
\node[gp node center] at (4.056,1.889) {$10^{5}$};
\draw[gp path] (4.221,2.197)--(4.221,2.287);
\draw[gp path] (4.439,2.197)--(4.439,2.287);
\draw[gp path] (4.551,2.197)--(4.551,2.287);
\draw[gp path] (4.604,2.197)--(4.604,2.377);
\draw[gp path] (4.768,2.197)--(4.768,2.287);
\draw[gp path] (1.320,5.785)--(1.320,2.197)--(4.908,2.197);
\node[gp node center] at (3.114,5.426) {33\% Labelled};
%% coordinates of the plot area
\gpdefrectangularnode{gp plot 1}{\pgfpoint{1.320cm}{2.197cm}}{\pgfpoint{4.908cm}{5.785cm}}
\draw[gp path] (6.780,5.785)--(6.780,2.197)--(10.368,2.197);
\node[gp node center,rotate=-270] at (5.706,3.991) {Clique Runtime (ms)};
\node[gp node center] at (8.574,1.427) {MAC Runtime (ms)};
\begin{scope}
\clip (6.780,5.785) rectangle (10.368,2.197);
\def\gprawrgbimagedata{%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffffffffffffffac88ffac88ffac88%
  ffac88ffac88ff8c72ffffffffac88ff9576ff9f7dff8c72ff7869ff7e6cff7367ff6f65ff8c72ff9576ff6f65ffac88%
  ff846eff846eff9576ff7e6cff7e6cff6f65ff7869fc6260fc6260ff7869fe6862fe6862fc6260f85a5dfb605ff34e5a%
  f7565cfa5e5ff5535bf5515bf04557ec4055ea3d54e83b53e93c53e23450e53651db2a4cdc2b4cd9274bdb2a4cd7254a%
  d8264ad32047d32048d11f468b0000ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffac88%
  ffac88ffffffffffffff9576ffac88ffac88ffffffffac88ffffffffffffff9f7dffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffff9576ff9f7dff9576ff8c72ff9f7dff9576ffac88ff9f7dff7e6cff9f7dff9f7dffac88%
  ffac88ff7367ff846efe6862ff8c72ff7e6cff846eff846eff7367ef4356ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffac88ffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffff9f7dff9f7dffffffffffffff9f7dffffffffffffffac88ffffffff9f7dff9f7dff846eff8c72ff9f7d%
  ff8c72ff9576ff9f7dff7869ffffffffac88ff6f65ff7869ff8c72ff9f7dff7869ff9f7dff7e6cff7869f5515bffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffff9f7dffffffffac88ffffffffffffffac88ff9f7dffffffff9f7dffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffac88ffffffff9f7dff9f7dffffffffffffffac88ffffffffffffff9f7d%
  ffac88ffac88ff9f7dff9f7dff9576ff9576ff846eff7e6cff9f7dff8c72ffffffff7869ff8c72ff7869ff7367ff9f7d%
  ff9576fa5e5fff7e6cfd6561ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffff9f7dffffffffac88ffffffffac88ffac88ffffff%
  ffffffff9f7dffffffffac88ffffffffac88ffffffff9f7dff9576ffffffffffffffffffffffffffffffffffffffac88%
  ffac88ffffffffffffff9f7dffac88ff9576ff846eff9576ff9576ff9f7dff9576ff8c72ff9576ffac88ff7e6cff6b63%
  ffac88ff846eff7e6cffffffff9f7dff6f65ff7367ff846efd6561ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffff%
  ffffffffac88ffffffffffffffffffffac88ffffffff9f7dffffffffac88ff9f7dffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffac88ff9f7dff9576ffac88ffac88ff9f7dff7e6cffac88ff9f7dff846eff8c72ff8c72%
  ff8c72ff846eff7869ff846eff8c72ff9576ff8c72ff7e6cffac88ff7367ff8c72ff7367ff7e6cfd6561ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffff9f7dffac88ff9f7dffffffffffffffffffff9f7dffac88ffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffff9f7dffac88ffac88ff9f7dffac88ff9f7dffac88ffac88ffffffff8c72ff9576ff8c72%
  ff9576ff9f7dff9576ff6b63ff846eff7e6cff8c72ff7869ffffffff7e6cff8c72ff7e6cff7e6cff8c72ff8c72ff7367%
  ff7e6cff9576ff7869ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffff9f7dffffffffffffffffffffac88ffffffffffffff9f7dffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffffff9f7dffffffff9f7dff9f7dff9f7d%
  ff9576ff9576ff7e6cffac88ff8c72ff9f7dff846eff6b63ff9f7dff7869fd6561ff7e6cff846eff7869ff9576ff9576%
  ff846eff7869ff7869ff7e6cff9f7dff9f7dff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffac88ff9f7dffffffffac88ffffffffac88ffac88ffffffff9f7dffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffac88ffac88%
  ff7e6cffac88ff9f7dff9f7dff9576ff9f7dfe6862ff8c72ff7367ff846efd6561fc6260ff8c72fc6260ff6b63ff7869%
  ff7367ff9576ff7e6cff8c72ffac88ff8c72ffac88ff7e6cff9576ff8c72ff9576ffac88ff9576ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ffffffffac88ffac88ffffffffffffffac88%
  ffac88ffffffff9f7dffffffff846effac88ff9576ffffffff8c72ffac88ff8c72ffac88fd6561ff8c72ff8c72ff7367%
  ff7367ff7869ff7367ff6f65fe6862ff6b63fd6561ff7367ff9576ff7869ff9576ff846eff846eff9f7dff9f7dff9576%
  ffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffac88ffffffffac88ff9f7dffac88ffac88ffffffffac88ffffffffffffffffffffffff%
  ffffffff9576ff9f7dffac88ff9f7dffffffff9576ffffffff9f7dff846eff9f7dff9f7dff9576ff9576ff8c72ff846e%
  ff846eff7367ff6b63ff6b63fc6260ff846efd6561ff6b63fb605ff95c5ef7585dff6b63ff7367ff7e6cff8c72ff9f7d%
  ff9f7dff9576ffffffff846eff9576ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffff9f7dffffffffac88ffffffff8c72ff9f7dffffffff9f7dff9576%
  ff846eff9f7dff8c72ffac88ff7e6cff846efe6862fd6561fe6862ff7367ff6b63f7565cf7585dff6b63fc6260fb605f%
  ff6b63ff7e6cff7e6cff8c72ff846effac88ffffffff7e6cffffffff9f7dffac88ff9576ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffac88ffffffff9f7d%
  ffac88ffffffff8c72ffac88ff9f7dff7869ff6b63ff9576ff9576ff7e6cff7e6cff6f65fe6862fd6561f7565cf95c5e%
  fa5e5ff24b59f7565cf34e5afd6561fc6260ff7869fe6862ff9f7dff846eff9f7dffac88ffac88ffffffffac88ffffff%
  ff6b63ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffffffffffffffffffffffffff9f7dffffffffffffffffffffffffffac88ffac88ff9576ffac88ffffffffffffffffff%
  ffac88ffac88ffffffff9576ffffffff9576ff9576ffffffff8c72ff9576ff846eff8c72ffffffff8c72ff7869fd6561%
  ff846eff6f65fe6862fd6561f85a5df7565cff6b63fd6561f6545cfd6561f85a5dff846eff9576ff8c72ff9f7dff9f7d%
  ffac88ff9f7dffffffffffffff9576e93c53ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffac88ffffffffffffffffffffffffffffffffac88ffffffff9f7dffac88ffffffffac88ff9f7dffac88ffac88ff9f7d%
  ffffffffffffffffffff846effac88ffffffffffffffac88ff9f7dff7367ffac88ff9f7dff846eff8c72ff8c72ff9f7d%
  ff846eff846eff9576ff7e6cff7869f95c5efa5e5ffe6862fb605ff85a5dfd6561fb605fff7e6cf85a5df95c5efd6561%
  ff7e6cff7869ff846eff9576ffac88ff9576ff9f7dff9576ffac88ff9f7de33550ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffff9576ffffffffffffff9f7d%
  ffffffffffffff9576ffffffffac88ffffffffffffffac88ffac88ffac88ff9f7dffffffff9f7dff9576ff9f7dffac88%
  ffac88ff9f7dff8c72ff8c72ff846eff7367ff7e6cff9576ff7869ff6b63fa5e5ffc6260f4505afa5e5ff5515bf5535b%
  f95c5efb605ff5515bf4505afb605ffc6260ff846eff9576ffac88ffac88ffffffffffffff9f7dff8c72ff7e6ce83a53%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffff9f7dffffffffac88ffffffffac88%
  ffffffff8c72ffac88ffac88ffffffffac88ffffffff9f7dff9f7dffffffffac88ff9576ffffffff9f7dffac88ff9576%
  ff846effac88ff9f7dff8c72ffac88ff9576ff8c72ff8c72ff7869ff8c72ff7367ff846eff7869ff7e6cff7e6cff6b63%
  ff6b63fe6862f6545cf7565cfe6862f7585df7585df5535bfe6862fc6260ff7e6cff8c72ffac88ff9f7dff9f7dff9f7d%
  ffac88ff8c72ff846efe6862f14858ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffac88ffac88ffac88ffffffffffffffffff%
  ffac88ffac88ffffffffffffff9f7dffac88ffac88ffac88ffffffffac88ff9576ffffffff9f7dffac88ffffffffac88%
  ff9f7dff9f7dffac88ffac88ff9576ff8c72ff9f7dff9f7dff9f7dff8c72ff9f7dff7e6cffffffff7367ff8c72ff7e6c%
  ff846eff7e6cfa5e5ffe6862ff7367fe6862f85a5dfd6561fb605ffb605ff5535bf5535bf7585dfb605fff846eff7869%
  ff846eff8c72ff9576ffffffff846eff9f7dff7e6cff7e6cfe6862fb605fffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffff9f7dff9f7dffac88ffffffffffffffffff%
  ffac88ffffffffac88ffffffff9576ffffffffac88ffac88ffffffffac88ffac88ff9f7dff9f7dff846eff9576ff9f7d%
  ff9f7dff8c72ff7869ff846eff8c72ff7869ff7869ff7869ff7869fd6561fd6561f34e5af7585dfd6561f4505af95c5e%
  fc6260fa5e5ffc6260ff7e6cff8c72ff846effac88ffac88ff9f7dff9f7dff9f7dff7e6cff7e6cff7367ff9576ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffac88ff9576ffac88ffac88ff9f7dffffffffac88%
  ffffffffffffff9f7dffffffff9f7dffffffffac88ffac88ffac88ffffffffac88ffac88ff8c72ff9f7dff9f7dff846e%
  ffac88ffac88ff7e6cff9f7dff9f7dff846eff9576ff9f7dff8c72ff7869ff6f65ff7869ff7e6cff7869fe6862fb605f%
  fe6862fa5e5ffb605ffd6561f85a5dfd6561ff7869ff7869ff846eff8c72ff9576ff846eff9f7dffac88ff9f7dff9f7d%
  ff9f7dff9f7dff9f7dff9576ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffff%
  ff9f7dffffffff8c72ff9576ff9f7dffffffffffffff9f7dff9f7dffac88ff9576ff9f7dffffffff9f7dffffffffac88%
  ff9f7dffac88ff9f7dff9576ff8c72ff9f7dff9f7dff8c72ff7869ff9f7dff7869ff9576ff846eff7869fa5e5fff9576%
  ff7e6cff7869ff7869fe6862fd6561fb605ffa5e5ff95c5ef7565cf7585dfa5e5fff7869fe6862ff9576ffac88ff9576%
  ffffffffac88ff9f7dffac88ffffffff9f7dffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffac88ffffffffffffffac88ffac88ff9f7dff9f7dff9576ff9f7dffffffffac88ffffffffac88ffac88%
  ff9f7dff9f7dffffffffffffff9f7dffac88ffac88ffffffff9576ffffffffac88ff7869ff9576ff9f7dffac88ff9576%
  ff7869ff7e6cff9576ff9f7dff6f65ff7869fb605fff7e6cff6f65fd6561fb605ffb605fff7e6cff7869fe6862ff6b63%
  ff846eff7869ffac88ff846eff9f7dff9f7dff9f7dffac88ffffffffac88ffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffac88ffffffff9f7dff9f7dffac88ff7e6cff8c72ffac88ff846eff846effac88%
  ffffffffac88ff9f7dffffffffac88ffffffffac88ffffffffffffffffffffac88ff9f7dff8c72ffffffff846eff9f7d%
  ff9576ff8c72ff9576ff9576ff7e6cff9576ff9576ff846eff7e6cff6f65fd6561ff6f65fe6862fc6260fe6862ff7367%
  fd6561ff6f65ff6b63ff9576ff9576ff7e6cff9576ffac88ffffffff9f7dff9f7dffac88ffac88ffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ff9576ff9f7dff9f7dff8c72%
  ffac88ff7869ffac88ff9f7dffac88ffffffffac88ffac88ffffffffac88ffffffffffffffac88ffac88ff9f7dff846e%
  ffac88ffac88ff9576ff8c72ff9576ffffffff9f7dff9f7dff846eff9f7dff8c72ff9f7dff8c72ff8c72ff846eff846e%
  ff7869fc6260fa5e5fff6b63fd6561ff6f65ff7367ff7e6cff846effac88ff846effac88ff8c72ffac88ffac88ff9576%
  ffac88ffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffffff%
  ff9f7dff9f7dff8c72ff9f7dffac88ff9576ff8c72ff8c72ff7e6cff9f7dff9576ff9f7dffffffffac88ffac88ff9f7d%
  ffffffffffffff9f7dffffffffffffffffffff7869ffac88ff7e6cff9f7dffffffff9f7dff7e6cff9576ff8c72ffac88%
  ff8c72ff8c72ff6f65ff9576ff7367ff6f65ff7367ff7869fc6260ff7e6cff7367ff6f65ff7e6cff7e6cff7e6cff8c72%
  ff9f7dff9f7dffffffffac88ffffffff9576ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffac88%
  ffffffff9f7dffac88ffac88ffac88ffac88ff9f7dff8c72ff9f7dff8c72ff7e6cff8c72ffac88ffac88ff9576ffac88%
  ffac88ff9f7dffac88ff9f7dffffffffac88ffac88ff9576ff9f7dff8c72ff8c72ff9f7dffffffff8c72ff9f7dffffff%
  ff846eff8c72ff7367ff8c72ff9f7dff9576ff7e6cff7367ff9576ff6f65fd6561ff9576ff6f65ff6f65ff7367fe6862%
  ff7e6cff8c72ff8c72ff9576ff9576ff8c72ffac88ffac88ff9f7dff9f7dffffffffffffffffffffac88ffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffff9f7dffffffff9f7dffffffffffffff8c72ffac88ff7e6cff8c72ff8c72ff9576ff846e%
  ffac88ff846effffffffffffff8c72ffffffffffffffac88ffffffff9576ffffffffffffffac88ffac88ff8c72ff9576%
  ff9f7dff9f7dffac88ff8c72ff9f7dff7e6cff7869ff9f7dff9f7dff8c72ff846eff9576ff7367ffac88ff7367ff7367%
  fe6862ff7367ff6b63fe6862ff6f65ff6f65ff9f7dff9f7dff846eff9576ff9f7dff9f7dff9f7dff9f7dffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffac88ffac88ff8c72ffffffff7869%
  ff846eff8c72ff9576ff8c72ff9576ffffffff9f7dffffffff9f7dffac88ffffffffffffffffffffac88ff9f7dffac88%
  ff9f7dff9f7dff9f7dff8c72ff9576ff9f7dff9576ff9576ff8c72ff6b63ff9f7dff6f65ff9576ff6f65ff846eff8c72%
  ff9576ff8c72ff846eff9576fd6561fd6561ff8c72ff8c72ff9576ff846eff7e6cff6f65ff9576ffffffffac88ffac88%
  ff9f7dffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffff9f7dff9576ffac88ffac88ff9f7d%
  ffac88ff8c72ff9f7dffac88ffac88ff9f7dff9576ffac88ffac88ff8c72ff9f7dffffffffac88ff9f7dffffffffac88%
  ff9f7dff8c72ffffffff9f7dff9576ffac88ff9576ffac88ff9f7dff9576ff8c72ff9576ff9f7dff9f7dff7e6cff6f65%
  ff7869ff7869ff7e6cff8c72ff7e6cff8c72ff9576ff7e6cff8c72ff846eff7869ff6f65ff6f65ff7e6cff6b63ff7869%
  ff846eff9f7dff9f7dff9f7dff9f7dff9f7dffac88ffffffffac88ffac88ffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffac88ffac88ff9f7dffac88ff9f7dffffffffac88ff8c72ff9f7dffac88ff9576ff8c72ff8c72%
  ff9f7dff9f7dff9f7dffac88ffffffffffffffac88ff9f7dff8c72ffffffff9f7dffac88ff9576ffac88ff9576ff9f7d%
  ff8c72ff9576ff7e6cff7367fd6561ff9f7dff7869ff9576ff846eff8c72ff846eff8c72ff7e6cff8c72ff7869ff846e%
  ff7e6cff6f65ff7367ff9576ff846eff846eff9576ff7e6cffffffff9f7dff8c72ffac88ffac88ffac88ffffffffffff%
  ffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ffffffffffffff9576ff9576ffac88ff9576ff9f7d%
  ff9576ffac88ffac88ffac88ffac88ffffffffffffffffffffffffff9f7dffac88ff9f7dffac88ffffffff8c72ff9f7d%
  ff9576ffffffff9f7dff846effac88ff8c72ff9576ff7367ff7869ff8c72ff8c72ff7367ffac88ff846eff8c72ff9f7d%
  ff8c72ff9576ff7e6cff7e6cffac88ff9f7dfb605fff846eff9f7dff8c72ff9576ff9576ff9f7dff7e6cff8c72ffffff%
  ffffffffac88ffac88ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffffffffffffffffffffffff9f7dff9f7dffac88%
  ffac88ffac88ffffffff9f7dffac88ffac88ffac88ff9f7dffffffffac88ff9f7dff9f7dffffffffffffffac88ff9f7d%
  ffac88ff9f7dffac88ffac88ffac88ffac88ff9576ff9f7dff9576ff8c72ffac88ff9576ff9576ff7e6cff7869ff8c72%
  ff6f65ff9576ff8c72ff7e6cff9576ff8c72ff846eff8c72ffac88ff7e6cff8c72ff6f65ff8c72ff9f7dff8c72ff7367%
  ff9576ff846ef6545cff7e6cffac88ffac88ffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88%
  ffac88ff8c72ffac88ff9f7dff846effffffffffffffffffffac88ffffffffac88ff9f7dff9576ff9576ff9576ff7e6c%
  ffffffffac88ffac88ffffffffffffff9576ff9f7dff846eff9576ffac88ff8c72ff9576ff846eff8c72ff9576ff7869%
  ff9f7dff9576ff7e6cff9576ff8c72ff7869ff9576ff9f7dffffffff846eff7367ff7869ff7e6cff846eff9f7dff7e6c%
  ff846eff9576ff9576ff6f65ff7367ff846ef6545cf95c5eff9f7dffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ff9576ffac88ffac88ff9f7dffac88ffac88ffac88ff9f7dffac88ffac88%
  ffffffff8c72ff7869ff6f65ff7e6cff9f7dff8c72ffac88ff9f7dffffffff7869ffffffff9f7dffac88ff9576ff846e%
  ff8c72ff8c72ff7367ff7367ff7367ff6f65ff9576ff846eff9576ff846eff7367ff8c72ff8c72ff8c72ff9f7dff9576%
  ff7e6cff846eff7e6cff8c72ffac88ffac88ff8c72ff7e6cfa5e5ffd6561fd6561f24a59ff7e6cffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffac88ff9f7dffffffffac88ff9f7dffffffffffffff9f7dff9f7d%
  ffffffffac88ff9576ffac88ffffffffffffff9576ff8c72ff6f65fe6862ff8c72ff9f7dff9576ffffffffffffff8c72%
  ff8c72ff8c72ff846eff8c72ffac88ff8c72ff9576ff7367ff846eff9576ff7367ff7e6cff7869ff846eff9f7dff846e%
  ffac88ff9f7dff846eff7e6cff8c72ff846eff7367ff9f7dff846eff846eff9f7dff7e6cff7367f7565cfe6862ff7367%
  ff7367ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffff%
  ffac88ffffffffac88ffffffffffffffac88ffffffffac88ffffffff9576ffffffffac88ff9576ff7e6cff846effffff%
  ffac88ff9f7dff9576ffac88ff9f7dff9576ffac88ff7e6cff846eff7869ff7869ff6b63ff846eff6f65fe6862ff6b63%
  ff7869ff6b63ff7e6cff6f65ff7869ff8c72ff7e6cff9576ff7e6cff8c72ffac88ff9f7dffffffffffffff8c72ffac88%
  ff7e6cff6f65ff6b63ff846eff7e6cffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ff9f7dff9f7dffac88ffac88ffffffffffffff9f7dffffffffac88ffac88ffffffffac88ffac88ff9f7dffffff%
  ffffffff9f7dffac88ffffffff9576ffffffffffffffffffffac88ffffffff9f7dffac88ff7e6cff9576ff846eff7e6c%
  fd6561ff7367fc6260ff6f65fe6862ff6b63ff6f65ff6b63ff7367ff846eff8c72ff9f7dff9576ff9576ffac88ff846e%
  ff9f7dffffffffac88ffac88ff9f7dff8c72ff8c72ff846eff9576ffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffff9f7dffac88ffffffff9f7dffffffffac88ffac88ffffff%
  ffac88ffffffffac88ffac88ffac88ff9576ff9576ff9f7dffac88ff9f7dff9f7dff9576ff8c72ffac88ff9576ff8c72%
  ff9576ff8c72ff7869ff7367fd6561fa5e5fff6b63ff6f65f95c5efd6561f6545cf85a5df5515bfa5e5fff6f65ff7869%
  ff9576ff8c72ff9f7dff846eff8c72ff9f7dff8c72ff9576ffffffffffffffffffff9f7dffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffac88ffffffffac88%
  ff9f7dffac88ffac88ff7e6cffac88ff9f7dffffffff9f7dffffffffffffffffffff9f7dffffffffac88ffac88ff9f7d%
  ff9f7dffffffffac88ff7869ff7e6cff846eff7869ff846eff846eff6f65fc6260f95c5efc6260f85a5df34d59f6545c%
  f4505af5535bfd6561ff7e6cff7869ff7e6cff846effac88ffac88ffac88ff9f7dffac88ffac88ffac88ff9f7dffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffac88%
  ffffffffffffffac88ffffffffac88ffffffffffffffac88ffffffff9f7dffac88ffac88ffffffffffffffac88ffffff%
  ff9f7dffac88ffffffffac88ff9576ff7e6cff9f7dff9f7dff9576ffac88ff8c72ff7869fe6862ff6f65ff7367ff6f65%
  fa5e5ff85a5df34d59fb605ffa5e5ff7585dfc6260f95c5eff6f65ff6b63ff9576ff9f7dff9f7dffac88ffffffffac88%
  ff9f7dffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ff9576ffac88ffac88ff9f7dffffffff9f7dffffffffac88ffffff%
  ffac88ffac88ffac88ffffffffffffffffffffffffffffffffac88ffac88ff9f7dff846eff7e6cff7367ff7e6cff8c72%
  ff7869ff846eff7367fd6561f95c5ef7585df7565cf4505af6545cf5515bf4505af85a5dfe6862ff8c72ff9576ff7869%
  ff9f7dff8c72ff9f7dffffffffac88ff9f7dffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffffffffffffffac88ffac88%
  ffac88ffac88ff9f7dffffffffffffff9f7dffffffffffffffffffffffffffac88ff9576ff9f7dff9576ffac88ffffff%
  ffac88ffac88ff6f65ff8c72ff846efd6561ff7869fb605ff7585df5515bfc6260f34e5af7585df6545cf4505af7565c%
  f95c5eff6b63ff7e6cff9576ff846eff8c72ff9f7dff7869ffffffffffffffac88ffffffffffffffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffac88ff9f7dffffffff9f7dff9f7dff9f7dffffffff846effac88ffffffffffffffffffffac88ffac88ff9576%
  ffac88ffffffff9f7dff9576ff8c72ffffffff7e6cff846eff7869ff7367fe6862ff7e6cf5535bfd6561f34e5afb605f%
  f5515bf7585df4505af5515bf5515bff6b63ff6f65ff8c72ff9576ff9f7dff9f7dffac88ffac88ffac88ffffffffac88%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffff9f7dffffffffac88ff9f7dff8c72ff9f7dff9f7dffac88ffffffffffff%
  ffac88ffffffffffffffffffffffffff9f7dffac88ffac88ff9576ff8c72ff7869ff846eff8c72ff846eff7869f7565c%
  ff6b63f7585df6545cf6545cf95c5ef24b59f04757f5535bfd6561f5535bfd6561ff7367ff7367ff7e6cff846eff9f7d%
  ff8c72ff8c72ff9f7dffac88ffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ff9f7dff9f7dff846e%
  ff7e6cff9f7dff9f7dff9f7dffffffffffffffac88ff9f7dffffffff9576ffac88ffac88ff9f7dff9f7dff846eff9f7d%
  ff9f7dff9576ff7869ff7367fc6260fe6862fc6260fa5e5ff34d59fb605ff7565cf5535bf6545cf7585dff7367ff7869%
  ff7367ff7869ff6f65ff7e6cffffffff9f7dff9f7dffffffffac88ffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ff7869ff9f7dff9f7dff9f7dffac88ff9f7dffffffffffffffffffff9f7dffac88ff9576%
  ffac88ff9f7dff846eff9576ff8c72ff7367ff7869ff6f65ff7869ff7e6cfb605ffa5e5ffa5e5ff7585df7565cf5515b%
  f7565cf4505afe6862ff7367ff6f65ff7367ff9f7dffac88ff9576ff9f7dff9f7dff9f7dffffffffac88ffac88ffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ff9f7dff9f7dff9f7dffac88ff9f7dff8c72ff7869ff846eff8c72ffffffffffffffac88%
  ffffffffffffffffffff9f7dffffffff8c72ff9576ffffffffac88ff7367ff846eff7367ff846eff7869fc6260f5535b%
  fb605ff85a5df7565cf85a5dfe6862fa5e5ffc6260ff7e6cff846eff7367ffac88ff9f7dffac88ffac88ff9f7dff8c72%
  ffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffffffffff9f7dffac88ff9576ff9576ffac88ffac88ff9576ff846effac88ff8c72%
  ffac88ffac88ffac88ffffffffffffffffffff9f7dff9f7dff8c72ffffffffffffff9f7dffac88ff9f7dff7e6cff8c72%
  ff846eff6b63fc6260fe6862f7565cf4505afd6561fa5e5ffa5e5ffa5e5ffd6561ff7367ff8c72ff846eff6f65ff7367%
  ffac88ff9576ffac88ff9f7dffac88ffffffffffffffffffffffffff9576ffffffffac88ffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ff9f7dff9576ff8c72ff846e%
  ff8c72ff9f7dff9576ffac88ff9f7dff9f7dffffffffffffffffffff9f7dffffffff9f7dffac88ffffffffac88ff9576%
  ff846eff9576ff8c72ff9f7dff7e6cff7869ff6b63ff7367fd6561f6545cfe6862f95c5ef7565cf5535bff6b63fe6862%
  ff7367ff846eff9f7dff7367f95c5eff7367ffac88ff9f7dff9f7dffac88ff9f7dffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9f7dffffffffffffffffff%
  ff9576ffffffff7869ff846eff9576ff9f7dff9f7dffac88ffac88ffac88ffac88ffffffffac88ffac88ff9f7dffac88%
  ff9f7dffac88ff9576ffffffff9576ffac88ffac88ff7e6cff6f65ff8c72ff7869ff9576fd6561f85a5dfb605ffb605f%
  f5515bfb605ffc6260fc6260fb605fff846eff846efc6260f34e5af5535bff9576ff9f7dffac88ffffffffffffff9f7d%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffff9576ffffffffac88ffac88ff9576ff9576ff8c72ffffffff9f7dffac88ffffffff9576ffac88ffac88ff7e6c%
  ff7869ff7e6cff9576ffac88ff9f7dff9576ffac88ffac88ffac88ffac88ffac88ffffffffac88ff9576ffffffff9f7d%
  ff7e6cff8c72ff7367ff6b63fe6862fd6561f7585dfa5e5fff6f65fd6561ff8c72fe6862f7565cea3d54f34d59ff9576%
  ffffffffffffffac88ffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffff9f7dffffffff9576ffac88ff9576ffac88ff9f7dffac88ff9576ff9576%
  ffffffff9576ffffffff6b63ff7869f95c5eff7e6cff7e6cff8c72ff9f7dffac88ffac88ff846eff8c72ff9f7dff9576%
  ff8c72ff9576ff9f7dff8c72ff846eff8c72ff7869ff846eff9576ff7367fc6260fe6862f5535bfe6862ff6f65ff7e6c%
  fe6862f7585def4457ff6b63ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff8c72ffac88ffac88ffac88%
  ff9f7dffffffffac88ffac88ffffffffac88ffac88ffac88ff9f7dff7367ff7e6cff9f7dff9f7dffac88ff9f7dffac88%
  ffffffffac88ff8c72ffac88ff9576ff9f7dff8c72ff9576ff846eff7e6cff7e6cfe6862ff846eff6f65ff6f65ff6f65%
  fd6561fd6561fd6561ff7e6cff9f7dff6b63fd6561ff7e6cff9f7dffac88ffffffffffffffffffffffffffffffffac88%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ff8c72ffffffffac88ffffffff9f7dffac88ffffffffffffffac88ffac88ffffffff9f7dff846eff9576ff9f7dff9f7d%
  ffffffff9f7dffac88ffac88ffffffffffffff9576ffffffff8c72ff9576ff9f7dffac88ff9f7dff9576ff9576ff9576%
  fd6561fd6561ff7869ff7e6cff7e6cf7565cff7367ff7869ff6f65ff8c72ff9576ff8c72ff9576ffac88ffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ffac88ffffffff9f7dffffffffffffffffffffac88%
  ffac88ff9f7dff9f7dff9f7dffac88ffac88ffffffffffffff9f7dff8c72ffac88ffac88ffac88ff9576ff7e6cff6b63%
  ff9f7dff9576ff9f7dff8c72ff9576ff9576ff846eff7367ff7e6cff7367ff7869ff8c72ff7869ff8c72ffac88ff9576%
  ff9576ff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88ffffffffac88ffac88%
  ffac88ffffffffffffffffffff9f7dffac88ff9f7dffac88ffffffffffffff9f7dffffffff9f7dffac88ff8c72ff9f7d%
  ff9576ff9576ff7869ff846eff9f7dff9f7dffffffff7e6cff9f7dff7e6cff846eff6b63ff6f65ff7e6cffac88ff846e%
  ff8c72ff9576ff7e6cffac88ffac88ffffffffffffffffffffffffffac88ffffffffac88ffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff9576ffffffffac88ffffffffac88ffffffff9576ffac88ff9f7dff9f7dffac88ff9f7dffffff%
  ff9f7dffac88ff9576ff9576ff9576ff9f7dff9576ff8c72ff8c72ff9576ff7367ff7367ff7e6cff7869ff8c72ff8c72%
  ff7869ff7e6cff846effac88ffac88ff9f7dff8c72ffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffac88ffac88ff9f7dffac88ffac88ffffffffffffffac88ffac88%
  ff846effac88ffffffffac88ffffffff9f7dffac88ffac88ff9f7dff846effac88ff9576ff8c72ff8c72ff7869ff9f7d%
  ff8c72ffac88ff7e6cff6f65ff7e6cff846eff846eff6f65ff846effac88ffac88ffac88ffac88ffffffffffffffffff%
  ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffffffffac88ffffffffac88ffffff%
  ff9f7dffac88ffffffff9576ff8c72ffac88ff846eff9f7dffac88ff9f7dff9f7dff8c72ff9576ff9576ff9f7dffac88%
  ff8c72ff9f7dff9f7dff9f7dffac88ffac88ff9f7dff9576ff8c72ff9576ff7e6cff846effffffff9f7dff9f7dffac88%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffff%
  ffffffffac88ffac88ff9576ff9f7dff846eff9f7dff9576ff8c72ffac88ff9576ff8c72ffac88ff9576ff9f7dffac88%
  ff9f7dff8c72ff9f7dff6f65ff8c72ff8c72ff9576ff846eff9576ffac88ffac88ff7e6cff8c72ff8c72ff7e6cff7869%
  ff9f7dff8c72ffffffffac88ffac88ffac88ffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffff9f7dff9f7dffac88ff9f7dff8c72ff7869ff7367ff7e6cff9f7dffffffff8c72ff846e%
  ff9f7dffac88ffffffffffffffffffff846eff7e6cff7e6cff9576ff8c72ffac88ff846eff9f7dff846effffffff9f7d%
  ff7869ffac88ff9f7dff8c72ffac88ffac88ffac88ff9576ffac88ffac88ffac88ffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9576ff9576ff7869ff8c72ff7367fa5e5f%
  ff9f7dffac88ff9f7dffffffff9576ffac88ffac88ff9576ff8c72ff9f7dff9f7dff7e6cff7869ff7e6cff8c72fb605f%
  ff846eff846effac88ff846eff846effac88ff9576ffac88ff9f7dffac88ffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffff9f7dffffff%
  ff9576ffffffff7367ff7869ff7e6cff8c72ff846eff9f7dff846eff9f7dffac88ff9576ff8c72ff8c72ff9576ff846e%
  ff8c72fc6260ff7367fb605ffc6260ff7367fe6862ff6f65ff846eff8c72ff7e6cff9f7dffffffff9f7dffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffac88ffac88ff8c72ff8c72ffffffff8c72ff846eff7e6cff8c72ff9576ff9576ff9576ffffffffffffffffff%
  ff846eff9576ff846eff7e6cfb605ffe6862fd6561f95c5ef6545cfb605fff6b63f7565cff7869ff846effffffff8c72%
  ff8c72ffac88ffac88ffac88ffffffffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffac88ffac88ffac88ffac88ffac88ff7e6cff9f7dff9576ff9576ffffffffac88ff9576%
  ff9576ff9576ff8c72ff7367ff7e6cff9576ff8c72ff9576f6545cff6b63f14958f95c5ef14958fc6260f5535bf4505a%
  fa5e5ff95c5eff8c72ff846eff9576ff9f7dffffffffac88ffffffffac88ffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9576ff9576ff9f7dff9f7d%
  ff9f7dffac88ff9f7dff8c72ffac88ff846effac88ff9576ff7869ff8c72ff6f65ff6f65ff7869fb605ffb605ff24a59%
  f5515bf24a59fa5e5ff5515bf95c5eff7869ff7367ff8c72ffac88ffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ffffffffac88ffffffffffffffac88ff9576%
  ffffffffac88ff8c72ffac88ff9576ffac88ff846eff9f7dff9576ff9f7dff8c72ff846eff7e6cff7869ff9f7dff9f7d%
  fc6260ff6b63f85a5df6545cf7565cf14958ef4457f14958fa5e5ffe6862ff6b63fd6561ff9f7dffac88ffac88ffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ffffffffac88ffffffffac88ffffffffffffffac88ff9576ff8c72ff6b63fa5e5ff5535bff7869ff7869ff7e6c%
  ff6f65ff7367ff7367ff7367ff7367ff6f65ff6f65f85a5df24b59f4505af04757f04557f24b59f4505aff6b63ff8c72%
  ff8c72ff8c72ff8c72ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88ffac88ff9576ff7e6cfb605f%
  f34d59f7585dfc6260ff7e6cff7869ff6b63ff7869fd6561fc6260ff7e6cf95c5eff6b63f5535bf6545ced4155ef4457%
  f34e5af6545cf5535bfd6561fe6862ff846eff7869ff9f7dffffffffffffffffffffac88ffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dff9576ff9f7dffffffff9576ffac88%
  ff9576ff8c72ff7e6cff6f65ff7869ff8c72ff6b63ff7869fc6260ff6f65ff7869ff6f65fc6260fc6260f95c5eff6b63%
  f34e5af4505af14958ef4356f4505af34e5af14958f34e5af7565cfe6862ff9576ff7869ff9f7dffac88ffffffffffff%
  ffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7d%
  ff9576ff846effac88ffffffff846eff8c72ff7367ff7367fd6561ff7869ff6f65ff7869ff7367ff7e6cff7e6cf85a5d%
  fb605fff7367f6545cf95c5ef85a5df85a5df7565cf24a59f5515bf7585df34e5af24a59ef4457fe6862ff6b63ff6f65%
  ff846effffffffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88%
  ff9f7dffffffffffffffffffff9576ff9f7dffac88ffac88ffac88ff6f65ff8c72ff9576ff9f7dff7367fe6862ff7367%
  ff7869ff7869ff7869fe6862ff7869ff6b63ff8c72fb605ffe6862f85a5df7565cfa5e5ff04757f6545cf34d59f14858%
  e73952f24b59fd6561ff8c72ff9576ff9f7dffffffff9576ffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffffffff9f7dffac88ff9576ffac88ff7e6cff846eff8c72ff846eff7869ff7e6cff9f7d%
  ff9576ff6f65fd6561ff6b63ff8c72fe6862ff8c72ff846eff7e6cfe6862ff6f65ff7869fd6561fc6260f7565cf95c5e%
  f14858f34d59f7585df95c5ef24a59f7585df7585dff6b63ff9576ff9576ffffffffac88ffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffffffff9576ffffffff7869ff846eff8c72ff8c72%
  ff8c72ff7869ff846eff846eff7869ff7367fe6862ff7367ff9576fc6260ff846efe6862f7585dfe6862ff7367fe6862%
  f95c5efe6862f85a5df5535bf7565cf24b59f5535bf7585df7585df95c5eff6f65ff8c72ff9576ff7e6cffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9576ff846effac88ff9576%
  ff846eff7869ff9576ff846eff7367ff7e6cff9576ff846eff7e6cff7869ff6b63ff6b63ff7869fe6862f85a5dff6b63%
  ff6f65ff7367fa5e5ff6545cf95c5efa5e5ff5535bf85a5df5515bf7585def4356f85a5df7585dfd6561fe6862ff8c72%
  ff9f7dffac88ff9f7dffffffffffffffffffffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffff%
  ffac88ffac88ffac88ff9f7dffffffffac88ff846eff9576ff8c72ff9576ff7e6cff846eff7e6cff846eff7869ff6b63%
  ff846eff846eff846efc6260fe6862fd6561fb605ff4505afd6561f7585df7565cfd6561fc6260f95c5ef6545cfe6862%
  ff7367ffac88ffac88ffffffffac88ff9f7dffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffff9f7dffffffffac88ffac88ffac88ffffffff846eff7e6cff6f65ff9576ff8c72ff7869ff6f65%
  ff7367ff9576ff7e6cff7367ff7367ff6f65fe6862fc6260ff6b63f7565cf85a5df34e5af7565cf24a59f7565cff6f65%
  fa5e5ffa5e5ffb605fff7e6cff9576ff7869ff8c72ffffffffffffff9f7dffac88ffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ff9576ffffffff7e6cff7367fc6260%
  fa5e5ffa5e5ffc6260f7585df95c5eff7e6cff846eff8c72ff6f65ff7e6cfe6862fd6561fc6260fd6561f5535bf85a5d%
  f7565cfb605ffe6862ff7367ff7869ff7869ff8c72ff7e6cff846eff9f7dffac88ff9576ffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffac88%
  ffffffff9576ff9f7dff8c72ff7367ff9576ff6b63fd6561f7585df7565cff9576ff7869ffac88ff8c72ff846eff9576%
  ff6f65ff7869ff6f65ff6b63ff846eff9576ff846eff9576ff9f7dffac88ffac88ff9576ff9576ffffffffac88ffac88%
  ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffac88ffac88ff9576ff6f65ff9576ff7869f95c5ef34e5af04557fd6561%
  ff7869ff7367ff9f7dffac88ff846effac88ff9f7dff9f7dffac88ffac88ff9f7dffffffffffffffffffffffffffac88%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffffffffffff%
  ffac88ffffffffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffac88ffac88ffffffffffffffac88ffffffffac88ffac88ffac88ffffffffffffffffffffffff%
  ffffffffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffff9f7dffffffffffffff9f7dffac88ffffffffac88ffac88ff9f7d%
  ffac88ffffffffac88ffac88ff9f7dff9576ff9576ff7869ffffffff8c72ffac88ffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffac88ffac88ff8c72ffac88%
  ffffffffffffffac88ff9f7dff9f7dffffffffac88ff9f7dff9576ff846eff7869ff6f65ff7367ff846eff8c72ffac88%
  ffffffff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffac88ff8c72ff7869ff6f65ff7e6cff8c72ff7e6cff7e6cf95c5eff6b63f7585df24b59ef4457e63852e43551df304e%
  e73952e53651f14858f95c5eff6b63ff8c72ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffff9f7dff8c72ff7367ff6b63f7565cf85a5df24b59f14858f14858f4505ae53751%
  ea3d54e33450da294bda2a4cd8274adc2b4cdd2d4dee4156ff6b63ff9576ffac88ff9f7dffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffac88ff9f7dff9576ff8c72ff9576ff7367fe6862f7585dee4256f24a59eb3e54%
  ef4356e73952e23350e23450de2e4ee43651dc2c4cdc2c4cd62449d01d45d52349d9274bf04557fa5e5ffc6260ff8c72%
  ffac88ffffffffac88ffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff9f7dffac88ff9f7dff7367ff8c72ff7869%
  f24b59ef4457e63852df304ee0314fe33550ea3d54e2324fe43651df2f4ee0314fe1314fe1314fdf304ee53651eb3e54%
  ff7367ff8c72ff9f7dffac88ffac88ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88%
  ffac88ffac88ff9f7dff9576f95c5efd6561fd6561f5515bf95c5efc6260f95c5efa5e5ffe6862fd6561ff7e6cff7869%
  fd6561ff9576ff9f7dffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffac88ffffffffffffffffffffac88%
  ffffffffac88ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff8b0000%
  ae062ab2072db80c32c1133ac91840ea3d54ff6b63ff7e6cff9f7dffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffffff%
  ffffffffffffffffffffffffffffffffffffffffffffffffffffff}%
\gprawimage{rgb}{6.762}{2.179}{101}{101}{3.624}{3.624}{\gprawrgbimagedata}{}
\end{scope}
\gpcolor{rgb color={0.000,0.000,0.000}}
\draw[gp path] (6.780,2.197)--(6.816,2.233)--(6.852,2.269)--(6.889,2.306)--(6.925,2.342)%
  --(6.961,2.378)--(6.997,2.414)--(7.034,2.451)--(7.070,2.487)--(7.106,2.523)--(7.142,2.559)%
  --(7.179,2.596)--(7.215,2.632)--(7.251,2.668)--(7.287,2.704)--(7.324,2.741)--(7.360,2.777)%
  --(7.396,2.813)--(7.432,2.849)--(7.469,2.886)--(7.505,2.922)--(7.541,2.958)--(7.577,2.994)%
  --(7.614,3.031)--(7.650,3.067)--(7.686,3.103)--(7.722,3.139)--(7.759,3.176)--(7.795,3.212)%
  --(7.831,3.248)--(7.867,3.284)--(7.904,3.321)--(7.940,3.357)--(7.976,3.393)--(8.012,3.429)%
  --(8.048,3.465)--(8.085,3.502)--(8.121,3.538)--(8.157,3.574)--(8.193,3.610)--(8.230,3.647)%
  --(8.266,3.683)--(8.302,3.719)--(8.338,3.755)--(8.375,3.792)--(8.411,3.828)--(8.447,3.864)%
  --(8.483,3.900)--(8.520,3.937)--(8.556,3.973)--(8.592,4.009)--(8.628,4.045)--(8.665,4.082)%
  --(8.701,4.118)--(8.737,4.154)--(8.773,4.190)--(8.810,4.227)--(8.846,4.263)--(8.882,4.299)%
  --(8.918,4.335)--(8.955,4.372)--(8.991,4.408)--(9.027,4.444)--(9.063,4.480)--(9.100,4.517)%
  --(9.136,4.553)--(9.172,4.589)--(9.208,4.625)--(9.244,4.661)--(9.281,4.698)--(9.317,4.734)%
  --(9.353,4.770)--(9.389,4.806)--(9.426,4.843)--(9.462,4.879)--(9.498,4.915)--(9.534,4.951)%
  --(9.571,4.988)--(9.607,5.024)--(9.643,5.060)--(9.679,5.096)--(9.716,5.133)--(9.752,5.169)%
  --(9.788,5.205)--(9.824,5.241)--(9.861,5.278)--(9.897,5.314)--(9.933,5.350)--(9.969,5.386)%
  --(10.006,5.423)--(10.042,5.459)--(10.078,5.495)--(10.114,5.531)--(10.151,5.568)--(10.187,5.604)%
  --(10.223,5.640)--(10.259,5.676)--(10.296,5.713)--(10.332,5.749)--(10.368,5.785);
\gpcolor{color=gp lt color border}
\draw[gp path] (6.780,2.197)--(6.960,2.197);
\node[gp node right] at (6.596,2.197) {1};
\draw[gp path] (6.780,2.744)--(6.960,2.744);
\node[gp node right] at (6.596,2.744) {10};
\draw[gp path] (6.780,5.481)--(6.960,5.481);
\draw[gp path] (6.780,5.785)--(6.960,5.785);
\node[gp node right] at (6.596,5.785) {1h};
\draw[gp path] (6.780,2.197)--(6.960,2.197);
\draw[gp path] (6.780,2.362)--(6.870,2.362);
\draw[gp path] (6.780,2.580)--(6.870,2.580);
\draw[gp path] (6.780,2.691)--(6.870,2.691);
\draw[gp path] (6.780,2.744)--(6.960,2.744);
\draw[gp path] (6.780,2.909)--(6.870,2.909);
\draw[gp path] (6.780,3.127)--(6.870,3.127);
\draw[gp path] (6.780,3.238)--(6.870,3.238);
\draw[gp path] (6.780,3.292)--(6.960,3.292);
\node[gp node right] at (6.596,3.292) {$10^{2}$};
\draw[gp path] (6.780,3.456)--(6.870,3.456);
\draw[gp path] (6.780,3.674)--(6.870,3.674);
\draw[gp path] (6.780,3.786)--(6.870,3.786);
\draw[gp path] (6.780,3.839)--(6.960,3.839);
\node[gp node right] at (6.596,3.839) {$10^{3}$};
\draw[gp path] (6.780,4.004)--(6.870,4.004);
\draw[gp path] (6.780,4.221)--(6.870,4.221);
\draw[gp path] (6.780,4.333)--(6.870,4.333);
\draw[gp path] (6.780,4.386)--(6.960,4.386);
\node[gp node right] at (6.596,4.386) {$10^{4}$};
\draw[gp path] (6.780,4.551)--(6.870,4.551);
\draw[gp path] (6.780,4.769)--(6.870,4.769);
\draw[gp path] (6.780,4.880)--(6.870,4.880);
\draw[gp path] (6.780,4.933)--(6.960,4.933);
\node[gp node right] at (6.596,4.933) {$10^{5}$};
\draw[gp path] (6.780,5.098)--(6.870,5.098);
\draw[gp path] (6.780,5.316)--(6.870,5.316);
\draw[gp path] (6.780,5.428)--(6.870,5.428);
\draw[gp path] (6.780,5.481)--(6.960,5.481);
\draw[gp path] (6.780,5.645)--(6.870,5.645);
\draw[gp path] (6.780,2.197)--(6.780,2.377);
\node[gp node center] at (6.780,1.889) {1};
\draw[gp path] (7.327,2.197)--(7.327,2.377);
\node[gp node center] at (7.327,1.889) {10};
\draw[gp path] (10.064,2.197)--(10.064,2.377);
\draw[gp path] (10.368,2.197)--(10.368,2.377);
\node[gp node center] at (10.368,1.889) {1h};
\draw[gp path] (6.780,2.197)--(6.780,2.377);
\draw[gp path] (6.945,2.197)--(6.945,2.287);
\draw[gp path] (7.163,2.197)--(7.163,2.287);
\draw[gp path] (7.274,2.197)--(7.274,2.287);
\draw[gp path] (7.327,2.197)--(7.327,2.377);
\draw[gp path] (7.492,2.197)--(7.492,2.287);
\draw[gp path] (7.710,2.197)--(7.710,2.287);
\draw[gp path] (7.821,2.197)--(7.821,2.287);
\draw[gp path] (7.875,2.197)--(7.875,2.377);
\node[gp node center] at (7.875,1.889) {$10^{2}$};
\draw[gp path] (8.039,2.197)--(8.039,2.287);
\draw[gp path] (8.257,2.197)--(8.257,2.287);
\draw[gp path] (8.369,2.197)--(8.369,2.287);
\draw[gp path] (8.422,2.197)--(8.422,2.377);
\node[gp node center] at (8.422,1.889) {$10^{3}$};
\draw[gp path] (8.587,2.197)--(8.587,2.287);
\draw[gp path] (8.804,2.197)--(8.804,2.287);
\draw[gp path] (8.916,2.197)--(8.916,2.287);
\draw[gp path] (8.969,2.197)--(8.969,2.377);
\node[gp node center] at (8.969,1.889) {$10^{4}$};
\draw[gp path] (9.134,2.197)--(9.134,2.287);
\draw[gp path] (9.352,2.197)--(9.352,2.287);
\draw[gp path] (9.463,2.197)--(9.463,2.287);
\draw[gp path] (9.516,2.197)--(9.516,2.377);
\node[gp node center] at (9.516,1.889) {$10^{5}$};
\draw[gp path] (9.681,2.197)--(9.681,2.287);
\draw[gp path] (9.899,2.197)--(9.899,2.287);
\draw[gp path] (10.011,2.197)--(10.011,2.287);
\draw[gp path] (10.064,2.197)--(10.064,2.377);
\draw[gp path] (10.228,2.197)--(10.228,2.287);
\draw[gp path] (6.780,5.785)--(6.780,2.197)--(10.368,2.197);
\gpfill{rgb color={1.000,0.998,0.996}} (10.920,2.197)--(11.138,2.197)--(11.138,2.226)--(10.920,2.226)--cycle;
\gpfill{rgb color={1.000,0.934,0.895}} (10.920,2.225)--(11.138,2.225)--(11.138,2.254)--(10.920,2.254)--cycle;
\gpfill{rgb color={1.000,0.871,0.793}} (10.920,2.253)--(11.138,2.253)--(11.138,2.282)--(10.920,2.282)--cycle;
\gpfill{rgb color={1.000,0.807,0.692}} (10.920,2.281)--(11.138,2.281)--(11.138,2.310)--(10.920,2.310)--cycle;
\gpfill{rgb color={1.000,0.743,0.590}} (10.920,2.309)--(11.138,2.309)--(11.138,2.313)--(10.920,2.313)--cycle;
\gpfill{rgb color={1.000,0.737,0.580}} (10.920,2.312)--(11.138,2.312)--(11.138,2.338)--(10.920,2.338)--cycle;
\gpfill{rgb color={1.000,0.731,0.575}} (10.920,2.337)--(11.138,2.337)--(11.138,2.366)--(10.920,2.366)--cycle;
\gpfill{rgb color={1.000,0.724,0.570}} (10.920,2.365)--(11.138,2.365)--(11.138,2.394)--(10.920,2.394)--cycle;
\gpfill{rgb color={1.000,0.717,0.564}} (10.920,2.393)--(11.138,2.393)--(11.138,2.422)--(10.920,2.422)--cycle;
\gpfill{rgb color={1.000,0.710,0.559}} (10.920,2.421)--(11.138,2.421)--(11.138,2.450)--(10.920,2.450)--cycle;
\gpfill{rgb color={1.000,0.703,0.553}} (10.920,2.449)--(11.138,2.449)--(11.138,2.478)--(10.920,2.478)--cycle;
\gpfill{rgb color={1.000,0.696,0.548}} (10.920,2.477)--(11.138,2.477)--(11.138,2.506)--(10.920,2.506)--cycle;
\gpfill{rgb color={1.000,0.689,0.542}} (10.920,2.505)--(11.138,2.505)--(11.138,2.534)--(10.920,2.534)--cycle;
\gpfill{rgb color={1.000,0.682,0.537}} (10.920,2.533)--(11.138,2.533)--(11.138,2.562)--(10.920,2.562)--cycle;
\gpfill{rgb color={1.000,0.675,0.531}} (10.920,2.561)--(11.138,2.561)--(11.138,2.590)--(10.920,2.590)--cycle;
\gpfill{rgb color={1.000,0.668,0.526}} (10.920,2.589)--(11.138,2.589)--(11.138,2.618)--(10.920,2.618)--cycle;
\gpfill{rgb color={1.000,0.661,0.520}} (10.920,2.617)--(11.138,2.617)--(11.138,2.646)--(10.920,2.646)--cycle;
\gpfill{rgb color={1.000,0.654,0.515}} (10.920,2.645)--(11.138,2.645)--(11.138,2.674)--(10.920,2.674)--cycle;
\gpfill{rgb color={1.000,0.647,0.509}} (10.920,2.673)--(11.138,2.673)--(11.138,2.702)--(10.920,2.702)--cycle;
\gpfill{rgb color={1.000,0.640,0.504}} (10.920,2.701)--(11.138,2.701)--(11.138,2.730)--(10.920,2.730)--cycle;
\gpfill{rgb color={1.000,0.633,0.498}} (10.920,2.729)--(11.138,2.729)--(11.138,2.758)--(10.920,2.758)--cycle;
\gpfill{rgb color={1.000,0.626,0.493}} (10.920,2.757)--(11.138,2.757)--(11.138,2.786)--(10.920,2.786)--cycle;
\gpfill{rgb color={1.000,0.619,0.487}} (10.920,2.785)--(11.138,2.785)--(11.138,2.814)--(10.920,2.814)--cycle;
\gpfill{rgb color={1.000,0.612,0.482}} (10.920,2.813)--(11.138,2.813)--(11.138,2.842)--(10.920,2.842)--cycle;
\gpfill{rgb color={1.000,0.605,0.476}} (10.920,2.841)--(11.138,2.841)--(11.138,2.870)--(10.920,2.870)--cycle;
\gpfill{rgb color={1.000,0.598,0.471}} (10.920,2.869)--(11.138,2.869)--(11.138,2.892)--(10.920,2.892)--cycle;
\gpfill{rgb color={1.000,0.592,0.467}} (10.920,2.891)--(11.138,2.891)--(11.138,2.898)--(10.920,2.898)--cycle;
\gpfill{rgb color={1.000,0.590,0.466}} (10.920,2.897)--(11.138,2.897)--(11.138,2.926)--(10.920,2.926)--cycle;
\gpfill{rgb color={1.000,0.581,0.462}} (10.920,2.925)--(11.138,2.925)--(11.138,2.954)--(10.920,2.954)--cycle;
\gpfill{rgb color={1.000,0.573,0.458}} (10.920,2.953)--(11.138,2.953)--(11.138,2.982)--(10.920,2.982)--cycle;
\gpfill{rgb color={1.000,0.564,0.454}} (10.920,2.981)--(11.138,2.981)--(11.138,3.010)--(10.920,3.010)--cycle;
\gpfill{rgb color={1.000,0.555,0.450}} (10.920,3.009)--(11.138,3.009)--(11.138,3.038)--(10.920,3.038)--cycle;
\gpfill{rgb color={1.000,0.546,0.446}} (10.920,3.037)--(11.138,3.037)--(11.138,3.066)--(10.920,3.066)--cycle;
\gpfill{rgb color={1.000,0.538,0.442}} (10.920,3.065)--(11.138,3.065)--(11.138,3.095)--(10.920,3.095)--cycle;
\gpfill{rgb color={1.000,0.529,0.438}} (10.920,3.094)--(11.138,3.094)--(11.138,3.123)--(10.920,3.123)--cycle;
\gpfill{rgb color={1.000,0.520,0.434}} (10.920,3.122)--(11.138,3.122)--(11.138,3.151)--(10.920,3.151)--cycle;
\gpfill{rgb color={1.000,0.511,0.430}} (10.920,3.150)--(11.138,3.150)--(11.138,3.179)--(10.920,3.179)--cycle;
\gpfill{rgb color={1.000,0.503,0.426}} (10.920,3.178)--(11.138,3.178)--(11.138,3.207)--(10.920,3.207)--cycle;
\gpfill{rgb color={1.000,0.494,0.422}} (10.920,3.206)--(11.138,3.206)--(11.138,3.235)--(10.920,3.235)--cycle;
\gpfill{rgb color={1.000,0.485,0.418}} (10.920,3.234)--(11.138,3.234)--(11.138,3.263)--(10.920,3.263)--cycle;
\gpfill{rgb color={1.000,0.476,0.414}} (10.920,3.262)--(11.138,3.262)--(11.138,3.291)--(10.920,3.291)--cycle;
\gpfill{rgb color={1.000,0.468,0.410}} (10.920,3.290)--(11.138,3.290)--(11.138,3.319)--(10.920,3.319)--cycle;
\gpfill{rgb color={1.000,0.459,0.406}} (10.920,3.318)--(11.138,3.318)--(11.138,3.347)--(10.920,3.347)--cycle;
\gpfill{rgb color={1.000,0.450,0.402}} (10.920,3.346)--(11.138,3.346)--(11.138,3.375)--(10.920,3.375)--cycle;
\gpfill{rgb color={1.000,0.441,0.398}} (10.920,3.374)--(11.138,3.374)--(11.138,3.403)--(10.920,3.403)--cycle;
\gpfill{rgb color={1.000,0.433,0.394}} (10.920,3.402)--(11.138,3.402)--(11.138,3.431)--(10.920,3.431)--cycle;
\gpfill{rgb color={1.000,0.424,0.390}} (10.920,3.430)--(11.138,3.430)--(11.138,3.459)--(10.920,3.459)--cycle;
\gpfill{rgb color={1.000,0.415,0.386}} (10.920,3.458)--(11.138,3.458)--(11.138,3.471)--(10.920,3.471)--cycle;
\gpfill{rgb color={1.000,0.412,0.384}} (10.920,3.470)--(11.138,3.470)--(11.138,3.487)--(10.920,3.487)--cycle;
\gpfill{rgb color={0.998,0.407,0.383}} (10.920,3.486)--(11.138,3.486)--(11.138,3.515)--(10.920,3.515)--cycle;
\gpfill{rgb color={0.995,0.400,0.381}} (10.920,3.514)--(11.138,3.514)--(11.138,3.543)--(10.920,3.543)--cycle;
\gpfill{rgb color={0.992,0.393,0.378}} (10.920,3.542)--(11.138,3.542)--(11.138,3.571)--(10.920,3.571)--cycle;
\gpfill{rgb color={0.988,0.385,0.376}} (10.920,3.570)--(11.138,3.570)--(11.138,3.599)--(10.920,3.599)--cycle;
\gpfill{rgb color={0.985,0.378,0.374}} (10.920,3.598)--(11.138,3.598)--(11.138,3.627)--(10.920,3.627)--cycle;
\gpfill{rgb color={0.982,0.370,0.372}} (10.920,3.626)--(11.138,3.626)--(11.138,3.655)--(10.920,3.655)--cycle;
\gpfill{rgb color={0.979,0.363,0.369}} (10.920,3.654)--(11.138,3.654)--(11.138,3.683)--(10.920,3.683)--cycle;
\gpfill{rgb color={0.975,0.356,0.367}} (10.920,3.682)--(11.138,3.682)--(11.138,3.711)--(10.920,3.711)--cycle;
\gpfill{rgb color={0.972,0.348,0.365}} (10.920,3.710)--(11.138,3.710)--(11.138,3.739)--(10.920,3.739)--cycle;
\gpfill{rgb color={0.969,0.341,0.362}} (10.920,3.738)--(11.138,3.738)--(11.138,3.767)--(10.920,3.767)--cycle;
\gpfill{rgb color={0.966,0.333,0.360}} (10.920,3.766)--(11.138,3.766)--(11.138,3.795)--(10.920,3.795)--cycle;
\gpfill{rgb color={0.963,0.326,0.358}} (10.920,3.794)--(11.138,3.794)--(11.138,3.823)--(10.920,3.823)--cycle;
\gpfill{rgb color={0.959,0.319,0.356}} (10.920,3.822)--(11.138,3.822)--(11.138,3.851)--(10.920,3.851)--cycle;
\gpfill{rgb color={0.956,0.311,0.353}} (10.920,3.850)--(11.138,3.850)--(11.138,3.879)--(10.920,3.879)--cycle;
\gpfill{rgb color={0.953,0.304,0.351}} (10.920,3.878)--(11.138,3.878)--(11.138,3.907)--(10.920,3.907)--cycle;
\gpfill{rgb color={0.950,0.296,0.349}} (10.920,3.906)--(11.138,3.906)--(11.138,3.935)--(10.920,3.935)--cycle;
\gpfill{rgb color={0.946,0.289,0.347}} (10.920,3.934)--(11.138,3.934)--(11.138,3.963)--(10.920,3.963)--cycle;
\gpfill{rgb color={0.943,0.282,0.344}} (10.920,3.962)--(11.138,3.962)--(11.138,3.992)--(10.920,3.992)--cycle;
\gpfill{rgb color={0.940,0.274,0.342}} (10.920,3.991)--(11.138,3.991)--(11.138,4.020)--(10.920,4.020)--cycle;
\gpfill{rgb color={0.937,0.266,0.340}} (10.920,4.019)--(11.138,4.019)--(11.138,4.048)--(10.920,4.048)--cycle;
\gpfill{rgb color={0.933,0.259,0.337}} (10.920,4.047)--(11.138,4.047)--(11.138,4.049)--(10.920,4.049)--cycle;
\gpfill{rgb color={0.933,0.259,0.337}} (10.920,4.048)--(11.138,4.048)--(11.138,4.076)--(10.920,4.076)--cycle;
\gpfill{rgb color={0.928,0.252,0.334}} (10.920,4.075)--(11.138,4.075)--(11.138,4.104)--(10.920,4.104)--cycle;
\gpfill{rgb color={0.923,0.246,0.332}} (10.920,4.103)--(11.138,4.103)--(11.138,4.132)--(10.920,4.132)--cycle;
\gpfill{rgb color={0.918,0.239,0.329}} (10.920,4.131)--(11.138,4.131)--(11.138,4.160)--(10.920,4.160)--cycle;
\gpfill{rgb color={0.912,0.232,0.326}} (10.920,4.159)--(11.138,4.159)--(11.138,4.188)--(10.920,4.188)--cycle;
\gpfill{rgb color={0.907,0.226,0.323}} (10.920,4.187)--(11.138,4.187)--(11.138,4.216)--(10.920,4.216)--cycle;
\gpfill{rgb color={0.902,0.219,0.320}} (10.920,4.215)--(11.138,4.215)--(11.138,4.244)--(10.920,4.244)--cycle;
\gpfill{rgb color={0.896,0.213,0.317}} (10.920,4.243)--(11.138,4.243)--(11.138,4.272)--(10.920,4.272)--cycle;
\gpfill{rgb color={0.891,0.206,0.315}} (10.920,4.271)--(11.138,4.271)--(11.138,4.300)--(10.920,4.300)--cycle;
\gpfill{rgb color={0.886,0.199,0.312}} (10.920,4.299)--(11.138,4.299)--(11.138,4.328)--(10.920,4.328)--cycle;
\gpfill{rgb color={0.880,0.193,0.309}} (10.920,4.327)--(11.138,4.327)--(11.138,4.356)--(10.920,4.356)--cycle;
\gpfill{rgb color={0.875,0.186,0.306}} (10.920,4.355)--(11.138,4.355)--(11.138,4.384)--(10.920,4.384)--cycle;
\gpfill{rgb color={0.870,0.179,0.303}} (10.920,4.383)--(11.138,4.383)--(11.138,4.412)--(10.920,4.412)--cycle;
\gpfill{rgb color={0.864,0.173,0.300}} (10.920,4.411)--(11.138,4.411)--(11.138,4.440)--(10.920,4.440)--cycle;
\gpfill{rgb color={0.859,0.166,0.297}} (10.920,4.439)--(11.138,4.439)--(11.138,4.468)--(10.920,4.468)--cycle;
\gpfill{rgb color={0.854,0.159,0.295}} (10.920,4.467)--(11.138,4.467)--(11.138,4.496)--(10.920,4.496)--cycle;
\gpfill{rgb color={0.848,0.153,0.292}} (10.920,4.495)--(11.138,4.495)--(11.138,4.524)--(10.920,4.524)--cycle;
\gpfill{rgb color={0.843,0.146,0.289}} (10.920,4.523)--(11.138,4.523)--(11.138,4.552)--(10.920,4.552)--cycle;
\gpfill{rgb color={0.838,0.139,0.286}} (10.920,4.551)--(11.138,4.551)--(11.138,4.580)--(10.920,4.580)--cycle;
\gpfill{rgb color={0.833,0.133,0.283}} (10.920,4.579)--(11.138,4.579)--(11.138,4.608)--(10.920,4.608)--cycle;
\gpfill{rgb color={0.827,0.126,0.280}} (10.920,4.607)--(11.138,4.607)--(11.138,4.628)--(10.920,4.628)--cycle;
\gpfill{rgb color={0.823,0.121,0.278}} (10.920,4.627)--(11.138,4.627)--(11.138,4.636)--(10.920,4.636)--cycle;
\gpfill{rgb color={0.822,0.120,0.277}} (10.920,4.635)--(11.138,4.635)--(11.138,4.664)--(10.920,4.664)--cycle;
\gpfill{rgb color={0.815,0.115,0.272}} (10.920,4.663)--(11.138,4.663)--(11.138,4.692)--(10.920,4.692)--cycle;
\gpfill{rgb color={0.809,0.111,0.267}} (10.920,4.691)--(11.138,4.691)--(11.138,4.720)--(10.920,4.720)--cycle;
\gpfill{rgb color={0.802,0.106,0.262}} (10.920,4.719)--(11.138,4.719)--(11.138,4.748)--(10.920,4.748)--cycle;
\gpfill{rgb color={0.796,0.101,0.256}} (10.920,4.747)--(11.138,4.747)--(11.138,4.776)--(10.920,4.776)--cycle;
\gpfill{rgb color={0.789,0.096,0.251}} (10.920,4.775)--(11.138,4.775)--(11.138,4.804)--(10.920,4.804)--cycle;
\gpfill{rgb color={0.783,0.092,0.246}} (10.920,4.803)--(11.138,4.803)--(11.138,4.832)--(10.920,4.832)--cycle;
\gpfill{rgb color={0.776,0.087,0.241}} (10.920,4.831)--(11.138,4.831)--(11.138,4.860)--(10.920,4.860)--cycle;
\gpfill{rgb color={0.770,0.082,0.236}} (10.920,4.859)--(11.138,4.859)--(11.138,4.889)--(10.920,4.889)--cycle;
\gpfill{rgb color={0.763,0.077,0.231}} (10.920,4.888)--(11.138,4.888)--(11.138,4.917)--(10.920,4.917)--cycle;
\gpfill{rgb color={0.757,0.073,0.225}} (10.920,4.916)--(11.138,4.916)--(11.138,4.945)--(10.920,4.945)--cycle;
\gpfill{rgb color={0.750,0.068,0.220}} (10.920,4.944)--(11.138,4.944)--(11.138,4.973)--(10.920,4.973)--cycle;
\gpfill{rgb color={0.744,0.063,0.215}} (10.920,4.972)--(11.138,4.972)--(11.138,5.001)--(10.920,5.001)--cycle;
\gpfill{rgb color={0.737,0.058,0.210}} (10.920,5.000)--(11.138,5.000)--(11.138,5.029)--(10.920,5.029)--cycle;
\gpfill{rgb color={0.731,0.054,0.205}} (10.920,5.028)--(11.138,5.028)--(11.138,5.057)--(10.920,5.057)--cycle;
\gpfill{rgb color={0.725,0.049,0.200}} (10.920,5.056)--(11.138,5.056)--(11.138,5.085)--(10.920,5.085)--cycle;
\gpfill{rgb color={0.718,0.044,0.195}} (10.920,5.084)--(11.138,5.084)--(11.138,5.113)--(10.920,5.113)--cycle;
\gpfill{rgb color={0.712,0.039,0.190}} (10.920,5.112)--(11.138,5.112)--(11.138,5.141)--(10.920,5.141)--cycle;
\gpfill{rgb color={0.705,0.035,0.184}} (10.920,5.140)--(11.138,5.140)--(11.138,5.169)--(10.920,5.169)--cycle;
\gpfill{rgb color={0.699,0.030,0.179}} (10.920,5.168)--(11.138,5.168)--(11.138,5.197)--(10.920,5.197)--cycle;
\gpfill{rgb color={0.692,0.025,0.174}} (10.920,5.196)--(11.138,5.196)--(11.138,5.207)--(10.920,5.207)--cycle;
\gpfill{rgb color={0.690,0.024,0.172}} (10.920,5.206)--(11.138,5.206)--(11.138,5.225)--(10.920,5.225)--cycle;
\gpfill{rgb color={0.686,0.023,0.167}} (10.920,5.224)--(11.138,5.224)--(11.138,5.253)--(10.920,5.253)--cycle;
\gpfill{rgb color={0.678,0.022,0.159}} (10.920,5.252)--(11.138,5.252)--(11.138,5.281)--(10.920,5.281)--cycle;
\gpfill{rgb color={0.671,0.020,0.150}} (10.920,5.280)--(11.138,5.280)--(11.138,5.309)--(10.920,5.309)--cycle;
\gpfill{rgb color={0.664,0.019,0.142}} (10.920,5.308)--(11.138,5.308)--(11.138,5.337)--(10.920,5.337)--cycle;
\gpfill{rgb color={0.657,0.018,0.134}} (10.920,5.336)--(11.138,5.336)--(11.138,5.365)--(10.920,5.365)--cycle;
\gpfill{rgb color={0.650,0.017,0.125}} (10.920,5.364)--(11.138,5.364)--(11.138,5.393)--(10.920,5.393)--cycle;
\gpfill{rgb color={0.643,0.016,0.117}} (10.920,5.392)--(11.138,5.392)--(11.138,5.421)--(10.920,5.421)--cycle;
\gpfill{rgb color={0.636,0.015,0.109}} (10.920,5.420)--(11.138,5.420)--(11.138,5.449)--(10.920,5.449)--cycle;
\gpfill{rgb color={0.629,0.014,0.100}} (10.920,5.448)--(11.138,5.448)--(11.138,5.477)--(10.920,5.477)--cycle;
\gpfill{rgb color={0.622,0.013,0.092}} (10.920,5.476)--(11.138,5.476)--(11.138,5.505)--(10.920,5.505)--cycle;
\gpfill{rgb color={0.615,0.011,0.083}} (10.920,5.504)--(11.138,5.504)--(11.138,5.533)--(10.920,5.533)--cycle;
\gpfill{rgb color={0.608,0.010,0.075}} (10.920,5.532)--(11.138,5.532)--(11.138,5.561)--(10.920,5.561)--cycle;
\gpfill{rgb color={0.601,0.009,0.067}} (10.920,5.560)--(11.138,5.560)--(11.138,5.589)--(10.920,5.589)--cycle;
\gpfill{rgb color={0.594,0.008,0.058}} (10.920,5.588)--(11.138,5.588)--(11.138,5.617)--(10.920,5.617)--cycle;
\gpfill{rgb color={0.587,0.007,0.050}} (10.920,5.616)--(11.138,5.616)--(11.138,5.645)--(10.920,5.645)--cycle;
\gpfill{rgb color={0.580,0.006,0.042}} (10.920,5.644)--(11.138,5.644)--(11.138,5.673)--(10.920,5.673)--cycle;
\gpfill{rgb color={0.573,0.005,0.033}} (10.920,5.672)--(11.138,5.672)--(11.138,5.701)--(10.920,5.701)--cycle;
\gpfill{rgb color={0.566,0.003,0.025}} (10.920,5.700)--(11.138,5.700)--(11.138,5.729)--(10.920,5.729)--cycle;
\gpfill{rgb color={0.559,0.002,0.017}} (10.920,5.728)--(11.138,5.728)--(11.138,5.757)--(10.920,5.757)--cycle;
\gpfill{rgb color={0.552,0.001,0.008}} (10.920,5.756)--(11.138,5.756)--(11.138,5.785)--(10.920,5.785)--cycle;
\draw[gp path] (10.920,2.197)--(11.138,2.197)--(11.138,5.785)--(10.920,5.785)--cycle;
\draw[gp path] (11.138,2.197)--(10.958,2.197);
\node[gp node left] at (11.139,2.197) {0};
\draw[gp path] (10.920,2.197)--(11.100,2.197);
\draw[gp path] (11.138,2.557)--(10.958,2.557);
\node[gp node left] at (11.139,2.557) {1};
\draw[gp path] (10.920,2.557)--(11.100,2.557);
\draw[gp path] (11.138,3.393)--(10.958,3.393);
\node[gp node left] at (11.139,3.393) {10};
\draw[gp path] (10.920,3.393)--(11.100,3.393);
\draw[gp path] (11.138,4.589)--(10.958,4.589);
\node[gp node left] at (11.139,4.589) {$10^2$};
\draw[gp path] (10.920,4.589)--(11.100,4.589);
\draw[gp path] (11.138,5.785)--(10.958,5.785);
\node[gp node left] at (11.139,5.785) {$\ge10^3$};
\draw[gp path] (10.920,5.785)--(11.100,5.785);
\node[gp node center] at (8.574,5.426) {Unlabelled};
%% coordinates of the plot area
\gpdefrectangularnode{gp plot 2}{\pgfpoint{6.780cm}{2.197cm}}{\pgfpoint{10.368cm}{5.785cm}}
\end{tikzpicture}
%% gnuplot variables