### 2016-04-08 paper/paper.tex
Rework definitions, and reorder

% vim: set spell spelllang=en tw=100 et sw=4 sts=4 :

\documentclass{llncs}

% \usepackage{showframe}

\usepackage{microtype}                 % typographic perfection
\usepackage{complexity}                % \P, \NP etc
\usepackage{tikz}                      % For pretty pictures
\usepackage{amsmath}                   % \operatorname
\usepackage{amsfonts}                  % \mathbb
\usepackage{hyperref}                  % clicky links
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{gnuplot-lua-tikz}          % graphs
\usepackage[ruled,vlined]{algorithm2e} % algorithms (after cleverref!)
\usepackage{cite}                      % automatically sort mass cites
\usepackage[bottom]{footmisc}          % fix footnote + bottom figure on same page
\usepackage{placeins}

\usetikzlibrary{arrows, shadows, calc, positioning, decorations, decorations.pathreplacing,
    patterns, matrix}

% lncs style
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}
\crefname{figure}{Fig.}{Figs.}
\Crefname{figure}{Fig.}{Figs.}
\crefname{table}{Table}{Tables}
\Crefname{table}{Table}{Tables}
\crefname{proposition}{Proposition}{Propositions}
\Crefname{proposition}{Proposition}{Propositions}

% pretty colours
\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofgsunshine}{rgb}{1.0, 0.862745, 0.211765}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}

\newcommand{\lineref}[1]{line~\ref{#1}}
\newcommand{\twolinesref}[2]{lines~\ref{#1} and~\ref{#2}}

\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}

\renewcommand{\floatpagefraction}{0.8}
\renewcommand{\topfraction}{0.9}
\renewcommand{\bottomfraction}{0.8}
\newcommand{\christine}[1]{{\color{green} Christine:#1}}

\title{Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems}

\author{Ciaran McCreesh\thanks{This work was supported by the Engineering and Physical Sciences
    Research Council [grant number EP/K503058/1]}\inst{1} \and Samba Ndojh Ndiaye\thanks{This work
was supported by the ANR project SoLStiCe (ANR-13-BS02-0002-01)}\inst{2} \and Patrick
Prosser\inst{1} \and Christine Solnon\samethanks[2] \inst{3}}
\institute{University of Glasgow, Glasgow, Scotland \and
Universit\'e Lyon 1, LIRIS, UMR5205, F-69621, France  \and INSA-Lyon, LIRIS, UMR5205, F-69621, France}

\begin{document}

\maketitle

\begin{abstract}
    The maximum common subgraph problem is to find the largest subgraph common to two given graphs.
    This problem can be solved either by constraint-based search, or by reduction to the maximum
    clique problem. We evaluate these two models using modern algorithms, and see that the best
    choice depends mainly upon whether the graphs have labelled edges. We also investigate a
    restricted branching technique for the connected version of the problem, show how to implement
    this in constraint-based and clique-inspired algorithms, and compare it to conventional
    propagation.
\end{abstract}

\section{Introduction}

Maximum common subgraph problems arise in biology and chemistry
\cite{DBLP:journals/jcamd/RaymondW02a,Ehrlich:2011}, in computer vision
\cite{DBLP:journals/jair/CookH94}, in the analysis of source code
\cite{DBLP:journals/tkde/DjokoCH97}, binary programs \cite{DBLP:conf/icics/GaoRS08}, and circuit
designs \cite{DBLP:journals/jair/CookH94}, in character recognition problems \cite{SIWEILU1991617},
and in many other domains \cite{Shasha:2002:AAT:543613.543620}, both directly and as a way of
measuring the similarity or difference between two graphs
\cite{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,KriegeThesis}.

\paragraph{Definition of the Maximum Common (Connected) Subgraph Problem.}
We introduce definitions and algorithms on undirected and unlabelled graphs. The extension to directed or labelled graphs is straightforward and discussed in Section \ref{extension}. An undirected graph $G$ is defined by a finite set of nodes $N_G$ and a set of edges $E_G\subseteq N_G\times N_G$ such that each edge is an undirected pair of nodes.

Given a graph $G$, two nodes $n_s, n_e \in N_G$ are connected by a path in $G$ if there exists a sequence of nodes $<n_0,\ldots n_k>$ such that $n_0=n_s$, $n_k=n_e$ and each couple of successive nodes is connected by an edge, {\em i.e.}, $\forall i\in [1,k], \{n_{i-1},n_i\}\in E_G$. A graph $G$ is \emph{connected} if every pair of nodes is connected by a path.

A graph $G$ is \emph{isomorphic} to another graph $G'$ if there exists a bijective function $f:N_G\rightarrow N_{G'}$ which preserves edges, \emph{i.e.}, $\forall (u,v)\in N_G\times N_G, \{u,v\}\in E_G\Leftrightarrow \{f(u),f(v)\}\in E_{G'}$.

A {\em subgraph} is obtained by removing nodes, and all their incident edges, {\em i.e.}, $G'$ is a \emph{subgraph} of $G$ if $N_{G'}\subseteq N_G$ and $E_{G'}=E_G\cap(N_{G'}\times N_{G'})$.
In this definition, we only remove edges incident to removed nodes. This is usually called induced subgraph in the litterature. We may also allow to remove edges which are not incident to removed nodes, thus leading to partial subgraphs. In this paper, we only consider induced subgraphs. However, all algorithms may be extended to the partial case, as explained in \cite{DBLP:conf/cp/NdiayeS11}.

A \emph{common subgraph} of two graphs $G$ and $G'$ is a graph isomorphic to subgraphs of $G$ and $G'$. A \emph{common connected subgraph} is a common subgraph which is connected.
A \emph{Maximum Common Subgraph} (MCS) (resp. \emph{Maximum Common Connected Subgraph} (MCCS) is a common subgraph (resp. common connected subgraph) which has a maximum number of nodes. Examples of MCS and MCCS are given in Fig. \ref{figure:mcsexample}.

\begin{figure}[t] % first page if possible
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}[rotate=90]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!40] (M4) -- (M6);
            \draw [color=uofgsandstone!40] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!40] (M6) -- (M7);
            \draw [color=uofgsandstone!40] (M6) -- (M8);
            \draw [color=uofgsandstone!40] (M6) -- (M9);
            \draw [very thick] (M7) -- (M9);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=7cm, rotate=90]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!40] (M4) -- (M6);
            \draw [color=uofgsandstone!40] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!40] (M6) -- (M7);
            \draw [very thick] (M7) -- (M8);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=18cm, rotate=90]
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!40] (M6) -- (M7);
            \draw [very thick] (M6) -- (M8);
            \draw [color=uofgsandstone!40] (M6) -- (M9);
            \draw [color=uofgsandstone!40] (M7) -- (M9);
            \draw [color=uofgsandstone!40] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=25cm, rotate=90]
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!40, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [very thick] (M6) -- (M7);
            \draw [draw=uofgsandstone!40] (M7) -- (M8);
            \draw [draw=uofgsandstone!40] (M8) -- (M9);
        \end{scope}
\end{tikzpicture}
\caption{A maximum common induced subgraph of the first two graphs has eight vertices, shaded.
However, if we require the common subgraph to be connected, only seven vertices may be
selected---one way to do this is shown in the third and fourth graphs.}\label{figure:mcsexample}
\end{figure}

\paragraph{Overview of the paper.} In Section \ref{existing}, we review existing approaches for solving the MCS problem, with a specific focus on Constraint Programming (CP)-based approaches and on clique-based approaches that reduce the problem to finding a maximum clique in an association graph.

Conventional wisdom is that clique-based approaches are less effective in practice, but previous experimental evaluations have used weak clique algorithms, or even maximal clique enumeration algorithms. 
Therefore, in Section \ref{eval1}, we  re-evaluate the clique-based approach using a modern algorithm, and we show that it outperforms CP on labelled graphs, and that it is competitive with CP on non labelled graphs.

Then, in Section \ref{mccs}, we consider the MCCS problem. For the CP approach, we may add a global connectedness constraint to the model. Alternatively, we may use a special branching rule to grow connected subgraphs only, as proposed in \cite{DBLP:conf/mco/VismaraV08}. These two techniques may be combined, and we experimentally show that the best results are obtained when combining them. 
When solving the MCCS problem with a clique-based approach, neither technique seems directly viable with an association graph encoding.
However, we show that it is possible to adapt the combined branching and bounding rule used by modern clique algorithms to preserve connectedness during search.
We compare the clique-based approach with the best CP variant for the MCCS, and we show that it outperforms CP on non labelled graphs, whereas it is outperformed by CP on labelled graphs.

\section{Existing Complete Approaches for MCS}\label{existing}

There exist two main approaches for solving MCS: The first approach (described in \ref{CP}) is based on CP; The second approach (described in \ref{clique}) is based on a reformulation of MCS into a maximum clique problem. Both approaches are described for non directed and non labelled graphs. Their extension to directed or labelled graphs is discussed in Section \ref{extension}

\subsection{CP model for MCS}
\label{CP}

McGregor \cite{McGreg82} proposed a Branch \& Bound approach: Each branch of the search tree corresponds to the matching of two nodes, and a  bounding function evaluates the number of nodes that still may be matched so that the current branch is pruned as soon as this bound becomes lower than the size of the largest known common subgraph. CP approaches may be viewed as enhancements of this Branch \& Bound approach.

Vismara and Valery  \cite{DBLP:conf/mco/VismaraV08} introduced a first CP model. Given two graphs $G$ and $G'$, this model associates a variable $x_u$ with every node $u$ of $G$, and the domain of this variable contains all nodes of $G'$, plus an additional value $\bot$: variable $x_u$ is assigned to $\bot$ if node $u$ is not matched to any node of $G'$; otherwise $x_u$ is assigned to the node of $G'$ it is matched with. Edge constraints are introduced in order to ensure that variable assignments preserve edges between matched nodes, {\em i.e.}, 
$\forall \{u,v\}\subseteq N_G, (x_u=\bot)\vee (x_v=\bot) \vee (\{u,v\}\in E_G\Leftrightarrow \{x_u,x_v\}\in E_{G'})$. Difference constraints are introduced in order to ensure that each node of $G'$ is assigned to at most one variable, {\em i.e.}, $\forall \{u,v\}\subseteq N_G, (x_u=\bot)\vee (x_v=\bot) \vee (x_u\neq x_v)$.

This CP model is improved in \cite{DBLP:conf/cp/NdiayeS11} by replacing binary difference constraints with a soft global allDifferent constraint which maximizes the number of $x_u$ variables that are assigned to values different from $\bot$, while ensuring they are all different when they are not assigned to $\bot$.
Different constraint propagation techniques are compared in \cite{DBLP:conf/cp/NdiayeS11}. The combination ``MAC+Bound'' (resp. ``FC+Bound'') obtains the best results on labelled (resp. non labelled) graphs and outperforms the approach of \cite{McGreg82} and the CP model of \cite{DBLP:conf/mco/VismaraV08}. The combination ``MAC+Bound'' maintains arc consistency~\cite{sabi94} of edge constraints, whereas the conbination ``FC+Bound'' simply performs forward-checking on these constraints. In both combinations, the ``Bound'' filtering checks whether it is possible to assign distinct values to enough $x_u$ variables to surpass the best cost found so far (it is a weaker version of GAC(\emph{softAllDiff}) \cite{peti01} which computes the maximum number of variables that can be assigned distinct values).

\subsection{Reformulation of MCS as a maximum clique problem}
\label{clique}

We may solve MCS by introducing an association graph and searching for a maximum clique in it \cite{bala86,dura99,DBLP:journals/jcamd/RaymondW02a}.

An association graph of two graphs $G$ and $G'$ is an undirected graph $A_{G,G'}$ whose set of nodes is $N_{A_{G,G'}} = N_G \times N_{G'}$. To avoid confusing nodes of $A_{G,G'}$ with initial graph nodes, nodes of $A_{G,G'}$ are called matching nodes, as each node $(u,u')$ of $A_{G,G'}$ denotes the matching of $u$ with $u'$. Edges of $A_{G,G'}$ connect matching nodes which are compatible, {\em i.e.}, $E_{A_{G,G'}} = \big\{\{(u,u'),(v,v')\}\subseteq N_{A_{G,G'}}$, $(u,u')$ and $(v,v')$ are compatible$\big\}$, where two matching nodes $(u,u')$ and $(v,v')$ of $N_{A_{G,G'}}$ are compatible if $u \neq v$ and $u' \neq v'$, and if they preserve edges (\emph{i.e.}, $\{u,v\} \in E_G \Leftrightarrow \{u',v'\} \in E_{G'}$).

A \emph{clique} is a subgraph whose nodes are all linked pairwise.
A clique is \emph{maximal} if it is not strictly included in any other clique, and it is \emph{maximum} if it is the largest clique of a given graph, with respect to the number of nodes.

A clique in $A_{G,G'}$ corresponds to a set of compatible matchings. Therefore, such a clique corresponds to a common subgraph, and a maximum clique of $A_{G,G'}$ is a MCS of $G$ and $G'$.
It follows that any method able to find a maximum clique in a graph can be used to solve the MCS problem.

Note that the association graph is a partial subgraph of the microstructure \cite{DBLP:conf/aaai/Jegou93a} associated with the CP model of  \cite{DBLP:conf/mco/VismaraV08}: The microstructure has more matching nodes than the association graph because it has a matching node $(u,\bot)$ for each node $u$ of $G$. Each clique of size $|N_G|$ in the microstructure corresponds to a common subgraph the size of which is defined by the number of nodes that do not contain $\bot$.

\subsection{Extension to Labeled or Directed Graphs}\label{extension}

In some applications, nodes and edges are associated with labels. We note $\lambda(u)$ and $\lambda(\{u,v\})$ the label of a node $u$ and an edge $\{u,v\}$, respectively. In this case, the common  subgraph must match components the labels of which are compatible, and we note $\simeq$ the compatibility relationship, {\em i.e.}, labels $\lambda_1$ and $\lambda_2$ are compatible if $\lambda_1\simeq\lambda_2$. This kind of label compatibility constraint is handled in a straightforward way in both CP and clique-based approaches:
\begin{itemize}
\item For CP, we restrict the domain of every variable $x_u$ to nodes with compatible labels, {\em i.e.}, $\forall u\in N_G, D(x_u) = \{\bot\}\cup \{u'\in N_{G'}, \lambda(u)\simeq\lambda(u')\}$. We also have to ensure that edge labels are preserved in edge constraints, {\em i.e.}, $\forall \{u,v\}\subseteq N_G, (x_u=\bot)\vee (x_v=\bot) \vee ((\{u,v\}\in E_G\Leftrightarrow \{x_u,x_v\}\in E_{G'}) \wedge (\lambda(\{u,v\}) \simeq \lambda(\{x_u,x_v\})))$.

\item For clique-based approaches, label compatibility is handled through the definition of the association graph, by restricting the set of matching nodes to pairs of nodes with compatible labels, {\em i.e.}, $N_{G_C} = \{(u,u')\in N_G\times N_{G'}, \lambda(u)\simeq\lambda(u')\}$. Also, edges of the association graph are restricted to pairs $\{(u,u'),(v,v')\}$ such that $\lambda(\{u,v\}) \simeq \lambda(\{u',v')\}$.
\end{itemize} 
Labels usually simplify the solution process, both for CP and clique-based approaches: Node labels reduce domain sizes for CP, and the number of matching nodes in the association graphs; Edge labels tighten edge constraints for CP, and make the association graph sparser for clique-based approaches.

The extension of MCS algorithms to directed graphs is straightforward: For CP, edge orientation is handled in edge constraints; for clique-based approaches, it is handled in the definition of the association graph edges. In both cases, adding orientations to edges usually simplifies the problem: It tightens edge constraints for CP, and it makes the association graph sparser for  clique-based approaches.

\section{Re-Evaluating the Clique Model}\label{eval1}

Previous experimental work has used a maximal clique enumeration algorithm, even for the
maximisation problem. ?? check this! I recall someone using a bad max clique algorithm too.
\cite{DBLP:conf/sspr/BunkeFGSV02,DBLP:journals/jgaa/ConteFV07}

We instead use a maximum clique algorithm.

Also sparse stuff, but association graphs aren't sparse enough.

Say which clique algorithm we're using.

\subsection{Datasets}

\cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}

?? More? Can we scale to some of the SIP graphs?

\subsection{Experimental Evaluation}

?? Break this down more by family, size, etc.

\begin{figure}[p]
    \centering
    \input{gen-graph-unconnected-cumulative}
    \vspace*{1em}

    \centering
    \input{gen-graph-unconnected-heatmap}
    \caption{The cumulative number of instances solved in under a certain time: on the top, 33\%
        labelled graphs with up to 100 vertices, and in the middle, unlabelled and undirected graphs
        with up to 35 vertices. On the bottom row, an instance-by-instance comparison of the two
        models, with 33\% labelled graphs on the left, and unlabelled and undirected graphs on
        the right.} \label{figure:unconnected-cumulative}
\end{figure}

\subsection{Filtering from Edge Labels}

\subsection{Parallel Search}

\cite{DBLP:journals/jcc/KoncDTRJ12,DBLP:journals/algorithms/McCreeshP13,DBLP:journals/topc/McCreeshP15,DBLP:journals/cor/SegundoLP16}

?? Stuff on EHPs, how work stealing matters here too \cite{DBLP:journals/jco/BatsynGMP14}

\section{Find Maximum Common Connected Subgraphs}\label{mccs}

Connectedness is not hereditary, but can be built up one vertex at a time. Alternatively,
connectedness is a constraint.

\subsection{Connectedness as a Constraint}

Connectedness \cite{Brown:2005}

Reachability \cite{DBLP:conf/cp/DoomsDD05,DBLP:conf/cp/QuesadaRD05}

\subsection{Connectedness via Restricted Branching}

\begin{figure}[tb]
    \newcounter{stepcounter}
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [draw=uofgsandstone!40] (Ma) -- (Mb);
            \draw [draw=uofgsandstone!40] (Ma) -- (Mc);
            \draw [draw=uofgsandstone!40] (Mb) -- (Mc);
            \draw [draw=uofgsandstone!40] (Mb) -- (Md);
            \draw [draw=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [color=uofgsandstone!40] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [color=uofgsandstone!40] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Initial problem };
        \end{scope}

        \begin{scope}[xshift=13cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [draw=uofgsandstone!40] (Ma) -- (Mb);
            \draw [draw=uofgsandstone!40] (Ma) -- (Mc);
            \draw [draw=uofgsandstone!40] (Mb) -- (Mc);
            \draw [dotted] (Mb) -- (Md);
            \draw [dotted] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [color=uofgsandstone!40] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [color=uofgsandstone!40] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Trying $a \mapsto 1$ };
        \end{scope}

        \begin{scope}[xshift=26cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [very thick] (Ma) -- (Mb);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Md);
            \draw [color=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [very thick] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [color=uofgsandstone!40] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Trying $b \mapsto 2$ };
        \end{scope}

        \begin{scope}[yshift=-9cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, dotted, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [very thick] (Ma) -- (Mb);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Mc);
            \draw [very thick] (Mb) -- (Md);
            \draw [color=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [very thick] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [very thick] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) Trying $d \mapsto 4$ };
        \end{scope}

        \begin{scope}[xshift=13cm, yshift=-9cm]
            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md) at (0, 0) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Me) at (2, 0) {\phantom{0}};

            \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [left = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [right = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [left = 0 of Md, font=\small] { \vphantom{0}d };
            \node [right = 0 of Me, font=\small] { \vphantom{0}e };

            \draw [very thick] (Ma) -- (Mb);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Mb) -- (Mc);
            \draw [very thick] (Mb) -- (Md);
            \draw [color=uofgsandstone!40] (Mc) -- (Me);

            \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (M1) at (5.6, 4) {\phantom{0}};
            \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (M2) at (7.6, 4) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (5.6, 2) {\phantom{0}};
            \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (M4) at (7.6, 2) {\phantom{0}};
            \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (5.6, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
            \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

            \draw [very thick] (M1) -- (M2);
            \draw [color=uofgsandstone!40] (M1) -- (M3);
            \draw [very thick] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);

            \node [anchor=base, yshift=-2.2cm] at ($(Ma)!0.5!(M2)$) {
                \stepcounter{stepcounter}\roman{stepcounter}) $e \mapsto \bot$ is forced };
        \end{scope}
    \end{tikzpicture}

    \caption{Suppose we are looking for a connected common subgraph, using the graph on the left
        for variables and the graph on the right (which has an isolated vertex) for values. We
        initially consider $a \mapsto 1$. Our restricted branching rule requires us to select either variable
        $b$ or variable $c$ subsequently, not $d$ or $e$. We try $b \mapsto 2$, which adds $d$
        to the branchable variables, and forces $c \mapsto \bot$. We may now only branch on $d$, and
        we try $d \mapsto 4$. Now the only remaining variable is unbranchable, and so $e = \bot$ is forced, even
        though $5$ remains in its domain and does not violate any constraints.}\label{figure:restricted}
\end{figure}

\cref{figure:restricted} The key idea is that whenever we branch (except for the first choice we
make), we select only from variables which correspond to vertices which are adjacent to at least one
other variable which has been instantiated to a value other than $\bot$. If no such variables
remain, the subgraph we are building cannot grow any further, and all remaining variables must be
instantiated to $\bot$.

?? Patrick reckons this reminds him of Prim's algorithm for minimum spanning trees.

\subsection{Connectedness in an Association Graph Algorithm}

\begin{algorithm}[p] \DontPrintSemicolon
\nl $\FuncSty{associationMCCIS}$ :: (Graph $G_1$, Graph $G_2$) $\rightarrow$ Map \;
\nl \Begin{
    \nl $\KwSty{global} ~ \mathit{incumbent} \gets \emptyset$ \label{incumbent} \;
    \nl $G \gets \FuncSty{associationGraph}(G_1, G_2)$ \label{buildassoc} \;
    \nl $\FuncSty{search}(G, \emptyset, \emptyset, \operatorname{V}(G))$ \label{firstsearch} \;
    \nl $\KwSty{return} ~ \mathit{incumbent}$ \;
}
\vspace{0.9ex}
\nl $\FuncSty{search}$ :: (Graph $G$, Set $\mathit{solution}$, Set $\mathit{branchable}$, Set $\mathit{remaining}$) \;
\nl \Begin{
    \nl $\mathit{colourClasses} \gets \FuncSty{concatenate}($ \label{makecolours} \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \setminus \mathit{branchable}),$ \;
    $\hspace*{2em}\FuncSty{colour}(G, \mathit{remaining} \cap \mathit{branchable}))$ \;
    \nl \While{$\FuncSty{length}(\mathit{colourClasses}) > 0$\label{outerloop}}{
        \nl \ForEach{$v \in \FuncSty{last}(\mathit{colourClasses}) \textnormal{~in reverse order}$\label{innerloop}}{
            \nl \lIf{$|\mathit{solution}| + \FuncSty{length}(\mathit{colourClasses}) \leq
            |\mathit{incumbent}|$}{\KwSty{return}\label{bound}}
            \nl \lIf{$v \notin \mathit{branchable} ~ \KwSty{and} ~ \mathit{solution} \ne
            \emptyset$}{\KwSty{return} \label{acceptbranchable}}
            \nl $\mathit{solution}' \gets \mathit{solution} \cup \{ v \} \label{addv} \label{addvtosolution} $\;
            \nl \lIf{$|\mathit{solution}'| > |\mathit{incumbent}|$}{$\mathit{incumbent} \gets
            \mathit{solution}'$ \label{newincumbent}}
            \nl $\mathit{branchable}' \gets \mathit{branchable} \cup \FuncSty{connected}(G, v)$ \label{addtobranchable} \;
            \nl $\mathit{remaining}' \gets \mathit{remaining} \cap \operatorname{N}(G, v)$ \label{filterremaining} \;
            \nl \lIf{$\mathit{remaining}' \neq \emptyset$}{$\FuncSty{search}(G, \mathit{solution}',
            \mathit{branchable}', \mathit{remaining}')$\label{recursesearch}}
         }
         \nl $\FuncSty{removeLast}(\mathit{colourClasses})$ \;
     }
}
\vspace{0.9ex}
\nl $\FuncSty{colour}$ :: (Graph $G$, Set $\mathit{uncoloured}$) $\rightarrow$ List of List of Vertex \;
\nl \Begin{
    \nl $\mathit{result} \gets []$ \;
    \nl \While{$\mathit{uncoloured} \ne \emptyset$\label{whileuncoloured}}{
        \nl $\mathit{colourClass} \gets []$ \label{newcolourclass} \;
        \nl $\mathit{colourable} \gets \mathit{uncoloured}$ \;
        \nl \While{$\mathit{colourable} \ne \emptyset$}{
            \nl $v \gets \textnormal{the first vertex from}~\mathit{uncoloured}~\textnormal{by
            static degree order from}~G$ \label{selectvtocolour} \;
            \nl $\FuncSty{append}(\mathit{colourClass}, v)$ \label{addtocolourclass} \;
            \nl $uncoloured \gets uncoloured \setminus \{ v \}$ \;
            \nl $\mathit{colourable} \gets \mathit{colourable} \setminus \operatorname{N}(G, v)$ \;
        }
        \nl $\FuncSty{append}(\mathit{result}, \mathit{colourClass})$ \;
    }
    \nl $\KwSty{return} ~ \mathit{result}$ \;
}
\vspace{0.9ex}
\nl $\FuncSty{connected}$ :: (Graph $G$, Vertex $v$) $\rightarrow$ Set \;
\nl \Begin{
    \nl $\KwSty{return} ~ \{ w \in G : \FuncSty{first}(w) \in \operatorname{N}(G, \FuncSty{first}(v)) \}$ \;
}
\vspace{0.9ex}
\nl $\FuncSty{associationGraph}$ :: (Graph $G_1$, Graph $G_2$) $\rightarrow$ Graph \;
\nl \Begin{
    \nl $V \gets \{ (v_1, v_2) \in G_1 \times G_2 : v_1 \simeq v_2 \}$ \;
    \nl $E \gets \{ ((v_1, v_2), (w_1, w_2)) \in V \times V : (v_1, w_1) \simeq (v_2, w_2) \}$ \;
    \nl $\KwSty{return} ~ (V, v \mapsto 1, (v, w) \mapsto 1~\KwSty{if}~(v, w) \in E~\KwSty{otherwise}~\bot )$ \;
}
\caption{An algorithm for a maximum common connected induced subgraph isomorphism via an association graph.}
\label{algorithm:mccis}
\end{algorithm}

\begin{figure}[tb]
    \newcounter{stepcounter2}
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \matrix (m) [matrix of nodes, column sep=0.5cm] {
            \begin{tikzpicture}[scale=0.33]
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Ma) at (0, 4) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mb) at (2, 4) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mc) at (0, 2) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Md) at (2, 2) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Me) at (0, 0) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (Mf) at (2, 0) {\phantom{0}};

                    \node [left = 0 of Ma, font=\small] { \vphantom{0}a };
                    \node [right = 0 of Mb, font=\small] { \vphantom{0}b };
                    \node [left = 0 of Mc, font=\small] { \vphantom{0}c };
                    \node [right = 0 of Md, font=\small] { \vphantom{0}d };
                    \node [left = 0 of Me, font=\small] { \vphantom{0}e };
                    \node [right = 0 of Mf, font=\small] { \vphantom{0}f };

                    \draw (Ma) -- (Mb);
                    \draw (Ma) -- (Mc);
                    \draw (Ma) -- (Md);
                    \draw (Mc) -- (Md);
                    \draw (Mc) -- (Me);
                    \draw (Md) -- (Mf);

                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M1) at (6, 4) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M2) at (8, 4) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M3) at (6, 2) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M4) at (8, 2) {\phantom{0}};
                    \node[draw, circle, fill=white, inner sep=0.5pt, font=\normalsize] (M5) at (6, 0) {\phantom{0}};

                    \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
                    \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
                    \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
                    \node [right = 0 of M4, font=\small] { \vphantom{0}4 };
                    \node [left = 0 of M5, font=\small] { \vphantom{0}5 };

                    \draw [color=uofgsandstone!40] (M1) -- (M2);
                    \draw [color=uofgsandstone!40] (M1) -- (M3);
                    \draw [color=uofgsandstone!40] (M2) -- (M4);
                    \draw [color=uofgsandstone!40] (M3) -- (M4);
            \end{tikzpicture}
            &
            \begin{tikzpicture}[scale=0.33]
                    \node [anchor=west] (Ns) at (0, 3.5) { $\mathit{solution}$ };
                    \node [anchor=west] (Nb) at (0, 2) { $\mathit{branchable}$ };
                    \node [anchor=west] (Nr) at (0, 0.5) { $\mathit{remaining}$ };

                    \node [anchor=west] (Vs) at (5.5, 3.5) { $\{ (a, 1) \}$ };
                    \node [anchor=west] (Vb) at (5.5, 2) { $\{ b, c, d \} \times \{1 \ldots 5 \}$ };
                    \node [anchor=west] (Vr) at (5.5, 0.5) { $(\{ b, c, d \} \times \{ 2, 3 \}) \cup (\{ e, f\} \times \{ 4, 5 \})$ };
            \end{tikzpicture}
            \\[0.1cm]
            \stepcounter{stepcounter2}\roman{stepcounter2}) Initial problem &
            \stepcounter{stepcounter2}\roman{stepcounter2}) Search variables after guessing $a \mapsto 1$
            \\[0.4cm]
            \begin{tikzpicture}[scale=0.33]
                    \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Me4) at (0, 3) {\phantom{0}};
                    \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Me5) at (3, 3) {\phantom{0}};
                    \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mf4) at (0, 0) {\phantom{0}};
                    \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mf5) at (3, 0) {\phantom{0}};

                    \node [above = 0 of Me4, font=\small] { \vphantom{0}(e, 4) };
                    \node [above = 0 of Me5, font=\small] { \vphantom{0}(e, 5) };
                    \node [below = 0 of Mf4, font=\small] { \vphantom{0}(f, 4) };
                    \node [below = 0 of Mf5, font=\small] { \vphantom{0}(f, 5) };

                    \draw (Me4) -- (Mf5);
                    \draw (Me5) -- (Mf4);
            \end{tikzpicture}
            &
            \begin{tikzpicture}[scale=0.33]
                    \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (Mb2) at (3, 3) {\phantom{0}};
                    \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (Mb3) at (6, 3) {\phantom{0}};
                    \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Mc2) at (0, 0) {\phantom{0}};
                    \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Mc3) at (3, 0) {\phantom{0}};
                    \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md2) at (6, 0) {\phantom{0}};
                    \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md3) at (9, 0) {\phantom{0}};

                    \node [above = 0 of Mb2, font=\small] { \vphantom{0}(b, 2) };
                    \node [above = 0 of Mb3, font=\small] { \vphantom{0}(b, 3) };
                    \node [below = 0 of Mc2, font=\small] { \vphantom{0}(c, 2) };
                    \node [below = 0 of Mc3, font=\small] { \vphantom{0}(c, 3) };
                    \node [below = 0 of Md2, font=\small] { \vphantom{0}(d, 2) };
                    \node [below = 0 of Md3, font=\small] { \vphantom{0}(d, 3) };

                    \draw (Mb2) -- (Mc3);
                    \draw (Mb2) -- (Md3);
                    \draw (Mb3) -- (Mc2);
                    \draw (Mb3) -- (Md2);
            \end{tikzpicture}
            \\[0.1cm]
            \stepcounter{stepcounter2}\roman{stepcounter2}) $\mathit{remaining} \setminus \mathit{branchable}$ &
            \stepcounter{stepcounter2}\roman{stepcounter2}) $\mathit{remaining} \cap \mathit{branchable}$
            \\
        };
    \end{tikzpicture}\\[0.3cm]\begin{tikzpicture}[scale=0.33]
        \matrix (m) [matrix of nodes] {
            \begin{tikzpicture}[scale=0.33]
                \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Me4) at (0, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgsunshine, inner sep=0.5pt, font=\normalsize] (Me5) at (3, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mf4) at (6, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgcobalt, inner sep=0.5pt, font=\normalsize] (Mf5) at (9, 0) {\phantom{0}};
                \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (Mb2) at (12, 0) {\phantom{0}};
                \node[draw, circle, fill=uofglawn, inner sep=0.5pt, font=\normalsize] (Mb3) at (15, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Mc2) at (18, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Mc3) at (21, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md2) at (24, 0) {\phantom{0}};
                \node[draw, circle, fill=uofgthistle, inner sep=0.5pt, font=\normalsize] (Md3) at (27, 0) {\phantom{0}};

                \node [above = 0 of Me4, font=\small, xshift=-0.5ex] { \vphantom{0}[[(e, 4), };
                \node [above = 0 of Me5, font=\small] { \vphantom{0}(e, 5)], };
                \node [above = 0 of Mf4, font=\small] { \vphantom{0}[(f, 4), };
                \node [above = 0 of Mf5, font=\small] { \vphantom{0}(f, 5)], };
                \node [above = 0 of Mb2, font=\small] { \vphantom{0}[(b, 2), };
                \node [above = 0 of Mb3, font=\small] { \vphantom{0}(b, 3)], };
                \node [above = 0 of Mc2, font=\small] { \vphantom{0}[(c, 2), };
                \node [above = 0 of Mc3, font=\small] { \vphantom{0}(c, 3), };
                \node [above = 0 of Md2, font=\small] { \vphantom{0}(d, 2), };
                \node [above = 0 of Md3, font=\small] { \vphantom{0}(d, 3)]]};
            \end{tikzpicture}
            \\[0.1cm]
            \stepcounter{stepcounter2}\roman{stepcounter2}) The resulting $\mathit{colourClasses}$ variable.
            \\
        };
    \end{tikzpicture}

    \caption{Solving a maximum common connected problem using an association graph. Suppose we
        have already mapped vertex $a$ to vertex $1$, giving the assignments on the right. Now we
        have two subgraphs to colour. We need two colours for $\mathit{remaining} \setminus
        \mathit{branchable}$, and we place these two colour classes first in the $\mathit{colourClasses}$
        variable. We can also colour $\mathit{remaining} \cap \mathit{branchable}$ using two
        colours, since we cannot simultaneously map $c$ to $2$ and $d$ to $3$, or vice-versa. Thus
        $\mathit{colourClasses}$ becomes a list of four colour classes, the first three containing
        two vertices each, and the last containing four vertices. This tells us that if we hope to
        extend the current common subgraph by another four vertices, we
        must pick one assignment from each of the four colour classes (which is not actually possible, so we
    see the bound here gives an overestimate). The algorithm thus guesses $d \mapsto 3$ as its next
assignment, and if that fails, $d \mapsto 2$, and so on; once $b \mapsto 3$ is reached, the bound
decreases by one, and if $f \mapsto 5$ were reached we would stop due to a lack of remaining
branchable vertices. }\label{figure:assocrestricted}
\end{figure}

It is not possible to determine connectedness from a raw association graph. However, we can take a
maximum clique algorithm and mimic the branching strategy if we have access to the underlying graphs
and can determine the ``meaning'' of the association graph vertices.

Most modern maximum clique algorithms for dense graphs use some variation of greedy graph colouring
as a bound---the underlying observation is that each vertex in a clique must be given a different
colour in a colouring, so if we can colour a subset of vertices using $k$ colours then a maximum
clique in this subset has at most $k$ vertices. However, the colouring is also used as a branching
heuristic: vertices are selected in reverse from their colour classes in turn, starting with the
last colour class created. Because of this coupling of branching and the bound (which is important
in practice because it mimics a ``smallest domain first'' branching heuristic if colour classes are
viewed as variables \cite{DBLP:conf/cp/McCreeshP14}), if we were to select only a subset of vertices for
branching at each stage inside a clique algorithm, we would lose completeness. Thus we must adapt
the bound to take into account restricted branching.

In \cref{algorithm:mccis} we present a clique-inspired algorithm which finds a maximum common
connected induced subgraph isomorphism via an
association graph. If the additional
branching restrictions are removed, the core of the algorithm is what Prosser
\cite{DBLP:journals/algorithms/Prosser12} calls the ``MCSa1'' variant of a series of algorithms due
to Tomita \textit{et al.}\
\cite{DBLP:conf/dmtcs/TomitaS03,DBLP:journals/jgo/TomitaK07,DBLP:conf/walcom/TomitaSHTW10}, using a
bitset encoding due to San Segundo \textit{et al.}\
\cite{DBLP:journals/cor/SegundoRJ11,DBLP:journals/ol/SegundoMRH13} (and we refer the reader to these
papers for implementation details on how to use bitsets and other data structures to implement the
colouring stage with very low constant factors).

We begin by building the association graph (\lineref{buildassoc}). The main part of the algorithm
then works by building up candidate cliques in the $\mathit{solution}$ variable, by recursive calls
to the $\mathit{search}$ procedure---starting from the empty set (\lineref{firstsearch}), each
recursive subcall adds one vertex to $\mathit{solution}$ (\lineref{addv}) in such a way that
$\mathit{solution}$ is always a clique which corresponds to a connected common subgraph. The
$\mathit{remaining}$ set contains the set of vertices which are adjacent to every vertex in
$\mathit{solution}$, and which have not yet been accepted or rejected (and so initially it contains
every vertex). The main loops in the $\FuncSty{search}$ procedure
(\twolinesref{outerloop}{innerloop}) have the effect of iterating over each vertex in this set in a
particular order---each vertex $v$ is selected in turn, and then a recursive call is made to
consider the effects of including $v$ in $\mathit{solution}$ (\lineref{recursesearch}), followed by
the next iteration where $v$ is instead rejected. When $v$ is accepted, we add it to the new
$\mathit{solution'}$ (\lineref{addvtosolution}), and create a new $\mathit{remaining'}$ containing
only the vertices in $\mathit{remaining}$ which are adjacent to $v$ (\lineref{filterremaining}).

The $\mathit{branchable}$ set contains the set of association graph vertices which correspond to
vertices adjacent to an already-accepted vertex in the first input graph---in constraint programming terms, it is the
set of assignments which could be made next which preserve connectedness. (Using only one of the two
input graphs is sufficient for correctness, and has the advantage that the $\FuncSty{connected}$
function may be implemented as a simple lookup into a precomputed array which maps each vertex in
the first input graph to a bitset.) At the top of search, this set is empty, and is not used (our
first vertex selection is special, and does not care about connectedness). At subsequent depths, we
may only accept vertices which are in this set, and if no such vertices remain then we return
immediately (\lineref{acceptbranchable}). When recursing, we extend $\mathit{branchable}$ with the
new vertices permitted by our acceptance of the branching $v$ (\lineref{addtobranchable}). Note that
we are assuming that inside the main loops, we encounter every vertex in $\mathit{remaining} \cap
\mathit{connected}$ before any vertex in $\mathit{remaining} \setminus \mathit{connected}$.

As we proceed, we keep track of the best solution we have found so far---this is stored in the
$\mathit{incumbent}$ variable (\twolinesref{incumbent}{newincumbent}). We use the incumbent,
together with a colour bound, to prune portions of the search space which cannot contain a better
solution. The colour bound operates as follows: at each entry to the $\FuncSty{search}$ procedure,
we produce a greedy colouring of the vertices in $\mathit{remaining}$ (\lineref{makecolours}). This
greedy colouring gives us a list of colour classes, each of which is a list of pairwise non-adjacent
vertices. The two loops (\twolinesref{outerloop}{innerloop}) then iterate over each colour class,
from last to first, and then over each vertex in that colour class, again from last to first.
(Rather than actually using a list of lists and removing items, this process should be implemented
using a pair of immutable flat arrays. This technique is described elsewhere
\cite{DBLP:conf/cp/McCreeshP14}, so we do not discuss it here.) Finally, if at any point the number
of remaining colour classes plus the number of vertices currently present in $\mathit{solution}$ is
not strictly greater than the size of the incumbent, then we may backtrack immediately
(\lineref{bound}).

Finally, we describe the colouring process. In conventional clique algorithms, a simple greedy
sequential colouring is used (possibly with the help of previous colourings to reduce the
computational cost \cite{DBLP:conf/lion/NikolaevBS15}, and possibly with shortcuts taken for certain
vertices \cite{DBLP:journals/cor/SegundoT14}, and possibly followed by a repair step to
improve the colouring \cite{DBLP:conf/walcom/TomitaSHTW10}, or stronger bounding rules based upon
MaxSAT inference \cite{DBLP:conf/ictai/LiFX13,DBLP:conf/lion/LiJX15,DBLP:journals/cor/SegundoNB15}).
Such colourings will not give us the required property that vertices in $\mathit{remaining} \cap
\mathit{connected}$ come last (so they are selected first by the reverse branching order). Thus we
produce two greedy sequential colourings, first considering the non-branching vertices in
$\mathit{remaining} \setminus \mathit{connected}$, followed by the branching vertices, and
concatenate them (\lineref{makecolours}). This produces a valid colouring, since we do not merge any
colour classes between the two stages, although it may use more colours than a single colouring
would\footnote{What if we did not guarantee that vertices in $\mathit{remaining} \cap
    \mathit{connected}$ came last, and just used a conventional colouring with the branching rule?
    Suppose we had four vertices in $\mathit{remaining}$, and produced a colouring $[[v_1, v_2],
    [v_3], [v_4]]$, and suppose that extending $\mathit{solution}$ with $\{ v_1, v_3, v_4 \}$ gives
an optimal solution. If $v_4$ was not $\mathit{connected}$ yet, we would not branch on that subtree,
and the bound could eliminate branching on $v_3$ and $v_1$, so we would miss the solution. Thus we
cannot simply add the branching rule without also adapting the combined bound / ordering heuristic.}.

Our $\FuncSty{colour}$ procedure is the same as the bit-parallel algorithm introduced by San Segundo
\textit{et al.}\ \cite{DBLP:journals/cor/SegundoRJ11}. While we have remaining vertices to colour
(\lineref{whileuncoloured}), we start a new colour class (\lineref{newcolourclass}), and then
repeatedly pick a legal vertex to add to that colour class (\lineref{addtocolourclass}).  The
selection of the next vertex to colour (\lineref{selectvtocolour}) may be performed efficiently in
hardware if the association graph is permuted to be in degree order at the top of search. Other
initial vertex orderings have been considered on general clique problems
\cite{DBLP:journals/algorithms/Prosser12,DBLP:conf/lion/SegundoLB14}; it is possible that special
properties of the association graph could be exploited in this step.

\subsection{Experimental Evaluation}

\paragraph{Connected, Undirected, 33\% Labelled}

\paragraph{Connected, Undirected, Unlabelled}

?? My gut feeling is that if we have edge labels, the clique model wins, and if we don't, it's
likely to lose. So maybe I should run some experiments with vertex labels but not edge labels to
test this.

\paragraph{By family} ?? Break this down more by family, size, etc.

\paragraph{Does connected make instances easier or harder?} We could plot connected vs not connected
difficulties, and result sizes?

\begin{figure}[p]
    \centering
    \input{gen-graph-connected-cp.tex}
    \caption{A pairwise comparison of different CP techniques. The colour at a point indicates the
    number of instances with these runtimes, with darker colours representing more instances. The
    top two rows are for 33\% labelled undirected graphs with up to 100 vertices, and the bottom two
    rows are for unlabelled and undirected graphs with up to 35 vertices.}
        \label{figure:connected-cp}
\end{figure}

\begin{figure}[p]
    \centering
    \input{gen-graph-connected-cumulative}
    \vspace*{1em}

    \centering
    \input{gen-graph-connected-heatmap}
    \caption{The cumulative number of connected instances solved in under a certain time: on the
        top, 33\% labelled undirected graphs with up to 100 vertices, and in the middle, unlabelled
        and undirected graphs with up to 35 vertices. On the bottom row, an instance-by-instance
        comparison of the association and ``MAC both'' approaches, with 33\% labelled graphs on the
        left, and unlabelled and undirected graphs on the right.} \label{figure:connected-cumulative}
\end{figure}

\subsection{Parallel Search}

\cite{DBLP:conf/ictai/MinotNS15}

\section{Conclusion}

Some immoral perspectives on closely coupling branching with a particular constraint. Whether this
is useful for other things, like ASP.

We haven't discussed partial subgraphs, or connected directed.

Possibly portfolios.

Better instances.

\FloatBarrier
\bibliographystyle{splncs}
\bibliography{paper}

\end{document}